---
title: "binary outcome study: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## How well do these estimators recover heterogeneous treatment effects?

Ten scenarios vary the structure of the CATE, crossed with four sample sizes
(100, 250, 500, 1000), 100 runs each. Every dataset also carries five
deliberately unrelated covariates (`X01`–`X05`), so the estimators have to
find the signal rather than being handed it. Full design in
`binary/README.md`.

| Scenario | CATE structure |
|---|---|
| 1: No HTE | $b_0 + b_1 X_1 + b_2 X_2 + W b_W$ |
| 2: Binary HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3)$ |
| 3: Continuous HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 X_4)$ |
| 4: Two HTE vars | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4)$ |
| 5: Cts x binary interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{34} X_3 X_4)$ |
| 6: Two vars + cts x binary | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{34} X_3 X_4)$ |
| 7: Cts x cts interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{45} X_4 X_5)$ |
| 8: Two vars + cts x cts | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$ |
| 9: Cosine HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_4 \cos(X_4))$ |
| 10: Exponential HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 \exp(-\lvert X_4 \rvert))$ |

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle`, `dr_superlearner` — all using double crossfitting for the
nuisances (`R/cate_models.R`). The oracle uses the true outcome model and a
known propensity of 0.5; the semi-oracle knows only the propensity. The
SuperLearner library shrinks at `n = 100` (`SL.earth`/`SL.ranger` are
dropped), so `dr_superlearner` is sometimes filtered out of the smallest-`n`
figures below.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "binary")

metrics <- readRDS(file.path(res_path, "bin_metrics.RDS"))

# display order and labels - binary has no shared label-dictionary script
# (unlike crossfitting/cf_metrics.R, which does double duty as pipeline +
# labels), so these stay inline, matching bin_results.Rmd's precedent
scenario_labels <- c(
  `1` = "1: No HTE",
  `2` = "2: Binary HTE",
  `3` = "3: Continuous HTE",
  `4` = "4: Two HTE vars",
  `5` = "5: Cts x binary interaction",
  `6` = "6: Two vars + cts x binary",
  `7` = "7: Cts x cts interaction",
  `8` = "8: Two vars + cts x cts",
  `9` = "9: Cosine HTE",
  `10` = "10: Exponential HTE"
)

model_labels <- c(
  causal_forest = "Causal forest",
  dr_random_forest = "DR-RandomForest",
  dr_oracle = "DR-oracle",
  dr_semi_oracle = "DR-semi-oracle",
  dr_superlearner = "DR-SuperLearner"
)

# tidy up
metrics <- metrics %>%
  # temporary filter for 4 scenarios in thesis
  filter(scenario %in% c(1,3,8,9)) %>%
  mutate(
    scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                      labels = unname(scenario_labels)),
    n = factor(n, levels = c(100, 250, 500, 1000)),
    model = factor(recode(model, !!!model_labels), levels = unname(model_labels))
  ) %>%
  droplevels()

# per (scenario, model, n) summaries
# mcse denominator is sqrt(sum(!is.na(x))), matching the non-NA count of that
# column - not sqrt(n()), the whole group's row count regardless of NAs.
# BLP_p/indep_cate/indep_po are NA_real_ whenever the HTE test wasn't run
# (R/metrics.R::hte_test_metrics()), so those columns are the most exposed.
metrics_summary <- metrics %>%
  group_by(scenario, model, n) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(sum(!is.na(bias))),
    mean_ate_bias = mean(ate_bias, na.rm = T),
    mcse_ate_bias = sd(ate_bias, na.rm = T) / sqrt(sum(!is.na(ate_bias))),
    mean_rel_ate_bias = mean(rel_ate_bias, na.rm = T),
    mcse_rel_ate_bias = sd(rel_ate_bias, na.rm = T) / sqrt(sum(!is.na(rel_ate_bias))),
    mean_rel_bias_cate = mean(rel_bias_cate, na.rm = T),
    mcse_rel_bias_cate = sd(rel_bias_cate, na.rm = T) / sqrt(sum(!is.na(rel_bias_cate))),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(sum(!is.na(mse))),
    mean_rmse = mean(rmse, na.rm = T),
    mcse_rmse = sd(rmse, na.rm = T) / sqrt(sum(!is.na(rmse))),
    mean_mae = mean(mae, na.rm = T),
    mcse_mae = sd(mae, na.rm = T) / sqrt(sum(!is.na(mae))),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(sum(!is.na(corr))),
    mean_spearman = mean(spearman, na.rm = T),
    mcse_spearman = sd(spearman, na.rm = T) / sqrt(sum(!is.na(spearman))),
    mean_sign_acc = mean(sign_acc, na.rm = T),
    mcse_sign_acc = sd(sign_acc, na.rm = T) / sqrt(sum(!is.na(sign_acc))),
    mean_BLP = mean(BLP_p, na.rm = T),
    mcse_BLP = sd(BLP_p, na.rm = T) / sqrt(sum(!is.na(BLP_p))),
    mean_indep_cate = mean(indep_cate, na.rm = T),
    mcse_indep_cate = sd(indep_cate, na.rm = T) / sqrt(sum(!is.na(indep_cate))),
    mean_indep_po = mean(indep_po, na.rm = T),
    mcse_indep_po = sd(indep_po, na.rm = T) / sqrt(sum(!is.na(indep_po))),
    # power: proportion of runs rejecting at alpha=0.05 - a genuine 0/1
    # indicator, so it gets the binomial MCSE, unlike the mean p-values above
    power_BLP = mean(BLP_p < 0.05, na.rm = T),
    mcse_power_BLP = sqrt(power_BLP * (1 - power_BLP) / sum(!is.na(BLP_p))),
    power_indep_cate = mean(indep_cate < 0.05, na.rm = T),
    mcse_power_indep_cate = sqrt(power_indep_cate * (1 - power_indep_cate) / sum(!is.na(indep_cate))),
    power_indep_po = mean(indep_po < 0.05, na.rm = T),
    mcse_power_indep_po = sqrt(power_indep_po * (1 - power_indep_po) / sum(!is.na(indep_po))),
    mean_n_na = mean(n_na, na.rm = T),
    total_n_na = sum(n_na, na.rm = T),
    .groups = "drop"
  )
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with a 95% CI (mean +/- qnorm(1 -
# alpha/2) x MCSE) error bar, faceted by scenario - see
# continuous/cts_results.R's summary_plot for why this isn't a raw +/- 1x
# MCSE (~68% coverage, not 95%)
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = 0, alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - z * .data[[mcse_col]],
           hi = .data[[mean_col]] + z * .data[[mcse_col]]) %>%
    ggplot(aes(x = n, y = est, colour = model, ymin = lo, ymax = hi)) +
    geom_hline(yintercept = hline, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_wrap(~scenario, scales = "free") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Sample size", colour = "Model")
}
```

## Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per
run, against the known truth.

```{r bias-plot, fig.height=16}
bias_plot <- metrics %>%
  ggplot(aes(x = n, y = bias, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bias in CATE estimates", y = "Bias", x = "Sample size", colour = "Model")

bias_sum_plot <- summary_plot(metrics_summary, "mean_bias", "mcse_bias",
                              "Mean bias in CATE estimates", "Mean bias")

(bias_plot / bias_sum_plot) + plot_layout(guides = "collect")
```

## ATE bias

Bias of the *average* treatment effect (`mean(est) - mean(true)`), as
opposed to Bias's per-unit average — the two can differ in sign or magnitude
when errors don't average out evenly across units.

```{r ate-bias-plot, fig.height=16}
ate_bias_plot <- metrics %>%
  ggplot(aes(x = n, y = ate_bias, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bias in the average treatment effect", y = "ATE bias",
       x = "Sample size", colour = "Model")

ate_bias_sum_plot <- summary_plot(metrics_summary, "mean_ate_bias", "mcse_ate_bias",
                                  "Mean ATE bias", "Mean ATE bias")

(ate_bias_plot / ate_bias_sum_plot) + plot_layout(guides = "collect")
```

## Relative bias

`rel_ate_bias` is the parameter-level relative bias (ATE bias divided by the
true ATE); `rel_bias_cate` is the per-unit version, `(est - true) / true`
averaged over units (undefined and set to `NA` wherever the true CATE is
exactly 0 for a unit — see `R/metrics.R::cate_metrics()`).

```{r rel-bias-plot, fig.height=16}
rel_ate_bias_sum_plot <- summary_plot(metrics_summary, "mean_rel_ate_bias",
                                      "mcse_rel_ate_bias",
                                      "Mean relative bias in the ATE",
                                      "Mean relative ATE bias")
rel_bias_cate_sum_plot <- summary_plot(metrics_summary, "mean_rel_bias_cate",
                                       "mcse_rel_bias_cate",
                                       "Mean relative bias in the CATE",
                                       "Mean relative CATE bias")

rel_ate_bias_sum_plot
rel_bias_cate_sum_plot
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth, for every model and sample size.

```{r mse-plot, fig.height=16}
mse_plot <- metrics %>%
  ggplot(aes(x = n, y = mse, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "MSE of CATE estimates", y = "MSE", x = "Sample size", colour = "Model")

mse_sum_plot <- summary_plot(metrics_summary, "mean_mse", "mcse_mse",
                             "Mean MSE of CATE estimates", "Mean MSE")

(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
```

## RMSE of CATE estimates

`sqrt(mse)` — the same ranking as MSE, shown on the outcome's own scale
rather than squared.

```{r rmse-plot, fig.height=16}
rmse_plot <- metrics %>%
  ggplot(aes(x = n, y = rmse, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "RMSE of CATE estimates", y = "RMSE", x = "Sample size", colour = "Model")

rmse_sum_plot <- summary_plot(metrics_summary, "mean_rmse", "mcse_rmse",
                              "Mean RMSE of CATE estimates", "Mean RMSE")

(rmse_plot / rmse_sum_plot) + plot_layout(guides = "collect")
```

## MAE of CATE estimates

Mean absolute error — a robustness check on MSE, less sensitive to outlying
runs.

```{r mae-plot, fig.height=16}
mae_plot <- metrics %>%
  ggplot(aes(x = n, y = mae, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "MAE of CATE estimates", y = "MAE", x = "Sample size", colour = "Model")

mae_sum_plot <- summary_plot(metrics_summary, "mean_mae", "mcse_mae",
                             "Mean MAE of CATE estimates", "Mean MAE")

(mae_plot / mae_sum_plot) + plot_layout(guides = "collect")
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in scenario 1 since there's no heterogeneity to correlate with —
excluded from these plots).

```{r corr-plot, fig.height=16}
corr_df <- filter(metrics, scenario != scenario_labels[["1"]])

corr_plot <- corr_df %>%
  ggplot(aes(x = n, y = corr, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  # free_y only: sample size is shared across panels, only the correlation
  # scale needs freeing per scenario
  facet_wrap(~scenario, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation between estimated and true CATEs",
       y = "Pearson correlation", x = "Sample size", colour = "Model")

corr_sum_plot <- summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
                              "mean_corr", "mcse_corr",
                              "Mean correlation with the true CATE", "Mean correlation")

(corr_plot / corr_sum_plot) + plot_layout(guides = "collect")
```

## Spearman rank correlation with the true CATE

Rank correlation alongside the Pearson correlation above — less sensitive to
outlying CATE estimates and to non-linear-but-monotone miscalibration. Same
scenario-1-undefined convention, so scenario 1 is excluded here too.

```{r spearman-plot, fig.height=16}
spearman_plot <- corr_df %>%
  ggplot(aes(x = n, y = spearman, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Spearman rank correlation between estimated and true CATEs",
       y = "Spearman correlation", x = "Sample size", colour = "Model")

spearman_sum_plot <- summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
                                  "mean_spearman", "mcse_spearman",
                                  "Mean Spearman correlation with the true CATE",
                                  "Mean Spearman correlation")

(spearman_plot / spearman_sum_plot) + plot_layout(guides = "collect")
```

## Sign accuracy

Proportion of units where the estimated CATE has the correct sign — a
direction-detection metric, distinct from magnitude accuracy. 0.5 (dashed
line) is the chance level for a coin-flip on sign.

```{r sign-acc-plot, fig.height=16}
sign_acc_plot <- metrics %>%
  ggplot(aes(x = n, y = sign_acc, colour = model)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Sign accuracy of CATE estimates", y = "Proportion correct sign",
       x = "Sample size", colour = "Model")

sign_acc_sum_plot <- summary_plot(metrics_summary, "mean_sign_acc", "mcse_sign_acc",
                                  "Mean sign accuracy", "Mean proportion correct sign",
                                  hline = 0.5)

(sign_acc_plot / sign_acc_sum_plot) + plot_layout(guides = "collect")
```

## True-CATE HTE test evaluation

The BLP and independence tests re-run on the true CATE and true nuisances
instead of an estimator's fitted ones (`bin_true_cate_tests.RDS` —
`run_true_cate_tests()`/`true_cate_test_row()`, `R/cate_models.R`) — the
ceiling each model's own BLP/CATE-permutation test is chasing, since nothing
here is estimated. Shown as an extra "True CATE" series alongside the five
models in the two sections below (and in "HTE test power"); no true-CATE
counterpart for the PO permutation test, since `dr_oracle`'s own `indep_po`
already tests the true pseudo-outcome.

```{r true-cate-data}
true_cate_tests <- readRDS(file.path(res_path, "bin_true_cate_tests.RDS")) %>%
  mutate(
    scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                      labels = unname(scenario_labels)),
    n = factor(n, levels = c(100, 250, 500, 1000))
  ) %>%
  droplevels()

true_cate_summary <- true_cate_tests %>%
  group_by(scenario, n) %>%
  summarise(
    mean_BLP = mean(BLP_p, na.rm = T),
    mcse_BLP = sd(BLP_p, na.rm = T) / sqrt(sum(!is.na(BLP_p))),
    mean_indep_cate = mean(indep_cate, na.rm = T),
    mcse_indep_cate = sd(indep_cate, na.rm = T) / sqrt(sum(!is.na(indep_cate))),
    power_BLP = mean(BLP_p < 0.05, na.rm = T),
    mcse_power_BLP = sqrt(power_BLP * (1 - power_BLP) / sum(!is.na(BLP_p))),
    power_indep_cate = mean(indep_cate < 0.05, na.rm = T),
    mcse_power_indep_cate = sqrt(power_indep_cate * (1 - power_indep_cate) / sum(!is.na(indep_cate))),
    .groups = "drop"
  ) %>%
  mutate(model = "True CATE")

true_cate_raw <- true_cate_tests %>% mutate(model = "True CATE")

model_levels_true <- c(levels(metrics$model), "True CATE")

# row-bind the true-CATE reference series onto a per-run or summarised data
# frame, extending `model`'s levels so it becomes one more colour/facet entry
bind_true_cate <- function(df, true_df, extra_cols) {
  bind_rows(
    mutate(df, model = factor(as.character(model), levels = model_levels_true)),
    true_df %>%
      mutate(model = factor(model, levels = model_levels_true)) %>%
      select(scenario, n, model, all_of(extra_cols))
  )
}
```

## BLP test p-values

Best linear predictor test of heterogeneity, against the nominal 5%
significance threshold (dashed line) — where the CATE truly varies, low
p-values indicate the model detects it; in scenario 1 (no HTE), p-values
should sit uniform on [0, 1]. "True CATE" has no defined BLP test in scenario
1: the true CATE is exactly constant there, and `GenericML::BLP()` cannot
identify its interaction coefficient when tau has zero variance.

```{r blp-plot, fig.height=16}
BLP_plot <- bind_true_cate(metrics, true_cate_raw, "BLP_p") %>%
  ggplot(aes(x = n, y = BLP_p, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "BLP test p-values", y = "p-value", x = "Sample size", colour = "Model")

BLP_sum_plot <- summary_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_BLP", "mcse_BLP")),
  "mean_BLP", "mcse_BLP", "Mean BLP test p-value", "Mean p-value", hline = 0.05)

(BLP_plot / BLP_sum_plot) + plot_layout(guides = "collect")
```

## CATE permutation test p-values

```{r indep-cate-plot, fig.height=16}
indep_cate_plot <- bind_true_cate(metrics, true_cate_raw, "indep_cate") %>%
  ggplot(aes(x = n, y = indep_cate, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "CATE permutation test p-values", y = "p-value", x = "Sample size",
       colour = "Model")

indep_cate_sum_plot <- summary_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_indep_cate", "mcse_indep_cate")),
  "mean_indep_cate", "mcse_indep_cate",
  "Mean CATE permutation test p-value", "Mean p-value", hline = 0.05)

(indep_cate_plot / indep_cate_sum_plot) + plot_layout(guides = "collect")
```

## PO permutation test p-values

```{r indep-po-plot, fig.height=16}
indep_po_plot <- metrics %>%
  ggplot(aes(x = n, y = indep_po, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "PO permutation test p-values", y = "p-value", x = "Sample size",
       colour = "Model")

indep_po_sum_plot <- summary_plot(metrics_summary, "mean_indep_po", "mcse_indep_po",
                                  "Mean PO permutation test p-value", "Mean p-value",
                                  hline = 0.05)

(indep_po_plot / indep_po_sum_plot) + plot_layout(guides = "collect")
```

## HTE test power

Power (proportion of runs rejecting at alpha=0.05) alongside the mean
p-values above — the rsimsum-standard measure for a hypothesis test's
simulated behaviour, with the binomial 95% CI rather than the continuous-
measure one.

```{r power-plot, fig.height=16}
BLP_power_plot <- summary_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("power_BLP", "mcse_power_BLP")),
  "power_BLP", "mcse_power_BLP", "BLP test power", "Power", hline = 0.05)
indep_cate_power_plot <- summary_plot(
  bind_true_cate(metrics_summary, true_cate_summary,
                c("power_indep_cate", "mcse_power_indep_cate")),
  "power_indep_cate", "mcse_power_indep_cate",
  "CATE permutation test power", "Power", hline = 0.05)
indep_po_power_plot <- summary_plot(metrics_summary, "power_indep_po",
                                    "mcse_power_indep_po",
                                    "PO permutation test power", "Power",
                                    hline = 0.05)

BLP_power_plot
indep_cate_power_plot
indep_po_power_plot
```

## Missing estimates

`n_na` is a data-quality diagnostic, not a performance metric: the count of
`NA` CATE estimates within a run. Mostly zero, so shown as a table rather
than a distribution plot; only cells with at least one missing estimate are
listed.

```{r na-table}
na_table <- metrics_summary %>%
  select(scenario, model, n, mean_n_na, total_n_na) %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, n, model)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("Scenario", "Model", "n", "Mean NA count per run",
                      "Total NA count"))
} else {
  cat("No missing CATE estimates in any run.")
}
```

## Headline table

Arranged by scenario, sample size, and mean MSE (best model first within
each cell).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, model, n, mean_bias, mean_ate_bias, mean_rel_ate_bias,
         mean_rel_bias_cate, mean_mse, mean_rmse, mean_mae,
         mean_sign_acc, mean_corr, mean_BLP, mean_indep_cate, mean_indep_po,
         power_BLP, power_indep_cate, power_indep_po) %>%
  arrange(scenario, n, mean_mse)

kable(headline, digits = 4,
      col.names = c("Scenario", "Model", "n", "Mean bias", "Mean ATE bias",
                    "Mean relative ATE bias", "Mean relative CATE bias",
                    "Mean MSE", "Mean RMSE", "Mean MAE", "Mean sign accuracy",
                    "Mean correlation", "Mean BLP p-value",
                    "Mean CATE perm. p-value", "Mean PO perm. p-value",
                    "BLP power", "CATE perm. power", "PO perm. power"))
```
