---
title: "Binary outcome results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(paletteer)
library(here)
library(cowplot)
library(purrr)

# paths
path <- here()
res_path <- file.path(dirname(here()), "results", "new_format")
```

```{r data, include=FALSE}
# data
metrics <- readRDS(file.path(res_path, "metrics_binary_tidy.RDS"))

metrics <- metrics %>%
  mutate(scenario = factor(scenario, levels = c(1:10)),
         n = factor(n, levels = c(100, 250, 500, 1000)),
         model = factor(
           recode(model,
                  causal_forest = "Causal forest",
                  dr_random_forest = "DR-RandomForest",
                  dr_oracle = "DR-oracle",
                  dr_semi_oracle = "DR-semi-oracle",
                  dr_superlearner = "DR-SuperLearner"),
           levels = c("Causal forest", "DR-RandomForest", "DR-oracle", "DR-semi-oracle", "DR-SuperLearner")))

# per scenario summaries
metrics_summary <- metrics %>%
  group_by(scenario, n, model) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T)/sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T)/sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T)/sqrt(n()),
    mean_BLP = mean(BLP_p, na.rm = T),
    mcse_BLP = sd(BLP_p, na.rm = T)/sqrt(n()),
    mean_indep = mean(indep_p, na.rm = T),
    mcse_indep = sd(indep_p, na.rm = T)/sqrt(n()),
    .groups = "drop"
  )
```

```{r plot_generatiom, include=FALSE}

make_all_plots <- function(scen) {
  bias_plot <- metrics %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = bias, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Bias distribution, scenario ", scen),
         y = "Bias",
         x = "Sample size") +
    theme(legend.position = "none")
  
  bias_sum_plot <- metrics_summary %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mean_bias, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_bias - mcse_bias,
                      ymax = mean_bias + mcse_bias),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean bias in CATEs, scenario ", scen),
         y = "Bias",
         x = "sample_size") +
    theme(legend.position = "none")
  
  mse_plot <- metrics %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mse, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("MSE distribution, scenario ", scen),
         y = "MSE",
         x = "Sample size") +
    theme(legend.position = "none")
  
  mse_sum_plot <- metrics_summary %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mean_mse, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_mse - mcse_mse,
                      ymax = mean_mse + mcse_mse),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean MSE in CATEs, scenario ", scen),
         y = "MSE",
         x = "sample_size") +
    theme(legend.position = "none")
  
  corr_plot <- metrics %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = corr, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Correlation distribution, scenario ", scen),
         y = "Correlation",
         x = "Sample size") +
    theme(legend.position = "none")
  
  corr_sum_plot <- metrics_summary %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mean_corr, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_corr - mcse_corr,
                      ymax = mean_corr + mcse_corr),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean correlation in CATEs, scenario ", scen),
         y = "Correlation",
         x = "sample_size") +
    theme(legend.position = "none")
  
  BLP_plot <- metrics %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = BLP_p, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("BLP p-values distribution, scenario ", scen),
         y = "p-value",
         x = "Sample size") +
    theme(legend.position = "none")
  
  BLP_sum_plot <- metrics_summary %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mean_BLP, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_BLP - mcse_BLP,
                      ymax = mean_BLP + mcse_BLP),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("mean BLP p-value, scenario ", scen),
         y = "p-value",
         x = "sample_size") +
    theme(legend.position = "none")
  
  indep_plot <- metrics %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = indep, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("PO permutation test p-value distribution, scenario ", scen),
         y = "p-value",
         x = "Sample size") +
    theme(legend.position = "none")
  
  indep_sum_plot <- metrics_summary %>%
    filter(scenario == scen) %>%
    ggplot(aes(x = n, y = mean_indep, colour = model)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_indep - mcse_indep,
                      ymax = mean_indep + mcse_indep),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean PO permutation test p-value, scenario ", scen),
         y = "p-value",
         x = "sample_size") +
    theme(legend.position = "none")
  
  legend <- get_legend(bias_plot + theme(legend.position = "right"))
  
  list(bias_plot = bias_plot, bias_sum_plot = bias_sum_plot, mse_plot = mse_plot,
       mse_sum_plot = mse_sum_plot, corr_plot = corr_plot, corr_sum_plot = corr_sum_plot,
       BLP_plot = BLP_plot, BLP_sum_plot = BLP_sum_plot, indep_plot = indep_plot,
       indep_sum_plot = indep_sum_plot, legend = legend)
}

scenarios <- unique(metrics$scenario)
all_plots <- map(scenarios, make_all_plots)
```


# Scenario 1 - no HTE
$b_0 + b_1 X_1 + b_2 X_2 + W*b_W$
```{r}
scenario <- 1
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 2 - simply HTE with a binary variable
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3)$
```{r}
scenario <- 2
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 3 - simple HTE with a continuous variable
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 X_4)$
```{r}
scenario <- 3
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 4 - two HTE variables
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4)$
```{r}
scenario <- 4
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 5 - cts-binary interaction in HTE variables
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{34} X_3 X_4)$
```{r}
scenario <- 5
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 6 - 2 HTE variables + cts-binary interaction
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{34} X_3 X_4)$
```{r}
scenario <- 6
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 7 - cts-cts interaction in HTE variables
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{45} X_4 X_5)$
```{r}
scenario <- 7
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 8 - 2 HTE variables + cts-cts interaction
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$
```{r}
scenario <- 8
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 9 - Cosine HTE
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_4 \cos(X_4))$
```{r}
scenario <- 9
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

# Scenario 10 - Exponential HTE
$b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 \exp(-\lvert X_4 \rvert))$

```{r}
scenario <- 10
plot_grid(all_plots[[scenario]]$bias_plot,
          all_plots[[scenario]]$bias_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$mse_plot,
          all_plots[[scenario]]$mse_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$corr_plot,
          all_plots[[scenario]]$corr_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$BLP_plot,
          all_plots[[scenario]]$BLP_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_cate_plot,
          all_plots[[scenario]]$indep_cate_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))

plot_grid(all_plots[[scenario]]$indep_plot,
          all_plots[[scenario]]$indep_sum_plot,
          all_plots[[scenario]]$legend,
          ncol = 3,
          rel_widths = c(1, 1, 0.2))
```

