---
title: "Single-event survival: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 11,
  fig.height = 7
)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(knitr)
```

## Pseudo-values, causal survival forest and RSF DR-learner on one event

One time-to-event outcome, distributed as event 1 of the competing-risk study
with event 2 removed (Weibull shape 2, scale 25, X1 and X2 prognostic, X3 the
effect modifier). The estimand is the CATE on RMST to day 28, which is the
competing-risk study's net RMST1 (`tau_RMST1_cs`).

| Scenario | Treatment effect | True τ, mean (SD) |
|---|---|---|
| 1 | none | 0 (0) |
| 2 | constant on the hazard scale, HR 1.42 | −2.16 (0.10) |
| 3 | heterogeneous, HR 1.42 / 2.86 at X3 = 0 / 1 | −5.24 (2.02) |

n = 500, censoring on (uniform + administrative at 180) or off (administrative
only), 500 runs. Design and arms in `ADEMP.md`.

- **Scenario 2's correlation and C-statistic** are measured against a truth
  with SD 0.10, all from X1 and X2. Read bias and RMSE there.
- **Scenario 1** has τ ≡ 0: correlation and C-statistic are set to 0 and 0.5
  by `cate_metrics()`. `tau_sd`, the within-run spread of the estimates, is the
  spurious heterogeneity each arm reports.
- **Without censoring** the pseudo-values are exactly min(Y, 28), so the
  pseudo-value arms are regressions on the observed truncated time.

```{r data}
res_path <- file.path(dirname(here()), "results", "single_event")
metrics <- readRDS(file.path(res_path, "se_metrics.RDS"))

scenario_labels <- c(`1` = "Null", `2` = "Constant", `3` = "Heterogeneous")
learner_levels <- c("CSF", "CF", "DR", "T")
nuisance_levels <- c("grf", "RF", "SL", "RSF", "RSF (scf)")

metrics <- metrics %>%
  mutate(
    scenario = factor(scenario, levels = 1:3, labels = unname(scenario_labels)),
    censoring = factor(censoring, levels = c(TRUE, FALSE),
                       labels = c("Censored", "Uncensored")),
    learner = factor(learner, levels = learner_levels),
    nuisance = factor(nuisance, levels = nuisance_levels)
  )

mcse <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

metrics_summary <- metrics %>%
  group_by(scenario, censoring, learner, nuisance, framework) %>%
  summarise(
    across(c(bias, ate_bias, rmse, corr, c_stat, tau_sd),
           list(mean = ~ mean(.x, na.rm = TRUE), mcse = mcse)),
    total_n_na = sum(n_na, na.rm = TRUE),
    .groups = "drop"
  )
```

### Results coverage

3,000 runs in all (3 scenarios × 2 censoring settings × 500 runs). Check this
before reading anything below.

```{r completeness-table}
metrics %>%
  group_by(scenario, censoring) %>%
  summarise(n_runs = n_distinct(run), n_arms = n_distinct(framework),
            n_na = sum(n_na), .groups = "drop") %>%
  kable(col.names = c("Scenario", "Censoring", "Runs collected",
                      "Arms present", "Missing estimates"))
```

```{r summary-plot-helper, include=FALSE}
# points with mean +/- 1.96 MCSE, coloured by the nuisance model, faceted
# scenario x censoring
summary_plot <- function(measure, title, ylab, hline = 0, scenarios = NULL) {
  df <- metrics_summary
  if (!is.null(scenarios)) df <- filter(df, scenario %in% scenarios)
  df %>%
    mutate(est = .data[[paste0(measure, "_mean")]],
           lo = est - 1.96 * .data[[paste0(measure, "_mcse")]],
           hi = est + 1.96 * .data[[paste0(measure, "_mcse")]]) %>%
    ggplot(aes(x = learner, y = est, ymin = lo, ymax = hi, colour = nuisance)) +
    { if (!is.null(hline)) geom_hline(yintercept = hline, linetype = "dashed") } +
    geom_point(position = position_dodge(width = 0.6), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.6), width = 0.3,
                  linewidth = 0.3) +
    facet_grid(censoring ~ scenario, scales = "free_y") +
    labs(title = title, x = NULL, y = ylab, colour = "Nuisance model") +
    theme_bw() +
    theme(legend.position = "bottom")
}
```

### Bias

```{r bias}
summary_plot("bias", "Mean CATE bias", "Bias (days)")
```

### RMSE

```{r rmse}
summary_plot("rmse", "RMSE against the true CATE", "RMSE (days)", hline = NULL)
```

### Spurious heterogeneity in the null

```{r tau-sd}
summary_plot("tau_sd", "Within-run SD of the estimated CATE (truth: 0)",
             "SD (days)", scenarios = "Null")
```

### Ranking (heterogeneous scenario)

```{r corr}
summary_plot("corr", "Correlation with the true CATE", "Pearson r",
             hline = NULL, scenarios = "Heterogeneous")
```

```{r c-stat}
summary_plot("c_stat", "C-statistic", "C", hline = 0.5,
             scenarios = "Heterogeneous")
```

### Table

```{r table}
metrics_summary %>%
  select(scenario, censoring, framework, bias_mean, rmse_mean, corr_mean,
         c_stat_mean, tau_sd_mean) %>%
  arrange(scenario, censoring) %>%
  kable(digits = 3)
```
