---
title: "Simulation Updates"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_width: 10
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(ggplot2)
library(paletteer)
library(gridExtra)
library(dplyr)
library(tidyr)
library(purrr)

res_path <- "/rds/general/user/evanvogt/projects/nihr_drf_simulations/live/results/"
```

```{r bin_results, include=FALSE}
cate_corr_bin <- readRDS(paste0(res_path, "binary/all/cate_correlations.RDS"))

cate_mse_bin <- readRDS(paste0(res_path, "binary/all/cate_mses.RDS"))

cate_coverage_bin <- readRDS(paste0(res_path, "binary/all/cate_coverage.RDS"))

cate_bias_bin <- readRDS(paste0(res_path, "binary/all/cate_bias.RDS"))

ci_length_bin <- readRDS(paste0(res_path, "binary/all/ci_length.RDS"))

BLP_df_bin <- readRDS(paste0(res_path, "binary/all/BLP_tests_metrics.RDS"))
```

```{r cont_results, include=FALSE}
cate_corr_cont <- readRDS(paste0(res_path, "continuous/all/cate_correlations.RDS"))

cate_mse_cont <- readRDS(paste0(res_path, "continuous/all/cate_mses.RDS"))

cate_coverage_cont <- readRDS(paste0(res_path, "continuous/all/cate_coverage.RDS"))

cate_bias_cont <- readRDS(paste0(res_path, "continuous/all/cate_bias.RDS"))

ci_length_cont <- readRDS(paste0(res_path, "continuous/all/ci_length.RDS"))

BLP_df_cont <- readRDS(paste0(res_path, "continuous/all/BLP_tests_metrics.RDS"))
```

```{r functions, include=FALSE}
create_long_df <- function(nested_list, summary_name) {
  map_dfr(names(nested_list), function(scenario) {
    map_dfr(names(nested_list[[scenario]]), function(n) {
      map_dfr(names(nested_list[[scenario]][[n]]), function(model) {
        values <- nested_list[[scenario]][[n]][[model]]
        df <- data.frame(
          scenario = as.numeric(gsub("scenario_", "", scenario)), 
          n = as.factor(n),
          model = model,
          sim = seq_along(values),
          value = values
        )
        names(df)[names(df) == "value"] <- summary_name
        return(df)
      })
    })
  })
}
filter_df <- function(df) {
  df <- df %>%
          mutate(
            model = ifelse(model == "CF", "causal forest",
                           ifelse(model == "DR_RF", "DR-learner",
                                  ifelse(model == "T_RF", "T-learner",
                                         ifelse(model == "DR_oracle", "Oracle DR-learner", NA))))
          )
  df <- df %>%
    mutate(
      model = factor(model, levels = c("causal forest", "T-learner", "DR-learner", "Oracle DR-learner"))
    )
  return(df)
}
```

```{r df_manipulation_bin, include=FALSE}
mse_df_bin <- create_long_df(cate_mse_bin, "mse")
coverge_df_bin <- create_long_df(cate_coverage_bin, "coverage")
bias_df_bin <- create_long_df(cate_bias_bin, "bias")
length_df_bin <- create_long_df(ci_length_bin, "length")

mse_df_bin <- filter_df(mse_df_bin)
coverage_df_bin <- filter_df(coverge_df_bin)
bias_df_bin <- filter_df(bias_df_bin)
length_df_bin <- filter_df(length_df_bin)
```

```{r df_manipulation_cont, include=FALSE}
mse_df_cont <- create_long_df(cate_mse_cont, "mse")
coverge_df_cont <- create_long_df(cate_coverage_cont, "coverage")
bias_df_cont <- create_long_df(cate_bias_cont, "bias")
length_df_cont <- create_long_df(ci_length_cont, "length")

mse_df_cont <- filter_df(mse_df_cont)
coverage_df_cont <- filter_df(coverge_df_cont)
bias_df_cont <- filter_df(bias_df_cont)
length_df_cont <- filter_df(length_df_cont)
```

```{r plotting functions, include=FALSE}

plot_violin <- function(df, var, title) {
  ggplot(df, aes(x = n, y = .data[[var]], colour = model, fill = model)) +
    geom_violin(alpha = 0.7, trim = FALSE) +
    facet_wrap(~scenario, nrow = 2) + 
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      axis.ticks.x = element_blank()
    ) +
    labs(
      title = title,
      x = "Sample size",
      y = var
    ) + 
    scale_color_paletteer_d("PrettyCols::Dark") +
    scale_fill_paletteer_d("PrettyCols::Dark")
}

plot_box <- function(df, var, title) {
  ggplot(df, aes(x = n, y = .data[[var]], colour = model, fill = model)) +
    geom_boxplot(alpha = 0.7) +
    facet_wrap(~scenario, nrow = 2) + 
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      axis.ticks.x = element_blank()
    ) +
    labs(
      title = title,
      x = "Sample size",
      y = var
    ) + 
    scale_color_paletteer_d("PrettyCols::Dark") + 
    scale_fill_paletteer_d("PrettyCols::Dark")
}

plot_average <- function(df, var, title) {
  measure <- sym(var)
  summary_df <- df %>%
  group_by(scenario, n, model) %>%
  summarise(
    mean_perf = mean(!!measure, na.rm = TRUE),
    mcse_perf = sd(!!measure, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
  summary_df <- summary_df %>% mutate(
    lb = mean_perf - 1.96*mcse_perf,
    ub = mean_perf + 1.96*mcse_perf
  )
  ggplot(summary_df, aes(x = n, y = mean_perf, colour = model)) +
    geom_point() + 
    geom_segment(aes(x = n, xend = n, y = lb, yend = ub)) +
    facet_wrap(~scenario, nrow = 2) + 
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
      axis.ticks.x = element_blank()
    ) +
    labs(
      title = title,
      x = "sample size",
      y = var
    ) + 
    scale_color_paletteer_d("PrettyCols::Dark") +
    scale_fill_paletteer_d("PrettyCols::Dark")
}
```

# Binary outcome

## Mean-squared Error of true CATE vs predicted CATE

```{r bin_MSE}
plot_box(mse_df_bin, "mse", "")
plot_violin(mse_df_bin, "mse", "")
plot_average(mse_df_bin, "mse", "")
```

## average bias of true CATE vs predicted CATE

```{r bin_bias}
plot_box(bias_df_bin, "bias", "")
plot_violin(bias_df_bin, "bias", "")
plot_average(bias_df_bin, "bias", "")
```

## coverage of true CATE vs predicted CATE

```{r bincov}
plot_box(coverage_df_bin, "coverage", "")
plot_violin(coverage_df_bin, "coverage", "")
plot_average(coverage_df_bin, "coverage", "")
```

## average CI length

```{r bin_ci}
plot_box(length_df_bin, "length", "")
plot_violin(length_df_bin, "length", "")
plot_average(length_df_bin, "length", "")
```

```{r BLP_bin_df, include=FALSE}
BLP_df_bin <- BLP_df_bin %>% mutate(
  TP = TP/1000,
  FP = FP/1000,
  TN = TN/1000,
  FN = FN/1000)

BLP_df_bin <- BLP_df_bin %>%
  mutate(agreement = ifelse(scenario == "scenario_1", TN, TP),
         disagreement = ifelse(scenario == "scenario_1", FP, FN)
  )
```

## BLP results

```{r - BLP_bin}
compact_bin <- BLP_df_bin %>%
  select(model, n, scenario, agreement) %>%
  pivot_wider(
    names_from = scenario,
    values_from = agreement
  ) %>%
  arrange(model, n)

knitr::kable(compact_bin)
```

# Continuous outcome

## Mean-squared Error of true CATE vs predicted CATE

```{r cont_MSE}
plot_box(mse_df_cont, "mse", "")
plot_violin(mse_df_cont, "mse", "")
plot_average(mse_df_cont, "mse", "")
```

## average bias of true CATE vs predicted CATE

```{r cont_bias}
plot_box(bias_df_cont, "bias", "")
plot_violin(bias_df_cont, "bias", "")
plot_average(bias_df_cont, "bias", "")
```

## coverage of true CATE vs predicted CATE

```{r contcov}
plot_box(coverage_df_cont, "coverage", "")
plot_violin(coverage_df_cont, "coverage", "")
plot_average(coverage_df_cont, "coverage", "")
```

## average CI length

```{r cont_ci}
plot_box(length_df_cont, "length", "")
plot_violin(length_df_cont, "length", "")
plot_average(length_df_cont, "length", "")
```

```{r BLP_cont_df, include=FALSE}
BLP_df_cont <- BLP_df_cont %>% mutate(
  TP = TP/1000,
  FP = FP/1000,
  TN = TN/1000,
  FN = FN/1000)

BLP_df_cont <- BLP_df_cont %>%
  mutate(agreement = ifelse(scenario == "scenario_1", TN, TP),
         disagreement = ifelse(scenario == "scenario_1", FP, FN)
  )
```

## BLP results

```{r - BLP_cont}
compact_cont <- BLP_df_cont %>%
  select(model, n, scenario, agreement) %>%
  pivot_wider( names_from = scenario, values_from = agreement ) %>%
  arrange(model, n)

knitr::kable(compact_cont)
```
