---
title: "Correlated binary CI study: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(purrr)
library(knitr)
```

## Do the bootstrap bands still cover when the covariates are correlated?

`confidence_intervals/binary/`'s design, with the half-sample bootstrap's
`CI_sf` sweep and the same estimators, run on `correlated/binary/`'s
scenarios 1–4. X1–X5 and X01–X03 are drawn from a Gaussian copula with
exchangeable latent correlation ρ. τ(x) and m0(x) are the parent studies'
functions, so only the covariates' joint distribution changes. Full design in
`../README.md`. The grid itself is defined once, in `bin_corr_ci_config.R`.

| | |
|---|---|
| scenarios | 1–4 |
| ρ | 0, 0.5 |
| n | 500, 1000 |
| `CI_sf` | 0.05 to 0.50 in steps of 0.05 |
| runs | 100 |
| array | **16,000 jobs**, split across two jobscripts |

| Scenario | true CATE $\tau(x)$, a risk difference |
|---|---|
| 1: No HTE | $b_W$ |
| 2: Continuous HTE | $b_W + b_4 \tanh(X_4)$ |
| 3: Two vars + cts x cts | $b_W + b_3 X_3 + b_4 \tanh(X_4) + b_{45} \tanh(X_4) \tanh(X_5)$ |
| 4: Cosine HTE | $b_W + b_4 \cos(X_4)$ |

The outcome is $P(Y = 1 \mid x, W) = m_0(x) + W \tau(x)$, with the control risk
$m_0(x) = 0.34 + 0.36\,\mathrm{plogis}(b_0 + b_1 X_1 + b_2 X_2)$, as in
`binary/`.

- **ρ = 0** is the paired independent arm: the copula with no correlation.
- **ρ = 0.5** is the correlated arm. The prognostic X1 and X2 move with the
  modifiers X3–X5 (cor(m0, τ) is 0.39 / −0.34 / 0.06 in scenarios 2 / 3 / 4),
  so the forests have to separate prognosis from effect modification. In
  scenario 3 the lowest treated risk falls to about 0.007 at small n
  (`../../README.md`, "Binary scenario 3 floor"). At this study's n ≥ 500
  |bW| is smaller, so the floor is higher.

**Pairing.** A run is seeded by its run index alone. Run *r* is the same
dataset at every `CI_sf`, and at the two values of ρ it shares W, the raw
latent normals and the categorical covariates. Y is coupled only through
`rbinom`'s shared uniforms, so most but not all outcomes agree across ρ
(`../../README.md`, "Pairing"). Every ρ = 0.5 − ρ = 0 comparison below is a per-run difference
with MCSE sd(diff)/√pairs. The level plots show each ρ's own (unpaired) MCSE,
so the paired differences are where to judge whether ρ matters.

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle` (`CI_MODELS` in `R/metrics.R`), on the parent's
`profile = "ci"` (`bin_corr_ci_models.R`).

### Performance measures

* **Marginal coverage**: the proportion of units their own interval covers.
* **Simultaneous coverage**: 0/1 per run, whether the band covered every unit
  at once. This is what the half-sample bootstrap is constructed to control.
* **CI length**: mean interval length within a run.
* **Bias-eliminated (BE) coverage** (query grid only): marginal and
  simultaneous coverage of the across-run mean estimate instead of the true
  τ. It separates the band's width from the point estimate's bias.

Nominal is 0.95 for both coverages.

### Three interval kinds, and the query grid at ρ = 0.5

As in the parent report, `bin_corr_ci_metrics.R` scores the per-unit bootstrap
band, grf's pointwise variance interval (`causal_forest` only) and the
bootstrap band over the fixed query grid (`build_query_grid()`).

**The query grid extrapolates at ρ = 0.5.** It crosses X4 and X5 over
[−2, 2] independently and holds every other covariate at 0. The true τ at
those points is still exact, but under correlation a corner like
(X4 = 2, X5 = −2) sits far out in the joint distribution, so the forests are
extrapolating there. A fall in grid coverage at ρ = 0.5 can be that, not a
bootstrap failure. **The per-unit band is the primary ρ comparison.**

```{r data}
# paths
path <- here()
res_root <- file.path(dirname(path), "results")
res_path <- file.path(res_root, "correlated", "confidence_intervals", "binary")

# the parent CI report's labels, so the two reports read side by side
scenario_labels <- c(
  `1` = "1: No HTE",
  `2` = "2: Continuous HTE",
  `3` = "3: Two vars + cts x cts",
  `4` = "4: Cosine HTE"
)

model_labels <- c(
  causal_forest = "Causal forest",
  dr_random_forest = "DR-RandomForest",
  dr_oracle = "DR-oracle",
  dr_semi_oracle = "DR-semi-oracle"
)

interval_labels <- c(
  half_boot      = "Half-sample bootstrap (per unit)",
  grf_normal     = "Causal forest variance (pointwise)",
  half_boot_grid = "Half-sample bootstrap (query grid)"
)

#' Split `model` into estimator x interval kind and apply the display labels
#'
#' The metrics scripts write the alternative intervals as extra rows of
#' `model` ("causal_forest_inbuilt", "<model>_grid"), as the parent study's do.
#' rho stays numeric (the paired differences pivot on it); rho_lab is for
#' facets. Also used on the parent study's metrics (no rho) and on the
#' optimal_sf metrics (no CI_sf).
tidy_ci <- function(df) {
  df <- df %>%
    filter(scenario %in% c(1:4)) %>%
    mutate(
      interval = case_when(
        model == "causal_forest_inbuilt" ~ "grf_normal",
        grepl("_grid$", model)           ~ "half_boot_grid",
        TRUE                             ~ "half_boot"
      ),
      model = sub("_inbuilt$|_grid$", "", model)
    ) %>%
    mutate(
      scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                        labels = unname(scenario_labels)),
      n = factor(n, levels = c(500, 1000)),
      model = factor(recode(model, !!!model_labels), levels = unname(model_labels)),
      interval = factor(recode(interval, !!!interval_labels),
                        levels = unname(interval_labels))
    )
  # the same seq() the study config uses, so the doubles match exactly
  if ("CI_sf" %in% names(df)) df$CI_sf <- factor(df$CI_sf, levels = seq(0.05, 0.5, 0.05))
  if ("rho" %in% names(df)) df$rho_lab <- factor(paste("ρ =", df$rho))
  droplevels(df)
}

#' Mean and MCSE of the three interval metrics per cell
#'
#' MCSE denominators are the non-NA counts, as the parent report: one failed
#' CI bound makes a run's marginal coverage and length NA (interval_metrics()
#' has no na.rm). Simultaneous coverage is a per-run 0/1 indicator, so it gets
#' the binomial MCSE.
summarise_ci <- function(df, by) {
  df %>%
    group_by(across(all_of(by))) %>%
    summarise(
      mean_marg_cov = mean(marginal_coverage, na.rm = T),
      mcse_marg_cov = sd(marginal_coverage, na.rm = T) / sqrt(sum(!is.na(marginal_coverage))),
      mean_simul_cov = mean(simultaneous_coverage, na.rm = T),
      mcse_simul_cov = sqrt(mean_simul_cov * (1 - mean_simul_cov) /
                              sum(!is.na(simultaneous_coverage))),
      mean_ci_len = mean(mean_ci_length, na.rm = T),
      mcse_ci_len = sd(mean_ci_length, na.rm = T) / sqrt(sum(!is.na(mean_ci_length))),
      runs = n(),
      .groups = "drop"
    )
}

metrics <- readRDS(file.path(res_path, "bin_corr_ci_metrics.RDS")) %>% tidy_ci()

CELL <- c("rho", "rho_lab", "scenario", "n", "CI_sf", "model", "interval")
metrics_summary <- summarise_ci(metrics, CELL)

# everything outside the interval-kind sections is about the per-unit bootstrap
metrics_hb <- filter(metrics, interval == interval_labels[["half_boot"]])
summary_hb <- filter(metrics_summary, interval == interval_labels[["half_boot"]])
```

```{r plot-helpers, include=FALSE}
ci_theme <- function() {
  list(scale_colour_paletteer_d("rcartocolor::Safe"),
       theme_bw(),
       theme(axis.text.x = element_text(angle = 45, hjust = 1),
             strip.background = element_rect(fill = "white"),
             strip.text = element_text(colour = "black"),
             legend.position = "bottom",
             legend.box = "vertical"))
}

with_ci <- function(df, mean_col, mcse_col, alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  mutate(df, est = .data[[mean_col]],
         lo = .data[[mean_col]] - z * .data[[mcse_col]],
         hi = .data[[mean_col]] + z * .data[[mcse_col]])
}

# the parent report's summary_plot, with rho added to the facet columns: each
# scenario keeps its own y scale and its four (n, rho) panels share it
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  p <- ggplot(with_ci(df, mean_col, mcse_col),
              aes(x = CI_sf, y = est, colour = model, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(n, rho_lab), scales = "free_y",
               labeller = labeller(n = function(x) paste("n =", x))) +
    ci_theme() +
    labs(title = title, y = ylab, x = "Subsampling ratio", colour = "Model")
}

# interval kind takes the colour and rho the shape, one plot per n
interval_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  p <- ggplot(with_ci(df, mean_col, mcse_col),
              aes(x = CI_sf, y = est, colour = interval, shape = rho_lab,
                  ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.6), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.6), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
    ci_theme() +
    labs(title = title, y = ylab, x = "Subsampling ratio",
         colour = "Interval kind", shape = "ρ")
}

#' Paired rho = 0.5 - rho = 0 difference in a per-run column, by cell
#'
#' Runs are matched on (by, run). The MCSE is sd(diff) / sqrt(pairs), which is
#' what the shared seeds buy over the unpaired sqrt(mcse_0^2 + mcse_0.5^2). A
#' run with the column NA at either rho drops out of its cell.
paired_diff <- function(df, col, by = c("scenario", "n", "CI_sf", "model", "interval")) {
  df %>%
    select(all_of(c(by, "run", "rho")), value = all_of(col)) %>%
    pivot_wider(names_from = rho, values_from = value, names_prefix = "rho_") %>%
    mutate(diff = rho_0.5 - rho_0) %>%
    group_by(across(all_of(by))) %>%
    summarise(mean_diff = mean(diff, na.rm = TRUE),
              mcse_diff = sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))),
              pairs = sum(!is.na(diff)), .groups = "drop")
}

diff_plot <- function(df, title, ylab, x = "CI_sf", colour = "model",
                      facet = facet_grid(rows = vars(scenario), cols = vars(n),
                                         scales = "free_y",
                                         labeller = labeller(n = function(x) paste("n =", x))),
                      xlab = "Subsampling ratio", colour_lab = "Model") {
  ggplot(with_ci(df, "mean_diff", "mcse_diff"),
         aes(x = .data[[x]], y = est, colour = .data[[colour]], ymin = lo, ymax = hi)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet +
    ci_theme() +
    labs(title = title, y = ylab, x = xlab, colour = colour_lab)
}
```

## Marginal coverage

The proportion of units their own interval covers, against the nominal 95%
(dashed line), across the sweep.

```{r marg-cov-plot, fig.height=9}
summary_plot(summary_hb, "mean_marg_cov", "mcse_marg_cov",
             "Mean marginal coverage", "Marginal coverage", hline = 0.95)
```

## Simultaneous coverage

The proportion of runs whose band covered every unit, the guarantee the
method claims.

```{r simul-cov-plot, fig.height=9}
summary_plot(summary_hb, "mean_simul_cov", "mcse_simul_cov",
             "Mean simultaneous coverage", "Simultaneous coverage", hline = 0.95)
```

## CI length

```{r ci-len-plot, fig.height=9}
summary_plot(summary_hb, "mean_ci_len", "mcse_ci_len",
             "Mean CI length", "Mean CI length")
```

## Effect of the correlation: paired ρ = 0.5 − ρ = 0 differences

Each point is the mean over runs of a run's metric at ρ = 0.5 minus the same
run's at ρ = 0, with a 95% CI from the paired MCSE. A CI clear of the dashed
line is a change the correlation made, not Monte Carlo noise. Per-unit
bootstrap band.

```{r diff-plots, fig.height=8}
diff_plot(paired_diff(metrics_hb, "marginal_coverage"), "Marginal coverage",
          "Δ marginal coverage (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics_hb, "simultaneous_coverage"), "Simultaneous coverage",
          "Δ simultaneous coverage (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics_hb, "mean_ci_length"), "CI length",
          "Δ mean CI length (ρ = 0.5 − ρ = 0)")
```

## The three interval kinds

As in the parent report: the pointwise interval is not competing on
simultaneous coverage, and does not depend on `CI_sf`. The query-grid band is
a different target, and at ρ = 0.5 a partly extrapolated one (see above). One
plot per sample size.

```{r interval-plots, fig.height=9}
for (nn in levels(metrics_summary$n)) {
  d <- filter(metrics_summary, n == nn)
  print(interval_plot(d, "mean_marg_cov", "mcse_marg_cov",
                      paste0("Mean marginal coverage by interval kind, n = ", nn),
                      "Marginal coverage", hline = 0.95))
  print(interval_plot(d, "mean_simul_cov", "mcse_simul_cov",
                      paste0("Mean simultaneous coverage by interval kind, n = ", nn),
                      "Simultaneous coverage", hline = 0.95))
  print(interval_plot(d, "mean_ci_len", "mcse_ci_len",
                      paste0("Mean CI length by interval kind, n = ", nn),
                      "Mean CI length"))
}
```

The paired ρ effect by interval kind. If the grid band loses more coverage
than the per-unit band, the extrapolation is the likelier cause.

```{r interval-diff-plots, fig.height=9}
interval_facet <- facet_grid(rows = vars(scenario), cols = vars(model, n), scales = "free_y",
                             labeller = labeller(n = function(x) paste("n =", x)))
diff_plot(paired_diff(metrics, "marginal_coverage"), "Marginal coverage, by interval kind",
          "Δ marginal coverage (ρ = 0.5 − ρ = 0)", colour = "interval",
          facet = interval_facet, colour_lab = "Interval kind")
diff_plot(paired_diff(metrics, "simultaneous_coverage"),
          "Simultaneous coverage, by interval kind",
          "Δ simultaneous coverage (ρ = 0.5 − ρ = 0)", colour = "interval",
          facet = interval_facet, colour_lab = "Interval kind")
```

### Bias-eliminated coverage on the query grid

Bias-eliminated (BE) coverage scores the grid band against the across-run
mean estimate at each grid point instead of the true τ. Its gap to the
ordinary grid coverage is the part of any undercoverage that comes from the
point estimate's bias, which is where extrapolation shows up. It is defined on
the grid only: the per-unit bands have no fixed point to average over runs.
The MCSEs treat the across-run mean as fixed, so they leave out that
reference's own Monte Carlo error.

```{r be-data}
grid_runs <- filter(metrics, interval == interval_labels[["half_boot_grid"]])

#' Mean and MCSE of true-tau and BE grid coverage per cell, long by reference
#'
#' The same MCSEs as summarise_ci(): sd / sqrt(non-NA) for the marginal
#' coverages, binomial for the 0/1 simultaneous ones.
summarise_be <- function(df, by) {
  binom_mcse <- function(x) {
    p <- mean(x, na.rm = TRUE)
    sqrt(p * (1 - p) / sum(!is.na(x)))
  }
  df %>%
    group_by(across(all_of(by))) %>%
    summarise(
      across(c(marginal_coverage, be_marginal_coverage),
             list(mean = ~ mean(.x, na.rm = TRUE),
                  mcse = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))))),
      across(c(simultaneous_coverage, be_simultaneous_coverage),
             list(mean = ~ mean(.x, na.rm = TRUE), mcse = binom_mcse)),
      .groups = "drop") %>%
    pivot_longer(matches("_coverage_(mean|mcse)$"),
                 names_to = c("reference", "measure", ".value"),
                 names_pattern = "^(be_)?(marginal|simultaneous)_coverage_(mean|mcse)$") %>%
    mutate(reference = factor(if_else(coalesce(reference, "") == "", "True τ",
                                      "Across-run mean (BE)"),
                              levels = c("True τ", "Across-run mean (BE)")))
}

be_summary <- summarise_be(grid_runs, CELL)

# reference takes the colour and rho the shape, as interval_plot()
be_plot <- function(df, title, ylab) {
  ggplot(with_ci(df, "mean", "mcse"),
         aes(x = CI_sf, y = est, colour = reference, shape = rho_lab,
             ymin = lo, ymax = hi)) +
    geom_hline(yintercept = 0.95, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.6), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.6), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
    ci_theme() +
    labs(title = title, y = ylab, x = "Subsampling ratio",
         colour = "Coverage of", shape = "ρ")
}
```

Coverage of the true τ and of the across-run mean, side by side with 95%
MCSE intervals. One plot per sample size and measure.

```{r be-plots, fig.height=9}
for (nn in levels(be_summary$n)) {
  for (m in c("simultaneous", "marginal")) {
    print(be_plot(filter(be_summary, n == nn, measure == m),
                  paste0("Query-grid band: ", m, " coverage, true τ vs BE, n = ", nn),
                  paste(tools::toTitleCase(m), "coverage")))
  }
}
```

The per-run gap, BE minus true-τ coverage. Both come from the same band in
the same run, so the MCSE is sd(gap)/√runs. A gap clear of zero is coverage
lost to the point estimate's bias rather than to the band's width.

```{r be-gap-plots, fig.height=9}
#' Per-run BE - true-tau coverage, mean and paired MCSE per cell
be_gap <- function(df, be_col, true_col) {
  df %>%
    mutate(diff = .data[[be_col]] - .data[[true_col]]) %>%
    group_by(rho_lab, scenario, n, CI_sf, model) %>%
    summarise(mean_diff = mean(diff, na.rm = TRUE),
              mcse_diff = sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))),
              .groups = "drop")
}

gap_facet <- facet_grid(rows = vars(scenario), cols = vars(n, rho_lab), scales = "free_y",
                        labeller = labeller(n = function(x) paste("n =", x)))
diff_plot(be_gap(grid_runs, "be_simultaneous_coverage", "simultaneous_coverage"),
          "Query-grid band: BE minus true-τ simultaneous coverage",
          "Δ simultaneous coverage (BE − true τ)", facet = gap_facet)
diff_plot(be_gap(grid_runs, "be_marginal_coverage", "marginal_coverage"),
          "Query-grid band: BE minus true-τ marginal coverage",
          "Δ marginal coverage (BE − true τ)", facet = gap_facet)
```

The paired ρ effect on BE coverage. With the bias taken out, a change here is
the band's width responding to the correlation, not the extrapolation.

```{r be-diff-plots, fig.height=8}
diff_plot(paired_diff(grid_runs, "be_simultaneous_coverage"),
          "Query-grid band: BE simultaneous coverage",
          "Δ BE simultaneous coverage (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(grid_runs, "be_marginal_coverage"),
          "Query-grid band: BE marginal coverage",
          "Δ BE marginal coverage (ρ = 0.5 − ρ = 0)")
```

```{r be-table}
# pooled over CI_sf. Run r is the same dataset at every CI_sf, so the rows are
# averaged within run first and the MCSE is over the run means.
be_cols <- c("simultaneous_coverage", "be_simultaneous_coverage",
             "marginal_coverage", "be_marginal_coverage")
grid_be <- grid_runs %>%
  group_by(scenario, n, rho, model, run) %>%
  summarise(across(all_of(be_cols), ~ mean(.x, na.rm = TRUE)), .groups = "drop_last") %>%
  summarise(across(all_of(be_cols),
                   list(mean = ~ mean(.x, na.rm = TRUE),
                        mcse = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))))),
            .groups = "drop") %>%
  arrange(scenario, n, model, rho)

kable(grid_be, digits = 3,
      caption = "Query-grid band, pooled over CI_sf: coverage vs bias-eliminated coverage",
      col.names = c("Scenario", "n", "ρ", "Model",
                    "Simultaneous coverage", "MCSE",
                    "BE simultaneous coverage", "MCSE",
                    "Marginal coverage", "MCSE",
                    "BE marginal coverage", "MCSE"))
```

## Choosing `CI_sf`

The distance from nominal simultaneous coverage, `|mean simultaneous coverage -
0.95|`, so **lower is better**. The table picks the arg-min per ρ, scenario,
sample size and estimator. A best ratio that moves with ρ means the right
`CI_sf` depends on the covariate structure, not just n.

```{r sf-distance-plot, fig.height=9}
sf_dist <- summary_hb %>%
  mutate(dist = abs(mean_simul_cov - 0.95))

sf_dist %>%
  ggplot(aes(x = CI_sf, y = dist, colour = model, linetype = rho_lab,
             group = interaction(model, rho_lab))) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 1.5) +
  geom_line(linewidth = 0.3) +
  facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y",
             labeller = labeller(n = function(x) paste("n =", x))) +
  ci_theme() +
  labs(title = "Distance from nominal simultaneous coverage",
       subtitle = "Lower is better; zero is exactly nominal",
       y = "|mean simultaneous coverage - 0.95|", x = "Subsampling ratio",
       colour = "Model", linetype = "ρ")
```

```{r sf-best-table}
best_sf <- sf_dist %>%
  group_by(rho, scenario, n, model) %>%
  slice_min(dist, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(scenario, n, model, rho, CI_sf, mean_simul_cov, mean_marg_cov, mean_ci_len) %>%
  arrange(scenario, n, model, rho)

kable(best_sf, digits = 4,
      col.names = c("Scenario", "n", "Model", "ρ", "Best CI_sf",
                    "Mean simultaneous coverage", "Mean marginal coverage",
                    "Mean CI length"))
```

## Data-driven `CI_sf` (`optimal_sf`)

```{r sf-data}
sf_file <- file.path(res_path, "sf_calibration", "bin_corr_ci_sf_metrics.RDS")
has_sf <- file.exists(sf_file)
if (has_sf) {
  sf <- readRDS(sf_file) %>% tidy_ci()
  sf_runs <- filter(sf, interval == interval_labels[["half_boot"]])
}
```

`r if (!has_sf) "**Skipped: the optimal_sf metrics (\x60sf_calibration/bin_corr_ci_sf_metrics.RDS\x60) are not on disk.**"`

`../optimal_sf/` picks `CI_sf` per run for `dr_random_forest`. It resamples
the pseudo-outcome residuals around the run's own estimate and takes the
ratio whose band covers that estimate closest to 90%. It then builds the final
band at that ratio. Two things to keep in mind:

- **α = 0.10 here, against the sweep's 0.05**, so the final bands are 90%
  bands and are judged against 0.90 (dashed lines). Compare the *picks* with
  the sweep's best ratio, not the coverage levels.
- **The calibration cannot see bias.** Its target is the estimate itself. At
  ρ = 0.5 the prognostic function tracks τ, which is where bias would grow, so
  the gap between the calibration's own coverage (`plugin coverage`) and the
  true-τ coverage is the thing to watch.

The distribution of the pick:

```{r sf-pick-plot, eval=has_sf, fig.height=8}
sf_runs %>%
  mutate(pick = factor(optimal_sf, levels = seq(0.05, 0.5, 0.05))) %>%
  ggplot(aes(x = pick, fill = rho_lab)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_x_discrete(drop = FALSE) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  facet_grid(rows = vars(scenario), cols = vars(n),
             labeller = labeller(n = function(x) paste("n =", x))) +
  ci_theme() +
  labs(title = "Calibrated CI_sf per run", x = "Chosen subsampling ratio",
       y = "Runs", fill = "ρ")
```

```{r sf-pick-diff, eval=has_sf, fig.height=4}
diff_plot(paired_diff(sf_runs, "optimal_sf", by = c("scenario", "n")),
          "Paired change in the calibrated CI_sf", "Δ CI_sf (ρ = 0.5 − ρ = 0)",
          x = "n", colour = "scenario", facet = facet_null(), xlab = "Sample size",
          colour_lab = "Scenario")
```

The final band's true-τ coverage, per unit and on the query grid:

```{r sf-coverage-plot, eval=has_sf, fig.height=8}
sf_summary <- summarise_ci(sf, c("rho", "rho_lab", "scenario", "n", "interval")) %>%
  left_join(sf %>%
              group_by(rho, scenario, n, interval) %>%
              summarise(mean_pick = mean(optimal_sf),
                        mean_plugin_cov = mean(plugin_coverage, na.rm = TRUE),
                        mean_be_simul_cov = mean(be_simultaneous_coverage, na.rm = TRUE),
                        mcse_be_simul_cov = sqrt(mean_be_simul_cov * (1 - mean_be_simul_cov) /
                                                   sum(!is.na(be_simultaneous_coverage))),
                        .groups = "drop"),
            by = c("rho", "scenario", "n", "interval"))

sf_summary %>%
  with_ci("mean_simul_cov", "mcse_simul_cov") %>%
  ggplot(aes(x = n, y = est, colour = rho_lab, ymin = lo, ymax = hi)) +
  geom_hline(yintercept = 0.90, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.4), size = 2) +
  geom_errorbar(position = position_dodge(width = 0.4), linewidth = 0.3, width = 0.2) +
  facet_grid(rows = vars(scenario), cols = vars(interval)) +
  ci_theme() +
  labs(title = "Simultaneous coverage of the true CATE at the calibrated CI_sf",
       y = "Simultaneous coverage", x = "Sample size", colour = "ρ")
```

```{r sf-coverage-diff, eval=has_sf, fig.height=6}
sf_diff_facet <- facet_grid(rows = vars(scenario), cols = vars(interval), scales = "free_y")
diff_plot(paired_diff(sf, "simultaneous_coverage", by = c("scenario", "n", "interval")),
          "Paired change in simultaneous coverage at the calibrated CI_sf",
          "Δ simultaneous coverage (ρ = 0.5 − ρ = 0)", x = "n", colour = "interval",
          facet = sf_diff_facet, xlab = "Sample size", colour_lab = "Interval kind")
diff_plot(paired_diff(sf, "marginal_coverage", by = c("scenario", "n", "interval")),
          "Paired change in marginal coverage at the calibrated CI_sf",
          "Δ marginal coverage (ρ = 0.5 − ρ = 0)", x = "n", colour = "interval",
          facet = sf_diff_facet, xlab = "Sample size", colour_lab = "Interval kind")
```

```{r sf-coverage-table, eval=has_sf}
kable(sf_summary %>%
        select(scenario, n, rho, interval, mean_pick, mean_plugin_cov, mean_marg_cov,
               mean_simul_cov, mcse_simul_cov, mean_be_simul_cov, mcse_be_simul_cov,
               mean_ci_len) %>%
        arrange(scenario, n, interval, rho),
      digits = 3,
      col.names = c("Scenario", "n", "ρ", "Interval", "Mean CI_sf picked",
                    "Plug-in coverage (calibration)", "Marginal coverage",
                    "Simultaneous coverage", "MCSE (simul.)",
                    "BE simultaneous coverage", "MCSE (BE simul.)", "Mean CI length"))
```

The pick against the sweep's best ratio for `dr_random_forest`. The sweep's
best is for a 95% band, the pick for a 90% one, so agreement is not expected
to be exact. What matters is whether the two move the same way with ρ.

```{r sf-vs-sweep, eval=has_sf}
pick_summary <- sf_runs %>%
  group_by(rho, scenario, n) %>%
  summarise(mean_pick = mean(optimal_sf),
            median_pick = median(optimal_sf),
            modal_pick = as.numeric(names(which.max(table(optimal_sf)))),
            .groups = "drop")

sweep_best <- best_sf %>%
  filter(model == model_labels[["dr_random_forest"]]) %>%
  transmute(rho, scenario, n, best_sweep_sf = as.numeric(as.character(CI_sf)),
            sweep_simul_cov = mean_simul_cov)

kable(left_join(pick_summary, sweep_best, by = c("rho", "scenario", "n")) %>%
        select(scenario, n, rho, everything()) %>%
        arrange(scenario, n, rho),
      digits = 3,
      col.names = c("Scenario", "n", "ρ", "Mean pick", "Median pick", "Modal pick",
                    "Sweep's best CI_sf", "Sweep simul. coverage there"))
```

## Sanity check: ρ = 0 against `confidence_intervals/binary/`

```{r parent-data}
parent_file <- file.path(res_root, "confidence_intervals", "binary", "ci_bin_metrics.RDS")
has_parent <- file.exists(parent_file)
```

`r if (!has_parent) "**Skipped: \x60confidence_intervals/binary/\x60's metrics are not on disk.**"`

The ρ = 0 arm has the same distribution as the parent's scenarios 1–4, but
different draws: the copula draws in another order, so the two are **not**
paired and the z-scores use the unpaired MCSE. With this many cells a few
`|z| > 3` are expected by chance; many more point to a DGM or pipeline
difference.

```{r sanity, eval=has_parent}
keys <- c("scenario", "n", "CI_sf", "model", "interval")
rho0 <- summarise_ci(filter(metrics, rho == 0), keys)
par <- summarise_ci(readRDS(parent_file) %>% tidy_ci(), keys)
cmp <- inner_join(rho0, par, by = keys, suffix = c("_rho0", "_parent"))
stems <- c(marg_cov = "marg_cov", simul_cov = "simul_cov", ci_len = "ci_len")
z <- sapply(stems, function(s) {
  (cmp[[paste0("mean_", s, "_rho0")]] - cmp[[paste0("mean_", s, "_parent")]]) /
    sqrt(cmp[[paste0("mcse_", s, "_rho0")]]^2 + cmp[[paste0("mcse_", s, "_parent")]]^2)
})
z_df <- bind_cols(select(cmp, all_of(keys)),
                  as.data.frame(round(z, 2)) %>% rename_with(~ paste0("z_", .x)))
# a simultaneous coverage of exactly 0 or 1 in both gives an MCSE of 0, so its
# z is NaN - not a flag
n_z <- sum(is.finite(z))
n_big <- sum(abs(z) > 3, na.rm = TRUE)
cat(sprintf("%d of %d finite z-scores have |z| > 3 (%.1f expected by chance).\n",
            n_big, n_z, n_z * 2 * pnorm(-3)))
kable(filter(z_df, if_any(starts_with("z_"), ~ abs(.x) > 3)) %>%
        arrange(scenario, n, model, CI_sf),
      caption = "Cells with |z| > 3, ρ = 0 vs confidence_intervals/binary/")
```

## Headline table

Every cell of the sweep for the per-unit bootstrap.

```{r headline-table}
headline <- summary_hb %>%
  select(scenario, n, model, CI_sf, rho, mean_marg_cov, mcse_marg_cov,
         mean_simul_cov, mcse_simul_cov, mean_ci_len, mcse_ci_len) %>%
  arrange(scenario, n, model, CI_sf, rho)

kable(headline, digits = 4,
      col.names = c("Scenario", "n", "Model", "CI_sf", "ρ", "Mean marginal coverage",
                    "MCSE (marg. cov.)", "Mean simultaneous coverage",
                    "MCSE (simul. cov.)", "Mean CI length", "MCSE (CI length)"))
```
