---
title: "ICTMC poster - candidate plots"
author: "Ellie Van Vogt"
date: today
format: html
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 7)
library(here)
source(here("results_processing", "poster_figs", "R", "candidates.R"))

#' Emit one Quarto tab per candidate plot, inside an already-open
#' `::: {.panel-tabset}` block
emit_plot_tabs <- function(plots) {
  for (nm in names(plots)) {
    cat("\n\n#### ", nm, "\n\n")
    print(plots[[nm]])
    cat("\n\n")
  }
}
```

Full-size candidates for the two panels piloted so far - see
`../ICTMC_poster_planning.md` for what each one is trying to settle, and
`explore_gallery.R` for the faster-to-scan contact-sheet version of the same
plots. Candidate-building logic lives in `R/candidates.R`; this document only
arranges the plots it returns into tabs.

## Sample size

Outer tabs are the outcome-scope choice (show both outcomes, or pick one);
inner tabs are the metric choice.

```{r ss-tabs}
#| results: asis
outcomes <- c(both = "Both outcomes", continuous = "Continuous only",
             binary = "Binary only")
cat("::: {.panel-tabset}\n\n")
for (oc in names(outcomes)) {
  cat("\n\n### ", outcomes[[oc]], "\n\n")
  cat("::: {.panel-tabset}\n\n")
  emit_plot_tabs(candidate_plots_sample_size(outcome = oc))
  cat("\n\n:::\n\n")
}
cat("\n\n:::\n\n")
```

## Missing data

Outer tabs are the two summarisation choices raised in the planning doc
(average over CATE model, or keep it split); inner tabs are the metric
choice.

```{r miss-tabs}
#| results: asis
variants <- c(FALSE, TRUE)
names(variants) <- c("Averaged over model", "Split by model")
cat("::: {.panel-tabset}\n\n")
for (vn in names(variants)) {
  cat("\n\n### ", vn, "\n\n")
  cat("::: {.panel-tabset}\n\n")
  emit_plot_tabs(candidate_plots_missing(by_model = variants[[vn]]))
  cat("\n\n:::\n\n")
}
cat("\n\n:::\n\n")
```
