---
title: "Crossfitting CI pilot: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Does the pilot's bootstrap deliver nominal coverage?

`crossfitting/` benchmarks its double-crossfitting procedure against cheaper
alternatives (see `cf_results.qmd`), but says nothing about whether any of
them, once fitted, produce trustworthy confidence intervals. This pilot adds
half-sample bootstrap intervals (`R/bootstrap_ci.R` — the same machinery
`confidence_intervals/` uses elsewhere) to **all 12 non-SuperLearner arms**.
The 3 SuperLearner arms stay out of scope: not RF-based.

The 12 split into two structural families, and the bootstrap differs between
them. The 5 arms whose stage 2 is a genuine per-fold crossfit use
`rf_half_boot` / `cf_half_boot`, which refit per fold against a fold-stratified
half sample. The 7 whole-sample / OOB arms have no fold structure to refit
against, so `rf_oob_half_boot` / `cf_oob_half_boot` refit one forest per draw
on an unstratified half sample instead.

Coverage from this method is already known, from elsewhere in
`confidence_intervals/`, to run below nominal. That's the actual research
question here: not *whether* it undershoots, but *how much*, and whether any
arm does noticeably better than the others for its cost.

### Design

Fixed at **n = 500, V = 10 folds, scenarios Null / Interaction / Non-linear,
50 runs** (150 array jobs) — a pilot, not the full production grid.
`CI_boot = 200` half-sample bootstrap draws, `CI_sf` (grf's `sample.fraction`)
fixed at 0.5 rather than swept.

Every arm is scored on the **training** sample only. Unlike the main
crossfitting comparison, no test-sample score exists here: no interval is
computed for `tau_test`, only for the training-sample `tau` the arm was
actually fit on.

Point estimates come from `run_all_crossfit_variants(sl_lib = NULL)` — the same
call `cf_analysis.R` makes, minus the SuperLearner family, which sits strictly
after the RF/CF arms in the RNG stream. Under the same seed these arms are
therefore **bit-identical** to the production study's, so any point-estimate
figure below should reproduce `cf_results.qmd`'s exactly.

### Arms compared

| id | family | stage 1 (nuisances) | stage 2 / bootstrap structure |
|---|---|---|---|
| `dcf` | DR-learner (RF) | double CF over fold pairs | crossfit, same folds, column `k` (**status quo**) |
| `scf_scf` | DR-learner (RF) | single CF, leave-one-fold-out | crossfit, same folds |
| `scf_scf_new` | DR-learner (RF) | single CF | crossfit, **fresh independent** split |
| `scf_oob` | DR-learner (RF) | single CF | whole sample, **OOB** |
| `scf_oob_t` | DR-learner (RF) | single CF, T-learner | whole sample, **OOB** |
| `oob_oob` | DR-learner (RF) | whole sample, **OOB**, T-learner | whole sample, **OOB** |
| `oob_oob_s` | DR-learner (RF) | whole sample, **OOB**, S-learner (`X.orig`) | whole sample, **OOB** |
| `oob_oob_manual` | DR-learner (RF) | whole sample, **OOB**, S-learner (tree loop) | whole sample, **OOB** |
| `cf_dcf` | Causal forest | double-CF matrices, column `k` | fold-wise (**status quo**) |
| `cf_scf` | Causal forest | single-CF vectors | fold-wise |
| `cf_full_oob` | Causal forest | single-CF vectors | whole sample, **OOB** |
| `cf_default` | Causal forest | grf's own internal OOB | whole sample, **OOB** |

Only the final-stage forest is refit per half-sample — nuisances are held fixed
and sliced, including `cf_default`'s, whose `Y.hat`/`W.hat` grf cross-fit for
itself. The bootstrap draws matrices are dropped before saving; only the
resulting bounds are kept.

### Three interval methods

The OOB arms carry three intervals, not one, and they are nearly free to have:

| `ci_method` | arms | kind | how |
|---|---|---|---|
| `half_boot` | all 12 | simultaneous | every unit scored in every draw |
| `half_boot_out` | the 7 OOB arms | simultaneous | same refits, only out-of-half units scored |
| `grf_normal` | the 7 OOB arms | **pointwise** | grf's own OOB variance, no bootstrap at all |

`half_boot` and `half_boot_out` come off one set of forest fits — a paired
contrast, differing only in whether a unit that was *in* the half sample gets
scored by the half forest's own OOB prediction. `half_boot_out` takes its
supremum over ~`n/2` rather than `n` units, so it is expected to run narrower.

`grf_normal` is the free comparator: for a whole-sample forest grf returns OOB
variance estimates (bootstrap of little bags) alongside the predictions, and
`R/metrics.R`'s `normal_interval` turns those into a CI at no extra compute.
**It is a pointwise interval** and is expected to score near zero on
simultaneous coverage — that is the method working as designed, not a defect.
It is natural for the OOB arms and awkward for the crossfit ones (whose `tau` is
stitched from `V` fold models), which is exactly the mirror image of the
bootstrap, and why the OOB arms are the first place both apply to the same arm.

Neither method reflects first-stage nuisance uncertainty: the bootstrap holds
nuisances fixed, and grf's variance treats the pseudo-outcomes as known. Whether
that is second-order is the Neyman-orthogonality claim this study probes.

Full design rationale is in `crossfitting/README.md`'s "Half-sample bootstrap CI
pilot" section.

```{r data}
# functions and label dictionaries - ci_variant_levels/ci_variant_labels here,
# family_labels transitively via cf_metrics.R
source(here("crossfitting", "confidence_intervals", "cf_ci_metrics.R"))

# paths - a separate results tree from crossfitting/'s
path <- here()
res_path <- file.path(dirname(path), "results", "crossfitting_ci")

metrics <- readRDS(file.path(res_path, "cf_ci_metrics.RDS"))

# tidy up - no `set` column: training-sample tau only
metrics_all <- metrics %>%
  mutate(
    scenario = factor(
      case_when(scenario == 1 ~ "Null",
                scenario == 6 ~ "Interaction",
                scenario == 9 ~ "Non-linear"),
      levels = c("Null", "Interaction", "Non-linear")),
    family = factor(recode(family, !!!family_labels), levels = unname(family_labels)),
    variant = factor(recode(variant, !!!ci_variant_labels),
                     levels = unname(ci_variant_labels[ci_variant_levels])),
    ci_method = factor(recode(ci_method, !!!ci_method_labels),
                       levels = unname(ci_method_labels[ci_method_levels]))
  ) %>%
  droplevels()

# every section below the method comparison is about the half-sample bootstrap,
# the one interval available for all 12 arms
metrics <- filter(metrics_all, ci_method == ci_method_labels[["half_boot"]])

# per variant summaries
metrics_summary <- metrics %>%
  group_by(scenario, family, variant) %>%
  summarise(
    mean_marg_cov = mean(marginal_coverage, na.rm = T),
    mcse_marg_cov = sd(marginal_coverage, na.rm = T) / sqrt(n()),
    mean_simul_cov = mean(simultaneous_coverage, na.rm = T),
    mcse_simul_cov = sd(simultaneous_coverage, na.rm = T) / sqrt(n()),
    mean_ci_len = mean(mean_ci_length, na.rm = T),
    mcse_ci_len = sd(mean_ci_length, na.rm = T) / sqrt(n()),
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(n()),
    mean_time_nuisance = mean(time_nuisance, na.rm = T),
    mean_time_stage2 = mean(time_stage2, na.rm = T),
    mean_time_boot = mean(time_boot, na.rm = T),
    mean_time_total = mean(time_nuisance + time_stage2 + time_boot, na.rm = T),
    .groups = "drop"
  )
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with MCSE error bars. Unlike
# cf_results.qmd's version there's no `set` (train/test) dimension to dodge
# on - one sample per arm/run - so colour by variant instead (redundant with
# x under faceting, so the legend is dropped). hline is optional: not every
# metric here has a natural 0/0.95 reference.
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = variant, y = est, colour = variant, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(size = 2) +
    geom_errorbar(linewidth = 0.3, width = 0.3) +
    facet_wrap(vars(family, scenario), nrow = n_distinct(df$family), scales = "free") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
    labs(title = title, y = ylab, x = "Crossfitting procedure")
}
```

## Marginal coverage

The headline metric: the proportion of units the interval covers, for every
arm, against the nominal 95% target (dashed line).

```{r marg-cov-plot, fig.height=16}
marg_cov_plot <- metrics %>%
  ggplot(aes(x = variant, y = marginal_coverage, colour = variant)) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "Marginal coverage of the half-sample bootstrap interval",
       y = "Marginal coverage", x = "Crossfitting procedure")

marg_cov_sum_plot <- summary_plot(metrics_summary, "mean_marg_cov", "mcse_marg_cov",
                                  "Mean marginal coverage", "Mean marginal coverage",
                                  hline = 0.95)

(marg_cov_plot / marg_cov_sum_plot) + plot_layout(guides = "collect")
```

## Simultaneous coverage

Whether the band covers *every* unit at once — the stronger guarantee the
half-sample bootstrap is constructed to control, and a per-run 0/1 indicator,
so only the mean-proportion summary is shown (a boxplot of a binary outcome
is uninformative).

```{r simul-cov-plot}
simul_cov_sum_plot <- summary_plot(metrics_summary, "mean_simul_cov", "mcse_simul_cov",
                                   "Mean simultaneous coverage", "Mean simultaneous coverage",
                                   hline = 0.95)

simul_cov_sum_plot
```

## CI width

Mean interval length per replicate. There is no natural reference line here
— width only matters relative to the coverage it buys.

```{r width-plot, fig.height=16}
width_plot <- metrics %>%
  ggplot(aes(x = variant, y = mean_ci_length, colour = variant)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "Width of the half-sample bootstrap interval",
       y = "Mean interval length", x = "Crossfitting procedure")

width_sum_plot <- summary_plot(metrics_summary, "mean_ci_len", "mcse_ci_len",
                               "Mean interval width", "Mean interval length")

(width_plot / width_sum_plot) + plot_layout(guides = "collect")
```

## Three interval methods on the OOB arms

The one place all three methods exist for the **same** arm, and the sharpest
question the pilot can answer: does an expensive bootstrap buy anything over the
variance grf hands back for free?

Read the two panels together, not separately. `grf_normal` is pointwise — it is
not built to control simultaneous coverage, so the marginal panel is the fair
reading of it. Both bootstrap bands are simultaneous and come off identical
refits, differing only in whether in-half units are scored; `half_boot_out`
takes its supremum over ~`n/2` rather than `n` units, so a narrower band is the
expected result, not a surprising one.

```{r method-comparison, fig.height=16}
oob_variants <- ci_variant_labels[c("scf_oob", "scf_oob_t", "oob_oob", "oob_oob_s",
                                    "oob_oob_manual", "cf_full_oob", "cf_default")]

method_df <- metrics_all %>%
  filter(variant %in% oob_variants) %>%
  droplevels() %>%
  group_by(scenario, family, variant, ci_method) %>%
  summarise(mean_marg_cov = mean(marginal_coverage, na.rm = T),
            mcse_marg_cov = sd(marginal_coverage, na.rm = T) / sqrt(n()),
            mean_ci_len = mean(mean_ci_length, na.rm = T),
            mcse_ci_len = sd(mean_ci_length, na.rm = T) / sqrt(n()),
            .groups = "drop")

method_cov_panel <- method_df %>%
  ggplot(aes(x = variant, y = mean_marg_cov, colour = ci_method,
             ymin = mean_marg_cov - mcse_marg_cov, ymax = mean_marg_cov + mcse_marg_cov)) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(linewidth = 0.3, width = 0.3, position = position_dodge(width = 0.5)) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(method_df$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Mean marginal coverage", x = NULL, colour = "Interval method")

method_width_panel <- method_df %>%
  ggplot(aes(x = variant, y = mean_ci_len, colour = ci_method,
             ymin = mean_ci_len - mcse_ci_len, ymax = mean_ci_len + mcse_ci_len)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  geom_errorbar(linewidth = 0.3, width = 0.3, position = position_dodge(width = 0.5)) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(method_df$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Mean interval length", x = "Whole-sample / OOB arm", colour = "Interval method")

(method_cov_panel / method_width_panel) + plot_layout(guides = "collect")
```

## Point-estimate accuracy

Not this report's focus, and per-arm rather than per-method. Under the same seed
these arms are bit-identical to the production study's, so this should reproduce
`cf_results.qmd`'s numbers exactly — the cheapest possible check that the pilot
is wired to the same estimators (bias and MSE of the CATE point estimate against
the known truth).

```{r point-plot, fig.height=10}
bias_plot <- metrics %>%
  ggplot(aes(x = variant, y = bias, colour = variant)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "Bias of CATE point estimates",
       y = "Average bias (estimate - truth)", x = "Crossfitting procedure")

mse_plot <- metrics %>%
  ggplot(aes(x = variant, y = mse, colour = variant)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "MSE of CATE point estimates",
       y = "MSE", x = "Crossfitting procedure")

bias_plot + mse_plot
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in the Null scenario since there's no heterogeneity to correlate with
— excluded from this plot).

```{r corr-plot, fig.height=10}
corr_df <- filter(metrics, scenario != "Null")

corr_plot <- corr_df %>%
  ggplot(aes(x = variant, y = corr, colour = variant)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  # free_x only: correlation is already on a bounded, comparable -1..1
  # scale, so y stays shared - only the sparse variant axis needs freeing
  facet_wrap(vars(family, scenario), nrow = n_distinct(corr_df$family), scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "Correlation between estimated and true CATEs",
       y = "Pearson correlation", x = "Crossfitting procedure")

corr_plot
```

## Runtime

What nuisances, the final model, and the bootstrap itself each cost.

```{r time-plot, fig.height=10}
time_df <- metrics_summary %>%
  select(scenario, family, variant, mean_time_nuisance, mean_time_stage2, mean_time_boot) %>%
  pivot_longer(c(mean_time_nuisance, mean_time_stage2, mean_time_boot),
               names_to = "stage", values_to = "seconds") %>%
  mutate(stage = recode(stage,
                        mean_time_nuisance = "Stage 1 (nuisances)",
                        mean_time_stage2 = "Stage 2 (final model)",
                        mean_time_boot = "Bootstrap CI"))

time_plot <- time_df %>%
  ggplot(aes(x = variant, y = seconds, fill = stage)) +
  geom_col() +
  facet_wrap(vars(family, scenario), nrow = n_distinct(time_df$family), scales = "free") +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Mean runtime per replicate",
       y = "Seconds", x = "Crossfitting procedure", fill = "Stage")

time_plot
```

## Coverage against cost: the decision plot

Marginal coverage against runtime, on a log scale — closer to the dashed
nominal-95% line is better. This is the plot that answers the pilot's
headline question: does the bootstrap buy closer-to-nominal coverage for any
arm, and at what cost?

```{r coverage-cost-plot}
coverage_cost_plot <- metrics_summary %>%
  ggplot(aes(x = mean_time_total, y = mean_marg_cov, colour = variant, shape = family)) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  geom_point(size = 3) +
  facet_wrap(~scenario, scales = "free") +
  scale_x_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  labs(title = "Marginal coverage against runtime",
       subtitle = "Closer to the dashed line is better",
       y = "Mean marginal coverage", x = "Mean runtime per replicate (s, log scale)",
       colour = "Procedure", shape = "Estimator")

coverage_cost_plot
```

## Headline table

Arranged by family, scenario, and mean marginal coverage (closest to nominal
first within each scenario).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, family, variant, mean_marg_cov, mcse_marg_cov, mean_simul_cov,
         mean_ci_len, mean_bias, mean_mse, mean_time_total) %>%
  arrange(family, scenario, desc(mean_marg_cov))

kable(headline, digits = 4,
      col.names = c("Scenario", "Family", "Procedure", "Mean marginal coverage",
                    "MCSE (marg. cov.)", "Mean simultaneous coverage",
                    "Mean CI length", "Mean bias", "Mean MSE", "Mean runtime (s)"))
```
