---
title: "Missing Data Handling for Continuous Outcomes"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(paletteer)
library(here)
library(cowplot)
library(purrr)

# paths
path <- here()
res_path <- file.path(dirname(here()), "results", "missing", "continuous")
```

```{r data, include=FALSE}
metrics <- readRDS(file.path(res_path, "cts_miss_metrics.RDS"))

metrics <- metrics %>%
  mutate(scenario = factor(scenario, levels = c(1:5)),
         n = factor(n, levels = c(500)),
         type = factor(type, levels = c("prognostic", "predictive", "both")),
         prop = factor(prop, levels = c(0.3)),
         mechanism = factor(mechanism, levels = c("MAR", "AUX", "AUX-Y")),
         method = factor(
           recode(method,
                  complete_cases = "Complete case",
                  mean_imputation = "Single mean imputation",
                  missforest = "Single forest-based imputation",
                  regression = "Single model-based imputation",
                  missing_indicator = "Missing indicators",
                  IPW = "IPW",
                  multiple_imputation = "Multiple imputation (rf)",
                  none = "Inbuilt missingness handling"),
           levels = c("Complete case", "Single mean imputation", "Single forest-based imputation", "Single model-based imputation", "Missing indicators", "IPW", "Multiple imputation (rf)", "Inbuilt missingness handling")
         ),
         model = factor(
           recode(model,
                  causal_forest = "Causal forest",
                  dr_random_forest = "DR-RandomForest",
                  dr_oracle = "DR-oracle",
                  dr_semi_oracle = "DR-semi-oracle",
                  dr_superlearner = "DR-SuperLearner"),
           levels = c("Causal forest", "DR-RandomForest", "DR-oracle", "DR-semi-oracle", "DR-SuperLearner")
         ))

metrics_summary <- metrics %>%
  group_by(scenario, n, type, prop, mechanism, method, model) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T)/sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T)/sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T)/sqrt(n()),
    .groups = "drop"
  )
```

```{r plot_generation, include=FALSE}
make_sum_plots <- function(scen, t) {
  bias_plot <- metrics_summary %>%
    filter(scenario == scen & type == t) %>%
    ggplot(aes(x = model, y = mean_bias, colour = method)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_bias - mcse_bias,
                      ymax = mean_bias + mcse_bias),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    facet_grid(~mechanism) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean bias in CATEs, scenario ", scen, ", ", t, " missingness type"),
         y = "Bias",
         x = "Model") +
  theme(axis.text = element_text(angle = 90))
  
  mse_plot <- metrics_summary %>%
    filter(scenario == scen & type == t) %>%
    ggplot(aes(x = model, y = mean_mse, colour = method)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_mse - mcse_mse,
                      ymax = mean_mse + mcse_mse),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    facet_grid(~mechanism) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean MSE in CATEs, scenario ", scen, ", ", t, " missingness type"),
         y = "MSE",
         x = "Model") +
  theme(axis.text = element_text(angle = 90))
  
  corr_plot <- metrics_summary %>%
    filter(scenario == scen & type == t) %>%
    ggplot(aes(x = model, y = mean_corr, colour = method)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(aes(ymin = mean_corr - mcse_corr,
                      ymax = mean_corr + mcse_corr),
                  position = position_dodge(width = 0.5), linewidth = 0.3) +
    facet_grid(~mechanism) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    labs(title = paste0("Mean Correlation in CATEs, scenario ", scen, ", ", t, " missingness type"),
         y = "Correlation",
         x = "Model") +
  theme(axis.text = element_text(angle = 90))
  
  list(bias_plot = bias_plot, mse_plot = mse_plot, corr_plot = corr_plot)
}

scen_types <- metrics %>% select(scenario, type) %>% unique()

sum_plots <- pmap(list(scen_types$scenario, scen_types$type), make_sum_plots)
```

In this document, 

# Scenario 1 - no HTE
$b_0 + b_1 X_1 + b_2 X_2 + W*b_W$
```{r s_1}
i <- which(scen_types$scenario == 1 & scen_types$type == "prognostic")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
```

# Scenario 2 - simple HTE binary variable
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3)$

## Prognostic missingness
```{r s_2_prog}
i <- which(scen_types$scenario == 2 & scen_types$type == "prognostic")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Predictive missingness
```{r s_2_pred}
i <- which(scen_types$scenario == 2 & scen_types$type == "predictive")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Both types of missingness
```{r s_2_both}
i <- which(scen_types$scenario == 2 & scen_types$type == "both")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

# Scenario 3 - two HTE variables
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3 + b_4 X_4)$

## Prognostic missingness
```{r s_3_prog}
i <- which(scen_types$scenario == 3 & scen_types$type == "prognostic")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Predictive missingness
```{r s_3_pred}
i <- which(scen_types$scenario == 3 & scen_types$type == "predictive")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Both types of missingness
```{r s_3_both}
i <- which(scen_types$scenario == 3 & scen_types$type == "both")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

# Scenario 4 - two way interaction
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$

## Prognostic missingness
```{r s_4_prog}
i <- which(scen_types$scenario == 4 & scen_types$type == "prognostic")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Predictive missingness
```{r s_4_pred}
i <- which(scen_types$scenario == 4 & scen_types$type == "predictive")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Both types of missingness
```{r s_4_both}
i <- which(scen_types$scenario == 4 & scen_types$type == "both")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

# Scenario 5 - non linear effect
$b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 \cos(X_4))$

## Prognostic missingness
```{r s_5_prog}
i <- which(scen_types$scenario == 5 & scen_types$type == "prognostic")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Predictive missingness
```{r s_5_pred}
i <- which(scen_types$scenario == 5 & scen_types$type == "predictive")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

## Both types of missingness
```{r s_5_both}
i <- which(scen_types$scenario == 5 & scen_types$type == "both")

sum_plots[[i]]$bias_plot
sum_plots[[i]]$mse_plot
sum_plots[[i]]$corr_plot
```

