---
title: "Interim-analysis validation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
source(here("validation/continuous/cts_val_config.R"))
```

## Does an interim finding survive the rest of the trial?

If a subgroup, a CATE ranking of covariates, or the spread of estimated
treatment effects looks real partway through a trial, does it still look real
once more data has accrued? This study fits two CATE estimators on the first
`interim_prop` of a simulated trial and checks whether what they find there —
top responder subgroups, the variance of the estimated CATEs, and the ranking
of covariates by treatment-effect variable importance (TE-VIM) — replicates
on the remaining chunk. It is a robustness check for CATE-based subgroup
discovery, not an estimator comparison.

### Design

| | |
|---|---|
| scenario | 3, continuous set (see below) |
| n | 1000, split into chunks of `n * interim_prop` and `n * (1 - interim_prop)` |
| `interim_prop` | 0.25, 0.5, 0.75 |
| runs | 100 — **300 array jobs** |
| folds | 5 below n=250 per chunk, else 10 |

**Estimators**: only `causal_forest` and `dr_random_forest`
(`R/cate_models.R`). The question is whether one estimator's own findings
persist across chunks, not which estimator is most accurate, so the
oracle/semi-oracle/SuperLearner arms this repo's other studies carry are not
run here.

### What the truth is

Scenario 3 of the continuous set is `TE_10[3]`: `tau = bW + b4 * X4`, with
`b4 = -1` and `X4 ~ N(0, 1)`. `X3` is not generated for this scenario
(`needs_X3 = FALSE`) — the study README's "continuous x binary interaction
(X3*X4)" description names scenario 5, not 3.

Two consequences used as reference lines below:

- **`Var(tau) = b4^2 * Var(X4) = 1`**, in both chunks and at every
  `interim_prop`. `bW` is recalibrated per chunk size, but it is an additive
  constant and does not enter the variance.
- **X4 is the only true HTE driver.** The other seven covariates in the
  fitted model (X1, X2, X01, X02, X03, X04, X05) are noise as far as the CATE
  is concerned, and X4 should rank as the single most important one in both
  chunks.

```{r data}
res_path <- study$res_path

metrics <- readRDS(file.path(res_path, "cts_val_metrics.RDS"))

subgroups <- metrics$subgroups %>% select(-scenario) %>% apply_labels()
variances <- metrics$variances %>% select(-scenario) %>% apply_labels()
var_imps  <- metrics$var_imps  %>% select(-scenario) %>% apply_labels()

interim_labels <- c(`0.25` = "Interim at 25% (250 -> 750)",
                    `0.5`  = "Interim at 50% (500 -> 500)",
                    `0.75` = "Interim at 75% (750 -> 250)")

subgroups <- subgroups %>%
  mutate(interim_prop = label_factor(interim_prop, interim_labels))
variances <- variances %>%
  mutate(interim_prop = label_factor(interim_prop, interim_labels))
var_imps <- var_imps %>%
  mutate(interim_prop = label_factor(interim_prop, interim_labels))

n_runs_expected <- study$n_sims
n_covariates <- length(unique(var_imps$variables))
```

## Data completeness

```{r completeness}
runs_present <- subgroups %>%
  distinct(interim_prop, model, run) %>%
  count(interim_prop, model, name = "runs") %>%
  pivot_wider(names_from = model, values_from = runs)

kable(runs_present,
      caption = paste0("Replicates present per cell (", n_runs_expected,
                       " expected)"))

na_counts <- bind_rows(
  subgroups %>% summarise(across(c(top_pval, bottom_pval), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "subgroups"),
  variances %>% summarise(across(c(vt1, vt2), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "variances"),
  var_imps %>% summarise(across(c(vi1, vi2), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "var_imps")
) %>% select(table, column, n_missing)

kable(na_counts, caption = "Missing values across all rows")
```

## Subgroup persistence

`cts_val_analysis.R` fits a tree on chunk 1's top/bottom-10%-CATE responder
groups, predicts them into chunk 2, and tests `Y ~ W * subgroup` there.
**`bottom_pval` is not usable**: the analysis script pulls the `top10`
column's interaction p-value correctly (`pvals_top[4]`) but pulls the
*intercept* p-value for `bottom_pval` (`pvals_bottom[1]`), not a subgroup
test at all. It is excluded below pending a fix to `cts_val_analysis.R` and a
re-run. Only `top_pval` — does the top-10% responder subgroup found in chunk
1 still show a treatment effect in chunk 2 — is reported.

```{r subgroup-hist}
ggplot(subgroups, aes(x = top_pval, fill = model)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  facet_wrap(~interim_prop) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme() +
  labs(title = "Distribution of top-10% subgroup p-values in chunk 2",
       x = "p-value (top-10% subgroup, W x subgroup interaction)",
       y = "Count", fill = "Model")
```

```{r subgroup-prop}
subgroup_prop <- subgroups %>%
  mutate(is_sig = top_pval < 0.05) %>%
  group_by(model, interim_prop) %>%
  summarise(runs = sum(!is.na(is_sig)),
            prop_sig = mean(is_sig, na.rm = TRUE),
            mcse = sqrt(prop_sig * (1 - prop_sig) / runs),
            .groups = "drop")

ggplot(subgroup_prop, aes(x = interim_prop, y = prop_sig, colour = model,
                          ymin = prop_sig - mcse, ymax = prop_sig + mcse)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.3), size = 2) +
  geom_errorbar(position = position_dodge(width = 0.3), linewidth = 0.3, width = 0.2) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Proportion of runs where the top-10% subgroup validates",
       subtitle = "Dashed line: nominal 5% false-positive rate. Scenario 3 has genuine HTE, so higher is better.",
       x = "Interim point", y = "Proportion with p < 0.05", colour = "Model")
```

## CATE variance stability

Does the spread of estimated treatment effects in chunk 1 (`vt1`) match the
spread in chunk 2 (`vt2`)? The dashed line is the true `Var(tau) = 1`.

```{r variance-box, fig.height=9}
tidy_variances <- variances %>%
  rename(stage1 = vt1, stage2 = vt2) %>%
  pivot_longer(cols = c(stage1, stage2), names_to = "stage", values_to = "variance")

ggplot(tidy_variances, aes(x = stage, y = variance, fill = model, colour = model)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  facet_grid(rows = vars(model), cols = vars(interim_prop)) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme() +
  labs(title = "CATE variance in early vs. late chunk",
       subtitle = "Dashed line: true Var(tau) = 1",
       x = "", y = "Variance of estimated CATE") +
  theme(legend.position = "none")
```

```{r variance-scatter, fig.height=9}
ggplot(variances, aes(x = vt1, y = vt2, colour = model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1) +
  facet_grid(rows = vars(model), cols = vars(interim_prop)) +
  drf_scale() +
  drf_theme() +
  labs(title = "CATE variance, chunk 1 vs. chunk 2, per run",
       subtitle = "Dashed line: vt1 = vt2 (perfect replication)",
       x = "Variance in chunk 1", y = "Variance in chunk 2") +
  theme(legend.position = "none")
```

```{r variance-change}
var_change_summary <- variances %>%
  group_by(model, interim_prop) %>%
  summarise(runs = sum(!is.na(var_change)),
            mean_change = mean(var_change, na.rm = TRUE),
            mcse = sd(var_change, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(var_change_summary, aes(x = interim_prop, y = mean_change, colour = model,
                               ymin = mean_change - mcse, ymax = mean_change + mcse)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.3), size = 2) +
  geom_errorbar(position = position_dodge(width = 0.3), linewidth = 0.3, width = 0.2) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean change in CATE variance, chunk 2 minus chunk 1",
       x = "Interim point", y = "Mean(vt2 - vt1)", colour = "Model")
```

## Variable-importance rank stability

`cts_val_analysis.R` ranks each covariate's TE-VIM within a chunk via
`rank(tevim)`: **rank 1 is the least important covariate, rank `r n_covariates`
is the most important.** X4 is the only true HTE driver, so a working
estimator should rank it `r n_covariates` (most important) in both chunks.

```{r rank-agreement}
rank_corr <- var_imps %>%
  group_by(model, interim_prop, run) %>%
  filter(n() >= 2) %>%
  summarise(rho = cor(vi1, vi2, method = "spearman"), .groups = "drop")

rank_corr_summary <- rank_corr %>%
  group_by(model, interim_prop) %>%
  summarise(runs = sum(!is.na(rho)),
            mean_rho = mean(rho, na.rm = TRUE),
            mcse = sd(rho, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(rank_corr_summary, aes(x = interim_prop, y = mean_rho, colour = model,
                              ymin = mean_rho - mcse, ymax = mean_rho + mcse)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.3), size = 2) +
  geom_errorbar(position = position_dodge(width = 0.3), linewidth = 0.3, width = 0.2) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Spearman correlation between chunk 1 and chunk 2 TE-VIM ranks",
       subtitle = "One value per run, across the 8 covariates. Dashed line: no agreement.",
       x = "Interim point", y = "Mean Spearman correlation (vi1 vs vi2)", colour = "Model")
```

```{r rank-mean, fig.height=9}
mean_vi <- var_imps %>%
  group_by(variables, model, interim_prop) %>%
  summarise(mean_vi1 = mean(vi1, na.rm = TRUE),
            mean_vi2 = mean(vi2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_longer(cols = c(mean_vi1, mean_vi2), names_to = "chunk", values_to = "mean_rank") %>%
  mutate(chunk = recode(chunk, mean_vi1 = "Chunk 1", mean_vi2 = "Chunk 2"),
         is_truth = variables == "X4")

ggplot(mean_vi, aes(x = variables, y = mean_rank, colour = chunk, group = chunk,
                    shape = is_truth)) +
  geom_point(size = 2) +
  geom_line(aes(group = interaction(chunk, model))) +
  facet_grid(rows = vars(model), cols = vars(interim_prop)) +
  scale_shape_manual(values = c(`TRUE` = 17, `FALSE` = 16), guide = "none") +
  drf_scale() +
  drf_theme() +
  coord_flip() +
  labs(title = "Mean TE-VIM rank per covariate, by chunk",
       subtitle = paste0("Rank 1 = least important, ", n_covariates,
                         " = most important. X4 (triangle) is the only true HTE driver."),
       x = "Covariate", y = "Mean rank", colour = "Chunk")
```

```{r rank-change, fig.height=9}
top_vars <- var_imps %>%
  group_by(variables, model, interim_prop) %>%
  summarise(mean_abs_change = mean(abs(diff), na.rm = TRUE), .groups = "drop")

ggplot(top_vars, aes(x = reorder(variables, mean_abs_change), y = mean_abs_change,
                     fill = model)) +
  geom_col(position = "dodge") +
  coord_flip() +
  facet_wrap(~interim_prop) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme() +
  labs(title = "Mean absolute change in TE-VIM rank, chunk 2 minus chunk 1",
       x = "Covariate", y = "Mean |rank change|", fill = "Model")
```

```{r x4-top-table}
x4_top <- var_imps %>%
  group_by(model, interim_prop, run) %>%
  summarise(top1 = variables[which.max(vi1)] == "X4",
            top2 = variables[which.max(vi2)] == "X4",
            .groups = "drop") %>%
  group_by(model, interim_prop) %>%
  summarise(runs = n(),
            p_top_chunk1 = mean(top1),
            p_top_chunk2 = mean(top2),
            p_top_both = mean(top1 & top2),
            .groups = "drop")

kable(x4_top, digits = 3,
      col.names = c("Model", "Interim point", "Runs", "X4 top in chunk 1",
                    "X4 top in chunk 2", "X4 top in both"),
      caption = "How often the true HTE driver (X4) is ranked most important")
```

## Headline table

```{r headline-table}
headline <- subgroup_prop %>%
  select(model, interim_prop, runs, prop_sig) %>%
  left_join(variances %>% group_by(model, interim_prop) %>%
              summarise(mean_vt1 = mean(vt1, na.rm = TRUE),
                       mean_vt2 = mean(vt2, na.rm = TRUE),
                       .groups = "drop"),
            by = c("model", "interim_prop")) %>%
  left_join(var_change_summary %>% select(model, interim_prop, mean_change),
            by = c("model", "interim_prop")) %>%
  left_join(rank_corr_summary %>% select(model, interim_prop, mean_rho),
            by = c("model", "interim_prop")) %>%
  left_join(x4_top %>% select(model, interim_prop, p_top_both),
            by = c("model", "interim_prop"))

kable(headline, digits = 3,
      col.names = c("Model", "Interim point", "Runs", "P(top subgroup sig.)",
                    "Mean vt1", "Mean vt2", "Mean var. change",
                    "Mean rank correlation", "P(X4 top in both)"))
```
