---
title: "Interim-analysis validation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 6)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
source(here("validation/continuous/cts_val_config.R"))
```

## Does an interim finding survive the rest of the trial?

If a subgroup, a CATE ranking of covariates, or the spread of estimated
treatment effects looks real partway through a trial, does it still look real
once more data has accrued? This study fits three CATE estimators on the first
`interim_prop` of a simulated trial and checks whether what they find there —
top and bottom responder subgroups, the variance of the estimated CATEs, the
ranking of covariates by variable importance, and whether the single most
important covariate still interacts with treatment — replicates on the remaining
stage. It is a robustness check for CATE-based subgroup discovery, not an
estimator comparison.

### Design

| | |
|---|---|
| scenario | 2, correlated continuous set at rho = 0.5 (see below) |
| n | 1000, one trial split into its first `n * interim_prop` participants and the rest |
| `interim_prop` | 0.25 to 0.75 in steps of 0.05 — 11 interim points |
| runs | 100 — **1100 array jobs** |
| folds | 5 below n=250 per stage, else 10 |

**Estimators**: `causal_forest`, `dr_random_forest` and `dr_superlearner`
(`R/cate_models.R`). The question is whether one estimator's own findings
persist across stages. The DR SuperLearner is crossfit, and its TE-VIMs come
from crossfit refits of its stage-2 SuperLearner on the same folds.

**Variable importance**: two measures, computed side by side for each estimator
in each stage (`cts_val_models.R`).

- **TE-VIM** — refit the second stage with each covariate dropped in turn and
  compare out-of-bag prediction error. Asks how much worse the CATE is predicted
  without that covariate.
- **Surrogate TreeSHAP** — fit a tuned xgboost surrogate to the estimated CATEs
  and take exact TreeSHAP of it, averaging |SHAP| over units ("Strategy 3" /
  indirect SHAP, Svensson et al.). Asks how much of each unit's estimated CATE
  is attributable to that covariate.

They need not agree, which is why both are carried: each gets its own
stage-1-vs-stage-2 rank comparison, and each nominates its own top covariate for
the interaction test below.

### What the truth is

Scenario 2 of the continuous set is `TE_10[2]`: `tau = bW + b4 * X4`, with
`b4 = -1` and `X4 ~ N(0, 1)`. `X3` and `X5` are generated (every scenario
draws X1–X5) but do not modify the effect. (Before the 2026-09-26 renumbering this was scenario 3,
and results produced then sit under `scenario_3/`.)

The covariates are correlated: X1–X5 and X01–X03 come from a Gaussian copula
with exchangeable latent correlation 0.5 (`R/dgm_scenarios.R`, CORRELATED
COVARIATES), X1 and X3 thresholded to binary. X04 and X05 (a three-level
categorical) are the only covariates independent of X4.

Each run is one trial of 1000, and stage 1 is its first `n * interim_prop`
participants, so a run's stage 1 at 0.25 is nested in its stage 1 at 0.30 and
curves across `interim_prop` are paired within run. (Before this rerun the two
stages were drawn as two separate trials, each with its own `bW`.)

Consequences used as reference lines below:

- **`Var(tau) = b4^2 * Var(X4) = 1`**, in both stages and at every
  `interim_prop`. The copula scales X4's latent column, so its variance is
  unchanged by the correlation.
- **X4 is the only true HTE driver.** The other nine covariates in the
  fitted model (X1, X2, X3, X5, X01–X05) are noise as far as the CATE
  is concerned, and X4 should rank as the single most important one in both
  stages, under either importance measure.
- **The noise covariates are not exchangeable.** Those correlated with X4 carry
  some of its signal, so they should rank consistently above X04/X05 in both
  stages. That structure inflates the stage-to-stage rank correlation without
  any real discovery behind it. Dropping one covariate at a time also lowers X4's TE-VIM,
  because its correlated proxies stand in for part of it.
- **X4 is therefore the right answer to the top-variable test.** A run that
  nominates X4 and then finds a significant `W x X4` interaction in stage 2 has
  done the whole job correctly. A run that nominates a proxy of X4 will still
  often find a significant *marginal* interaction (`p_cts`), because E[tau | X]
  moves with any covariate correlated with X4. The adjusted test (`p_cts_adj`)
  conditions on the other covariates and should not.

```{r data}
res_path <- study$res_path
fig_path <- here("../results/all_figures/validation")
dir.create(fig_path, showWarnings = FALSE)
metrics <- readRDS(file.path(res_path, "cts_val_metrics.RDS"))

# every estimator is shown: causal forest, DR random forest and DR SuperLearner.
# scenario and the covariate correlation rho take one value each, so they are
# dropped; any_of() because older metrics files carry no rho column.
subgroups <- metrics$subgroups %>% select(-any_of(c("scenario", "rho"))) %>% apply_labels()
variances <- metrics$variances %>% select(-any_of(c("scenario", "rho"))) %>% apply_labels()
var_imps  <- metrics$var_imps  %>% select(-any_of(c("scenario", "rho"))) %>% apply_labels()
top_var   <- metrics$top_var   %>% select(-any_of(c("scenario", "rho"))) %>% apply_labels()

n_models <- length(unique(subgroups$model))

# interim_prop stays numeric so it can carry an axis; `interim_lab` is the short
# form for the few plots that still need one panel per interim point. Built from
# the data rather than a literal vector - there are 11 of them now, and the
# grid is free to change again.
MEASURE_LABELS <- c(tevim = "TE-VIM", shap = "TreeSHAP")

add_interim_lab <- function(d) {
  d %>% mutate(interim_lab = factor(paste0(round(interim_prop * 100), "%"),
                                    levels = paste0(round(sort(unique(interim_prop)) * 100), "%")))
}

subgroups <- add_interim_lab(subgroups)
variances <- add_interim_lab(variances)
var_imps  <- add_interim_lab(var_imps)  %>%
  mutate(measure = label_factor(measure, MEASURE_LABELS))
top_var   <- add_interim_lab(top_var)   %>%
  mutate(measure = label_factor(measure, MEASURE_LABELS))

n_runs_expected <- study$n_sims
n_covariates <- length(unique(var_imps$variables))

# the three interim points the study ran before the grid was refined - used to
# keep the summary tables to a readable length
landmarks <- c(0.25, 0.5, 0.75)
```

## Data completeness

```{r completeness}
runs_present <- subgroups %>%
  distinct(interim_prop, model, run) %>%
  count(interim_prop, model, name = "runs") %>%
  pivot_wider(names_from = model, values_from = runs)

kable(runs_present,
      caption = paste0("Replicates present per cell (", n_runs_expected,
                       " expected)"))

# na_counts <- bind_rows(
#   subgroups %>% summarise(across(c(top_pval, bottom_pval), ~ sum(is.na(.x)))) %>%
#     pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
#     mutate(table = "subgroups"),
#   variances %>% summarise(across(c(vt1, vt2), ~ sum(is.na(.x)))) %>%
#     pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
#     mutate(table = "variances"),
#   var_imps %>% summarise(across(c(vi1, vi2), ~ sum(is.na(.x)))) %>%
#     pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
#     mutate(table = "var_imps"),
#   top_var %>% summarise(across(c(p_cts, p_split), ~ sum(is.na(.x)))) %>%
#     pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
#     mutate(table = "top_var")
# ) %>% select(table, column, n_missing)
# 
# kable(na_counts, caption = "Missing values across all rows")
```


## Subgroup persistence

`cts_val_analysis.R` fits a tree on stage 1's top/bottom-10%-CATE responder
groups, predicts them into stage 2, and tests the `W x subgroup` interaction
there.


```{r subgroup-long}
subgroups_long <- subgroups %>%
  pivot_longer(c(top_pval, bottom_pval), names_to = "which", values_to = "pval") %>%
  mutate(which = recode(which, top_pval = "Top 10%", bottom_pval = "Bottom 10%"),
         which = factor(which, levels = c("Top 10%", "Bottom 10%")))
```

```{r subgroup-ecdf, fig.width=10, fig.height=8}
ggplot(subgroups_long, aes(x = pval, colour = interim_prop, group = interim_prop)) +
  stat_ecdf(geom = "step", linewidth = 0.5, na.rm = TRUE) +
  geom_vline(xintercept = 0.1, linetype = "dashed") +
  facet_grid(model ~ which) +
  scale_colour_viridis_c(option = "viridis") +
  drf_theme() +
  labs(title = "Cumulative distribution of stage 2 subgroup p-values",
       x = "p-value (W x subgroup interaction)", y = "Cumulative proportion",
       colour = "Interim\nproportion")
ggsave("cts_val_sg_cumu.png", path = fig_path, width = 20, height = 6 * n_models,
       units = "cm")
```

```{r subgroup-prop, fig.width=10, fig.height=8}
subgroup_prop <- subgroups_long %>%
  tidyr::crossing(threshold = c(0.05, 0.1)) %>%
  mutate(is_sig = pval < threshold,
         threshold = factor(threshold, levels = c(0.05, 0.1), labels = c("0.05", "0.10"))) %>%
  group_by(model, which, interim_prop, threshold) %>%
  summarise(runs = sum(!is.na(is_sig)),
            prop_sig = mean(is_sig, na.rm = TRUE),
            mcse = sqrt(prop_sig * (1 - prop_sig) / runs),
            .groups = "drop")

ggplot(subgroup_prop, aes(x = interim_prop, y = prop_sig, colour = threshold,
                          ymin = prop_sig - qnorm(0.975) * mcse,
                          ymax = prop_sig + qnorm(0.975) * mcse)) +
  geom_ribbon(aes(fill = threshold), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_grid(model ~ which) +
  scale_x_continuous(breaks = sort(unique(subgroup_prop$interim_prop))) +
  # 0.05 steps as before, but over the whole range: the other estimators can
  # fall below the 0.45 the DR random forest alone never went under
  scale_y_continuous(breaks = seq(0, 1, 0.05)) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Proportion of successful stage 2 interaction tests",
       x = "Stage 1 proportion", y = "Proportion with p < threshold",
       colour = "Threshold", fill = "Threshold")
ggsave("cts_val_sg_props.png", path = fig_path, width = 20, height = 6 * n_models,
       units = "cm")

```

## CATE variance stability

Does the spread of estimated treatment effects in stage 1 (`vt1`) match the
spread in stage 2 (`vt2`)? The dashed line is the true `Var(tau) = 1`.

```{r variance-box, fig.height=3.5 * n_models}
tidy_variances <- variances %>%
  rename(stage1 = vt1, stage2 = vt2) %>%
  pivot_longer(cols = c(stage1, stage2), names_to = "stage", values_to = "variance") %>%
  mutate(stage = recode(stage, stage1 = "stage 1", stage2 = "stage 2"))

ggplot(tidy_variances, aes(x = interim_lab, y = variance, fill = stage)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_boxplot(alpha = 0.6, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  facet_wrap(~model, ncol = 1) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "CATE variance in early vs. late stage",
       subtitle = "Dashed line: true Var(tau) = 1. stage 1 grows and stage 2 shrinks left to right.",
       x = "Interim proportion", y = "Variance of estimated CATE", fill = "stage")
```

```{r variance-scatter, fig.height=9}
ggplot(variances, aes(x = vt1, y = vt2, colour = model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1) +
  facet_wrap(~interim_lab, ncol = 4) +
  drf_scale() +
  drf_theme() +
  theme(aspect.ratio = 1) +
  labs(title = "CATE variance, stage 1 vs. stage 2, per run",
       subtitle = "Dashed line: vt1 = vt2 (perfect replication)",
       x = "Variance in stage 1", y = "Variance in stage 2") +
  theme(legend.position = "bottom")
```

```{r variance-change}
var_change_summary <- variances %>%
  group_by(model, interim_prop) %>%
  summarise(runs = sum(!is.na(var_change)),
            mean_change = mean(var_change, na.rm = TRUE),
            mcse = sd(var_change, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(var_change_summary, aes(x = interim_prop, y = mean_change, colour = model,
                               ymin = mean_change - qnorm(0.975) * mcse,
                               ymax = mean_change + qnorm(0.975) * mcse)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = sort(unique(var_change_summary$interim_prop))) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean change in CATE variance, stage 2 minus stage 1",
       x = "Interim proportion", y = "Mean(vt2 - vt1)",
       colour = "Model", fill = "Model")
```

## Variable-importance rank stability

`cts_val_analysis.R` ranks each covariate's importance within a stage via
`rank()`: **rank 1 is the least important covariate, rank `r n_covariates`
is the most important.** X4 is the only true HTE driver, so a working
estimator should rank it `r n_covariates` (most important) in both stages,
under both measures.

```{r rank-agreement, fig.height=5}
rank_corr <- var_imps %>%
  group_by(model, measure, interim_prop, run) %>%
  filter(n() >= 2) %>%
  summarise(rho = cor(vi1, vi2, method = "spearman"), .groups = "drop")

rank_corr_summary <- rank_corr %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = sum(!is.na(rho)),
            mean_rho = mean(rho, na.rm = TRUE),
            mcse = sd(rho, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(rank_corr_summary, aes(x = interim_prop, y = mean_rho, colour = measure,
                              ymin = mean_rho - qnorm(0.975) * mcse,
                              ymax = mean_rho + qnorm(0.975) * mcse)) +
  geom_ribbon(aes(fill = measure), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_wrap(~model) +
  scale_x_continuous(breaks = sort(unique(rank_corr_summary$interim_prop))) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Spearman correlation between stage 1 and stage 2 importance ranks",
       x = "Interim proportion", y = "Mean Spearman correlation between stages",
       colour = "Measure", fill = "Measure")
```

```{r rank-mean, fig.height=3.5 * n_models}
mean_vi <- var_imps %>%
  group_by(variables, model, measure, interim_prop) %>%
  summarise(mean_vi1 = mean(vi1, na.rm = TRUE),
            mean_vi2 = mean(vi2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_longer(cols = c(mean_vi1, mean_vi2), names_to = "stage", values_to = "mean_rank") %>%
  mutate(stage = recode(stage, mean_vi1 = "Stage 1", mean_vi2 = "Stage 2"))

ggplot(mean_vi, aes(x = interim_prop, y = mean_rank, colour = variables,
                    linetype = stage)) +
  geom_line() +
  facet_grid(model ~ measure) +
  scale_x_continuous(breaks = sort(unique(mean_vi$interim_prop))) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean importance rank per covariate, by stage",
       subtitle = paste0("Rank 1 = least important, ", n_covariates,
                         " = most important. X4 is the only true HTE driver."),
       x = "Interim proportion", y = "Mean rank",
       colour = "Covariate", linetype = "stage")
```

```{r rank-change, fig.height=3.5 * n_models}
top_vars <- var_imps %>%
  group_by(variables, model, measure, interim_prop) %>%
  summarise(mean_abs_change = mean(abs(diff), na.rm = TRUE), .groups = "drop")

ggplot(top_vars, aes(x = interim_prop, y = mean_abs_change, colour = variables)) +
  geom_line() +
  facet_grid(model ~ measure) +
  scale_x_continuous(breaks = sort(unique(top_vars$interim_prop))) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean absolute change in importance rank, stage 2 minus stage 1",
       x = "Interim proportion", y = "Mean |rank change|", colour = "Covariate")
```

```{r x4-top, fig.height=5}
x4_top <- var_imps %>%
  group_by(model, measure, interim_prop, run) %>%
  summarise(top1 = variables[which.max(vi1)] == "X4",
            top2 = variables[which.max(vi2)] == "X4",
            .groups = "drop") %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = n(),
            p_top_stage1 = mean(top1),
            p_top_stage2 = mean(top2),
            p_top_both = mean(top1 & top2),
            .groups = "drop")

ggplot(x4_top, aes(x = interim_prop, y = p_top_both, colour = measure)) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_wrap(~model) +
  scale_x_continuous(breaks = sort(unique(x4_top$interim_prop))) +
  ylim(0, 1) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "How often the true HTE driver (X4) is ranked most important in both stages",
       x = "Interim proportion", y = "Proportion of runs", colour = "Measure")
```

```{r x4-top-table}
kable(x4_top %>% filter(interim_prop %in% landmarks), digits = 3,
      col.names = c("Model", "Measure", "Interim point", "Runs",
                    "X4 top in stage 1", "X4 top in stage 2", "X4 top in both"),
      caption = "How often X4 is ranked most important, at the three landmark interim points")
```

## Does the top-ranked driver replicate?

The point of ranking covariates is whether the winner means anything. For each
estimator and each importance measure, `cts_val_analysis.R` takes stage 1's
most important covariate and tests its treatment interaction on the remaining
trial participants — three ways: `p_cts`, the continuous `W x X_top` interaction;
`p_cts_adj`, the `W:X_top` coefficient of `Y ~ W * (all covariates)`, i.e. the
same interaction holding the other covariates fixed; and `p_split`, the
marginal test after dichotomising `X_top` at its stage-2 median.

```{r top-var-long}
top_var_long <- top_var %>%
  pivot_longer(c(p_cts, p_cts_adj, p_split), names_to = "form", values_to = "pval") %>%
  mutate(form = recode(form, p_cts = "Continuous W x X",
                       p_cts_adj = "Continuous W x X, adjusted",
                       p_split = "Median split"))

top_var_prop <- top_var_long %>%
  mutate(is_sig = pval < 0.05) %>%
  group_by(model, measure, form, interim_prop) %>%
  summarise(runs = sum(!is.na(is_sig)),
            prop_sig = mean(is_sig, na.rm = TRUE),
            mcse = sqrt(prop_sig * (1 - prop_sig) / runs),
            .groups = "drop")
```

```{r top-var-prop, fig.height=10}
ggplot(top_var_prop, aes(x = interim_prop, y = prop_sig, colour = model,
                         ymin = prop_sig - qnorm(0.975) * mcse,
                         ymax = prop_sig + qnorm(0.975) * mcse)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_grid(form ~ measure) +
  scale_x_continuous(breaks = sort(unique(top_var_prop$interim_prop))) +
  # coord_cartesian, not ylim(): ylim() drops any ribbon that pokes past 0 or 1
  coord_cartesian(ylim = c(0, 1)) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Proportion of runs where the interim's top covariate still interacts in stage 2",
       subtitle = "Ribbon: 95% Monte Carlo interval. Dashed line: nominal 5%. The continuous form should dominate the median split.",
       x = "Interim proportion", y = "Proportion with p < 0.05",
       colour = "Model", fill = "Model")
```

The median split throws away information and, for the already-binary X1, X3,
X04 and X05, is not a split at all — so the gap between the continuous and
median-split panels is the cost of dichotomising a discovered driver rather
than testing it as it stands. The gap between the continuous and adjusted
panels is mostly how much of the "replication" comes from nominating a proxy
of X4 that interacts with treatment only through its correlation with X4 —
partly, too, the power that adjusting for correlated covariates costs when X4
itself is nominated.

```{r top-var-agreement}
top_var_agree <- top_var %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = n(),
            p_x_top_is_X4 = mean(x_top == "X4"),
            p_stages_agree = mean(x_top == x_top2),
            .groups = "drop")

ggplot(top_var_agree, aes(x = interim_prop, colour = model)) +
  geom_line(aes(y = p_x_top_is_X4, linetype = "Nominates X4")) +
  geom_line(aes(y = p_stages_agree, linetype = "stages agree")) +
  facet_wrap(~measure) +
  scale_x_continuous(breaks = sort(unique(top_var_agree$interim_prop))) +
  ylim(0, 1) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Does the interim nominate the right covariate?",
       subtitle = "'Nominates X4': stage 1's top covariate is the true driver. 'stages agree': both stages pick the same one, right or wrong.",
       x = "Interim proportion", y = "Proportion of runs",
       colour = "Model", linetype = NULL)
```

```{r top-var-table}
kable(top_var_agree %>% filter(interim_prop %in% landmarks), digits = 3,
      col.names = c("Model", "Measure", "Interim point", "Runs",
                    "Nominates X4", "stages agree"),
      caption = "Which covariate the interim nominates, at the three landmark interim points")
```

## Headline table

At the three landmark interim points, to keep it readable — every panel above
carries all eleven.

```{r headline-table}
# `runs` is counted per subgroup and the two can differ (interaction_pval
# returns NA more often for bottom10), so it is joined on separately rather than
# carried through the pivot, where it would split each cell into two rows.
runs_per_cell <- subgroups %>%
  group_by(model, interim_prop) %>%
  summarise(runs = n(), .groups = "drop")

# subgroup_prop carries both thresholds; the table reports p < 0.05, as every
# other column does - pivoting both would make each cell a list
headline <- subgroup_prop %>%
  filter(interim_prop %in% landmarks, threshold == "0.05") %>%
  select(model, which, interim_prop, prop_sig) %>%
  pivot_wider(names_from = which, values_from = prop_sig, names_prefix = "sig_") %>%
  left_join(runs_per_cell, by = c("model", "interim_prop")) %>%
  relocate(runs, .after = interim_prop) %>%
  left_join(variances %>% group_by(model, interim_prop) %>%
              summarise(mean_vt1 = mean(vt1, na.rm = TRUE),
                       mean_vt2 = mean(vt2, na.rm = TRUE),
                       .groups = "drop"),
            by = c("model", "interim_prop")) %>%
  left_join(var_change_summary %>% select(model, interim_prop, mean_change),
            by = c("model", "interim_prop")) %>%
  left_join(rank_corr_summary %>%
              select(model, measure, interim_prop, mean_rho) %>%
              pivot_wider(names_from = measure, values_from = mean_rho,
                          names_prefix = "rho_"),
            by = c("model", "interim_prop")) %>%
  left_join(top_var_prop %>%
              filter(form == "Continuous W x X") %>%
              select(model, measure, interim_prop, prop_sig) %>%
              pivot_wider(names_from = measure, values_from = prop_sig,
                          names_prefix = "topvar_"),
            by = c("model", "interim_prop")) %>%
  arrange(model, interim_prop)

kable(headline, digits = 3,
      caption = paste("Landmark summary. sig_* = proportion of runs where that",
                      "subgroup validates; rho_* = mean rank correlation between",
                      "stages; topvar_* = proportion where the nominated top",
                      "covariate still interacts (continuous form)."))
```
