---
title: "Interim-analysis validation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
source(here("validation/continuous/cts_val_config.R"))
```

## Does an interim finding survive the rest of the trial?

If a subgroup, a CATE ranking of covariates, or the spread of estimated
treatment effects looks real partway through a trial, does it still look real
once more data has accrued? This study fits two CATE estimators on the first
`interim_prop` of a simulated trial and checks whether what they find there —
top and bottom responder subgroups, the variance of the estimated CATEs, the
ranking of covariates by variable importance, and whether the single most
important covariate still interacts with treatment — replicates on the remaining
chunk. It is a robustness check for CATE-based subgroup discovery, not an
estimator comparison.

### Design

| | |
|---|---|
| scenario | 3, continuous set (see below) |
| n | 1000, split into chunks of `n * interim_prop` and `n * (1 - interim_prop)` |
| `interim_prop` | 0.25 to 0.75 in steps of 0.05 — 11 interim points |
| runs | 100 — **1100 array jobs** |
| folds | 5 below n=250 per chunk, else 10 |

**Estimators**: only `causal_forest` and `dr_random_forest`
(`R/cate_models.R`). The question is whether one estimator's own findings
persist across chunks (can add superlearner for larger interim proportions)

**Variable importance**: two measures, computed side by side for each estimator
in each chunk (`cts_val_models.R`).

- **TE-VIM** — refit the second stage with each covariate dropped in turn and
  compare out-of-bag prediction error. Asks how much worse the CATE is predicted
  without that covariate.
- **Surrogate TreeSHAP** — fit a tuned xgboost surrogate to the estimated CATEs
  and take exact TreeSHAP of it, averaging |SHAP| over units ("Strategy 3" /
  indirect SHAP, Svensson et al.). Asks how much of each unit's estimated CATE
  is attributable to that covariate.

They need not agree, which is why both are carried: each gets its own
chunk-1-vs-chunk-2 rank comparison, and each nominates its own top covariate for
the interaction test below.

### What the truth is

Scenario 3 of the continuous set is `TE_10[3]`: `tau = bW + b4 * X4`, with
`b4 = -1` and `X4 ~ N(0, 1)`. `X3` is not generated for this scenario
(`needs_X3 = FALSE`) — the study README's "continuous x binary interaction
(X3*X4)" description names scenario 5, not 3.

Three consequences used as reference lines below:

- **`Var(tau) = b4^2 * Var(X4) = 1`**, in both chunks and at every
  `interim_prop`. `bW` is recalibrated per chunk size, but it is an additive
  constant and does not enter the variance.
- **X4 is the only true HTE driver.** The other seven covariates in the
  fitted model (X1, X2, X01, X02, X03, X04, X05) are noise as far as the CATE
  is concerned, and X4 should rank as the single most important one in both
  chunks, under either importance measure.
- **X4 is therefore the right answer to the top-variable test.** A run that
  nominates X4 and then finds a significant `W x X4` interaction in chunk 2 has
  done the whole job correctly.

```{r data}
res_path <- study$res_path

metrics <- readRDS(file.path(res_path, "cts_val_metrics.RDS"))

subgroups <- metrics$subgroups %>% select(-scenario) %>% apply_labels()
variances <- metrics$variances %>% select(-scenario) %>% apply_labels()
var_imps  <- metrics$var_imps  %>% select(-scenario) %>% apply_labels()
top_var   <- metrics$top_var   %>% select(-scenario) %>% apply_labels()

# interim_prop stays numeric so it can carry an axis; `interim_lab` is the short
# form for the few plots that still need one panel per interim point. Built from
# the data rather than a literal vector - there are 11 of them now, and the
# grid is free to change again.
MEASURE_LABELS <- c(tevim = "TE-VIM", shap = "TreeSHAP")

add_interim_lab <- function(d) {
  d %>% mutate(interim_lab = factor(paste0(round(interim_prop * 100), "%"),
                                    levels = paste0(round(sort(unique(interim_prop)) * 100), "%")))
}

subgroups <- add_interim_lab(subgroups)
variances <- add_interim_lab(variances)
var_imps  <- add_interim_lab(var_imps)  %>%
  mutate(measure = label_factor(measure, MEASURE_LABELS))
top_var   <- add_interim_lab(top_var)   %>%
  mutate(measure = label_factor(measure, MEASURE_LABELS))

n_runs_expected <- study$n_sims
n_covariates <- length(unique(var_imps$variables))

# the three interim points the study ran before the grid was refined - used to
# keep the summary tables to a readable length
landmarks <- c(0.25, 0.5, 0.75)
```

## Data completeness

```{r completeness}
runs_present <- subgroups %>%
  distinct(interim_prop, model, run) %>%
  count(interim_prop, model, name = "runs") %>%
  pivot_wider(names_from = model, values_from = runs)

kable(runs_present,
      caption = paste0("Replicates present per cell (", n_runs_expected,
                       " expected)"))

na_counts <- bind_rows(
  subgroups %>% summarise(across(c(top_pval, bottom_pval), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "subgroups"),
  variances %>% summarise(across(c(vt1, vt2), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "variances"),
  var_imps %>% summarise(across(c(vi1, vi2), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "var_imps"),
  top_var %>% summarise(across(c(p_cts, p_split), ~ sum(is.na(.x)))) %>%
    pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
    mutate(table = "top_var")
) %>% select(table, column, n_missing)

kable(na_counts, caption = "Missing values across all rows")
```

An NA in `subgroups` or `top_var` is not a failure: `interaction_pval()`
(`cts_val_models.R`) returns NA when the covariate it is handed carries no
contrast in chunk 2 — most often because the chunk-1 tree predicted no
`bottom10` leaf into chunk 2 at all.

## Subgroup persistence

`cts_val_analysis.R` fits a tree on chunk 1's top/bottom-10%-CATE responder
groups, predicts them into chunk 2, and tests the `W x subgroup` interaction
there.

Both subgroups are reported. `bottom_pval` used to be unusable — the analysis
script pulled the *intercept* p-value for it rather than the interaction — and
was excluded from earlier versions of this report. That is fixed
(`interaction_pval()` now indexes the coefficient by name), so these are the
first results in which the bottom-10% group is a real test.

```{r subgroup-long}
subgroups_long <- subgroups %>%
  pivot_longer(c(top_pval, bottom_pval), names_to = "which", values_to = "pval") %>%
  mutate(which = recode(which, top_pval = "Top 10%", bottom_pval = "Bottom 10%"),
         which = factor(which, levels = c("Top 10%", "Bottom 10%")))
```

```{r subgroup-ecdf}
ggplot(subgroups_long, aes(x = pval, colour = interim_prop, group = interim_prop)) +
  stat_ecdf(geom = "step", linewidth = 0.5, na.rm = TRUE) +
  geom_vline(xintercept = 0.05, linetype = "dashed") +
  facet_grid(model ~ which) +
  scale_colour_viridis_c(option = "viridis") +
  drf_theme() +
  labs(title = "Cumulative distribution of chunk-2 subgroup p-values",
       subtitle = "One curve per interim point. The height where a curve crosses the dashed line is that cell's proportion significant.",
       x = "p-value (W x subgroup interaction)", y = "Cumulative proportion of runs",
       colour = "Interim\nproportion")
```

```{r subgroup-prop}
subgroup_prop <- subgroups_long %>%
  mutate(is_sig = pval < 0.05) %>%
  group_by(model, which, interim_prop) %>%
  summarise(runs = sum(!is.na(is_sig)),
            prop_sig = mean(is_sig, na.rm = TRUE),
            mcse = sqrt(prop_sig * (1 - prop_sig) / runs),
            .groups = "drop")

ggplot(subgroup_prop, aes(x = interim_prop, y = prop_sig, colour = model,
                          ymin = prop_sig - mcse, ymax = prop_sig + mcse)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_wrap(~which) +
  scale_x_continuous(breaks = sort(unique(subgroup_prop$interim_prop))) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Proportion of runs where the interim subgroup validates",
       subtitle = "Ribbon: +/- 1 Monte Carlo SE. Dashed line: nominal 5% false-positive rate. Scenario 3 has genuine HTE, so higher is better.",
       x = "Interim proportion", y = "Proportion with p < 0.05",
       colour = "Model", fill = "Model")
```

## CATE variance stability

Does the spread of estimated treatment effects in chunk 1 (`vt1`) match the
spread in chunk 2 (`vt2`)? The dashed line is the true `Var(tau) = 1`.

```{r variance-box, fig.height=9}
tidy_variances <- variances %>%
  rename(stage1 = vt1, stage2 = vt2) %>%
  pivot_longer(cols = c(stage1, stage2), names_to = "stage", values_to = "variance") %>%
  mutate(stage = recode(stage, stage1 = "Chunk 1", stage2 = "Chunk 2"))

ggplot(tidy_variances, aes(x = interim_lab, y = variance, fill = stage)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_boxplot(alpha = 0.6, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  facet_wrap(~model, ncol = 1) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "CATE variance in early vs. late chunk",
       subtitle = "Dashed line: true Var(tau) = 1. Chunk 1 grows and chunk 2 shrinks left to right.",
       x = "Interim proportion", y = "Variance of estimated CATE", fill = "Chunk")
```

```{r variance-scatter, fig.height=9}
ggplot(variances, aes(x = vt1, y = vt2, colour = model)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.4, size = 1) +
  facet_wrap(~interim_lab, ncol = 4) +
  drf_scale() +
  drf_theme() +
  theme(aspect.ratio = 1) +
  labs(title = "CATE variance, chunk 1 vs. chunk 2, per run",
       subtitle = "Dashed line: vt1 = vt2 (perfect replication)",
       x = "Variance in chunk 1", y = "Variance in chunk 2") +
  theme(legend.position = "bottom")
```

```{r variance-change}
var_change_summary <- variances %>%
  group_by(model, interim_prop) %>%
  summarise(runs = sum(!is.na(var_change)),
            mean_change = mean(var_change, na.rm = TRUE),
            mcse = sd(var_change, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(var_change_summary, aes(x = interim_prop, y = mean_change, colour = model,
                               ymin = mean_change - mcse, ymax = mean_change + mcse)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_x_continuous(breaks = sort(unique(var_change_summary$interim_prop))) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean change in CATE variance, chunk 2 minus chunk 1",
       x = "Interim proportion", y = "Mean(vt2 - vt1)",
       colour = "Model", fill = "Model")
```

## Variable-importance rank stability

`cts_val_analysis.R` ranks each covariate's importance within a chunk via
`rank()`: **rank 1 is the least important covariate, rank `r n_covariates`
is the most important.** X4 is the only true HTE driver, so a working
estimator should rank it `r n_covariates` (most important) in both chunks,
under both measures.

```{r rank-agreement}
rank_corr <- var_imps %>%
  group_by(model, measure, interim_prop, run) %>%
  filter(n() >= 2) %>%
  summarise(rho = cor(vi1, vi2, method = "spearman"), .groups = "drop")

rank_corr_summary <- rank_corr %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = sum(!is.na(rho)),
            mean_rho = mean(rho, na.rm = TRUE),
            mcse = sd(rho, na.rm = TRUE) / sqrt(runs),
            .groups = "drop")

ggplot(rank_corr_summary, aes(x = interim_prop, y = mean_rho, colour = model,
                              ymin = mean_rho - mcse, ymax = mean_rho + mcse)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_wrap(~measure) +
  scale_x_continuous(breaks = sort(unique(rank_corr_summary$interim_prop))) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Spearman correlation between chunk 1 and chunk 2 importance ranks",
       subtitle = paste0("One value per run, across the ", n_covariates,
                         " covariates. Dashed line: no agreement."),
       x = "Interim proportion", y = "Mean Spearman correlation (vi1 vs vi2)",
       colour = "Model", fill = "Model")
```

```{r rank-mean, fig.height=9}
mean_vi <- var_imps %>%
  group_by(variables, model, measure, interim_prop) %>%
  summarise(mean_vi1 = mean(vi1, na.rm = TRUE),
            mean_vi2 = mean(vi2, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_longer(cols = c(mean_vi1, mean_vi2), names_to = "chunk", values_to = "mean_rank") %>%
  mutate(chunk = recode(chunk, mean_vi1 = "Chunk 1", mean_vi2 = "Chunk 2"))

ggplot(mean_vi, aes(x = interim_prop, y = mean_rank, colour = variables,
                    linetype = chunk)) +
  geom_line() +
  facet_grid(model ~ measure) +
  scale_x_continuous(breaks = sort(unique(mean_vi$interim_prop))) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean importance rank per covariate, by chunk",
       subtitle = paste0("Rank 1 = least important, ", n_covariates,
                         " = most important. X4 is the only true HTE driver."),
       x = "Interim proportion", y = "Mean rank",
       colour = "Covariate", linetype = "Chunk")
```

```{r rank-change, fig.height=9}
top_vars <- var_imps %>%
  group_by(variables, model, measure, interim_prop) %>%
  summarise(mean_abs_change = mean(abs(diff), na.rm = TRUE), .groups = "drop")

ggplot(top_vars, aes(x = interim_prop, y = mean_abs_change, colour = variables)) +
  geom_line() +
  facet_grid(model ~ measure) +
  scale_x_continuous(breaks = sort(unique(top_vars$interim_prop))) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Mean absolute change in importance rank, chunk 2 minus chunk 1",
       x = "Interim proportion", y = "Mean |rank change|", colour = "Covariate")
```

```{r x4-top}
x4_top <- var_imps %>%
  group_by(model, measure, interim_prop, run) %>%
  summarise(top1 = variables[which.max(vi1)] == "X4",
            top2 = variables[which.max(vi2)] == "X4",
            .groups = "drop") %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = n(),
            p_top_chunk1 = mean(top1),
            p_top_chunk2 = mean(top2),
            p_top_both = mean(top1 & top2),
            .groups = "drop")

ggplot(x4_top, aes(x = interim_prop, y = p_top_both, colour = model)) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_wrap(~measure) +
  scale_x_continuous(breaks = sort(unique(x4_top$interim_prop))) +
  ylim(0, 1) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "How often the true HTE driver (X4) is ranked most important in both chunks",
       x = "Interim proportion", y = "Proportion of runs", colour = "Model")
```

```{r x4-top-table}
kable(x4_top %>% filter(interim_prop %in% landmarks), digits = 3,
      col.names = c("Model", "Measure", "Interim point", "Runs",
                    "X4 top in chunk 1", "X4 top in chunk 2", "X4 top in both"),
      caption = "How often X4 is ranked most important, at the three landmark interim points")
```

## Does the top-ranked driver replicate?

The point of ranking covariates is whether the winner means anything. For each
estimator and each importance measure, `cts_val_analysis.R` takes chunk 1's
most important covariate and tests its treatment interaction on the remaining
trial participants — two ways: `p_cts`, the continuous `W x X_top` interaction,
and `p_split`, the same test after dichotomising `X_top` at its chunk-2 median.

```{r top-var-long}
top_var_long <- top_var %>%
  pivot_longer(c(p_cts, p_split), names_to = "form", values_to = "pval") %>%
  mutate(form = recode(form, p_cts = "Continuous W x X", p_split = "Median split"))

top_var_prop <- top_var_long %>%
  mutate(is_sig = pval < 0.05) %>%
  group_by(model, measure, form, interim_prop) %>%
  summarise(runs = sum(!is.na(is_sig)),
            prop_sig = mean(is_sig, na.rm = TRUE),
            mcse = sqrt(prop_sig * (1 - prop_sig) / runs),
            .groups = "drop")
```

```{r top-var-prop, fig.height=9}
ggplot(top_var_prop, aes(x = interim_prop, y = prop_sig, colour = model,
                         ymin = prop_sig - mcse, ymax = prop_sig + mcse)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_ribbon(aes(fill = model), alpha = 0.15, colour = NA) +
  geom_line() +
  geom_point(size = 1.5) +
  facet_grid(form ~ measure) +
  scale_x_continuous(breaks = sort(unique(top_var_prop$interim_prop))) +
  ylim(0, 1) +
  drf_scale() +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Proportion of runs where the interim's top covariate still interacts in chunk 2",
       subtitle = "Ribbon: +/- 1 Monte Carlo SE. Dashed line: nominal 5%. The continuous form should dominate the median split.",
       x = "Interim proportion", y = "Proportion with p < 0.05",
       colour = "Model", fill = "Model")
```

The median split throws away information and, for the already-binary X01-X05,
is not a split at all — so the gap between the two panels is the cost of
dichotomising a discovered driver rather than testing it as it stands.

```{r top-var-agreement}
top_var_agree <- top_var %>%
  group_by(model, measure, interim_prop) %>%
  summarise(runs = n(),
            p_x_top_is_X4 = mean(x_top == "X4"),
            p_chunks_agree = mean(x_top == x_top2),
            .groups = "drop")

ggplot(top_var_agree, aes(x = interim_prop, colour = model)) +
  geom_line(aes(y = p_x_top_is_X4, linetype = "Nominates X4")) +
  geom_line(aes(y = p_chunks_agree, linetype = "Chunks agree")) +
  facet_wrap(~measure) +
  scale_x_continuous(breaks = sort(unique(top_var_agree$interim_prop))) +
  ylim(0, 1) +
  drf_scale() +
  drf_theme(rotate_x = TRUE) +
  labs(title = "Does the interim nominate the right covariate?",
       subtitle = "'Nominates X4': chunk 1's top covariate is the true driver. 'Chunks agree': both chunks pick the same one, right or wrong.",
       x = "Interim proportion", y = "Proportion of runs",
       colour = "Model", linetype = NULL)
```

```{r top-var-table}
kable(top_var_agree %>% filter(interim_prop %in% landmarks), digits = 3,
      col.names = c("Model", "Measure", "Interim point", "Runs",
                    "Nominates X4", "Chunks agree"),
      caption = "Which covariate the interim nominates, at the three landmark interim points")
```

## Headline table

At the three landmark interim points, to keep it readable — every panel above
carries all eleven.

```{r headline-table}
# `runs` is counted per subgroup and the two can differ (interaction_pval
# returns NA more often for bottom10), so it is joined on separately rather than
# carried through the pivot, where it would split each cell into two rows.
runs_per_cell <- subgroups %>%
  group_by(model, interim_prop) %>%
  summarise(runs = n(), .groups = "drop")

headline <- subgroup_prop %>%
  filter(interim_prop %in% landmarks) %>%
  select(model, which, interim_prop, prop_sig) %>%
  pivot_wider(names_from = which, values_from = prop_sig, names_prefix = "sig_") %>%
  left_join(runs_per_cell, by = c("model", "interim_prop")) %>%
  relocate(runs, .after = interim_prop) %>%
  left_join(variances %>% group_by(model, interim_prop) %>%
              summarise(mean_vt1 = mean(vt1, na.rm = TRUE),
                       mean_vt2 = mean(vt2, na.rm = TRUE),
                       .groups = "drop"),
            by = c("model", "interim_prop")) %>%
  left_join(var_change_summary %>% select(model, interim_prop, mean_change),
            by = c("model", "interim_prop")) %>%
  left_join(rank_corr_summary %>%
              select(model, measure, interim_prop, mean_rho) %>%
              pivot_wider(names_from = measure, values_from = mean_rho,
                          names_prefix = "rho_"),
            by = c("model", "interim_prop")) %>%
  left_join(top_var_prop %>%
              filter(form == "Continuous W x X") %>%
              select(model, measure, interim_prop, prop_sig) %>%
              pivot_wider(names_from = measure, values_from = prop_sig,
                          names_prefix = "topvar_"),
            by = c("model", "interim_prop")) %>%
  arrange(model, interim_prop)

kable(headline, digits = 3,
      caption = paste("Landmark summary. sig_* = proportion of runs where that",
                      "subgroup validates; rho_* = mean rank correlation between",
                      "chunks; topvar_* = proportion where the nominated top",
                      "covariate still interacts (continuous form)."))
```
