---
title: "Continuous outcome study: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## How well do these estimators recover heterogeneous treatment effects?

Ten scenarios vary the structure of the CATE, crossed with four sample sizes
(100, 250, 500, 1000), 100 runs each. Every dataset also carries five
deliberately unrelated covariates (`X01`–`X05`), so the estimators have to
find the signal rather than being handed it. Full design in
`continuous/README.md`.

| Scenario | CATE structure |
|---|---|
| 1: No HTE | $b_0 + b_1 X_1 + b_2 X_2 + W b_W$ |
| 2: Binary HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3)$ |
| 3: Continuous HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 X_4)$ |
| 4: Two HTE vars | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4)$ |
| 5: Cts x binary interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{34} X_3 X_4)$ |
| 6: Two vars + cts x binary | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{34} X_3 X_4)$ |
| 7: Cts x cts interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{45} X_4 X_5)$ |
| 8: Two vars + cts x cts | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$ |
| 9: Cosine HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_4 \cos(X_4))$ |
| 10: Exponential HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 \exp(-\lvert X_4 \rvert))$ |

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle`, `dr_superlearner` — all using double crossfitting for the
nuisances (`R/cate_models.R`). The oracle uses the true outcome model and a
known propensity of 0.5; the semi-oracle knows only the propensity. The
SuperLearner library shrinks at `n = 100` (`SL.earth`/`SL.ranger` are
dropped), so `dr_superlearner` is sometimes filtered out of the smallest-`n`
figures below.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "continuous")

metrics <- readRDS(file.path(res_path, "cts_metrics.RDS"))

# display order and labels - continuous has no shared label-dictionary script
# (unlike crossfitting/cf_metrics.R, which does double duty as pipeline +
# labels), so these stay inline, matching cts_results.Rmd's precedent
scenario_labels <- c(
  `1` = "1: No HTE",
  `2` = "2: Binary HTE",
  `3` = "3: Continuous HTE",
  `4` = "4: Two HTE vars",
  `5` = "5: Cts x binary interaction",
  `6` = "6: Two vars + cts x binary",
  `7` = "7: Cts x cts interaction",
  `8` = "8: Two vars + cts x cts",
  `9` = "9: Cosine HTE",
  `10` = "10: Exponential HTE"
)

model_labels <- c(
  causal_forest = "Causal forest",
  dr_random_forest = "DR-RandomForest",
  dr_oracle = "DR-oracle",
  dr_semi_oracle = "DR-semi-oracle",
  dr_superlearner = "DR-SuperLearner"
)

# tidy up
metrics <- metrics %>%
  mutate(
    scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                      labels = unname(scenario_labels)),
    n = factor(n, levels = c(100, 250, 500, 1000)),
    model = factor(recode(model, !!!model_labels), levels = unname(model_labels))
  ) %>%
  droplevels()

# per (scenario, model, n) summaries
metrics_summary <- metrics %>%
  group_by(scenario, model, n) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(n()),
    mean_BLP = mean(BLP_p, na.rm = T),
    mcse_BLP = sd(BLP_p, na.rm = T) / sqrt(n()),
    mean_indep_cate = mean(indep_cate, na.rm = T),
    mcse_indep_cate = sd(indep_cate, na.rm = T) / sqrt(n()),
    mean_indep_po = mean(indep_po, na.rm = T),
    mcse_indep_po = sd(indep_po, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with MCSE error bars, faceted by
# scenario - see crossfitting/cf_results.qmd's summary_plot for the pattern
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = 0) {
  df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]]) %>%
    ggplot(aes(x = n, y = est, colour = model, ymin = lo, ymax = hi)) +
    geom_hline(yintercept = hline, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_wrap(~scenario, scales = "free") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Sample size", colour = "Model")
}
```

## Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per
run, against the known truth.

```{r bias-plot, fig.height=16}
bias_plot <- metrics %>%
  ggplot(aes(x = n, y = bias, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bias in CATE estimates", y = "Bias", x = "Sample size", colour = "Model")

bias_sum_plot <- summary_plot(metrics_summary, "mean_bias", "mcse_bias",
                              "Mean bias in CATE estimates", "Mean bias")

(bias_plot / bias_sum_plot) + plot_layout(guides = "collect")
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth, for every model and sample size.

```{r mse-plot, fig.height=16}
mse_plot <- metrics %>%
  ggplot(aes(x = n, y = mse, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "MSE of CATE estimates", y = "MSE", x = "Sample size", colour = "Model")

mse_sum_plot <- summary_plot(metrics_summary, "mean_mse", "mcse_mse",
                             "Mean MSE of CATE estimates", "Mean MSE")

(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in scenario 1 since there's no heterogeneity to correlate with —
excluded from these plots).

```{r corr-plot, fig.height=16}
corr_df <- filter(metrics, scenario != scenario_labels[["1"]])

corr_plot <- corr_df %>%
  ggplot(aes(x = n, y = corr, colour = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  # free_y only: sample size is shared across panels, only the correlation
  # scale needs freeing per scenario
  facet_wrap(~scenario, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation between estimated and true CATEs",
       y = "Pearson correlation", x = "Sample size", colour = "Model")

corr_sum_plot <- summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
                              "mean_corr", "mcse_corr",
                              "Mean correlation with the true CATE", "Mean correlation")

(corr_plot / corr_sum_plot) + plot_layout(guides = "collect")
```

## BLP test p-values

Best linear predictor test of heterogeneity, against the nominal 5%
significance threshold (dashed line) — where the CATE truly varies, low
p-values indicate the model detects it; in scenario 1 (no HTE), p-values
should sit uniform on [0, 1].

```{r blp-plot, fig.height=16}
BLP_plot <- metrics %>%
  ggplot(aes(x = n, y = BLP_p, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "BLP test p-values", y = "p-value", x = "Sample size", colour = "Model")

BLP_sum_plot <- summary_plot(metrics_summary, "mean_BLP", "mcse_BLP",
                             "Mean BLP test p-value", "Mean p-value", hline = 0.05)

(BLP_plot / BLP_sum_plot) + plot_layout(guides = "collect")
```

## CATE permutation test p-values

```{r indep-cate-plot, fig.height=16}
indep_cate_plot <- metrics %>%
  ggplot(aes(x = n, y = indep_cate, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "CATE permutation test p-values", y = "p-value", x = "Sample size",
       colour = "Model")

indep_cate_sum_plot <- summary_plot(metrics_summary, "mean_indep_cate", "mcse_indep_cate",
                                    "Mean CATE permutation test p-value", "Mean p-value",
                                    hline = 0.05)

(indep_cate_plot / indep_cate_sum_plot) + plot_layout(guides = "collect")
```

## PO permutation test p-values

```{r indep-po-plot, fig.height=16}
indep_po_plot <- metrics %>%
  ggplot(aes(x = n, y = indep_po, colour = model)) +
  geom_hline(yintercept = 0.05, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "PO permutation test p-values", y = "p-value", x = "Sample size",
       colour = "Model")

indep_po_sum_plot <- summary_plot(metrics_summary, "mean_indep_po", "mcse_indep_po",
                                  "Mean PO permutation test p-value", "Mean p-value",
                                  hline = 0.05)

(indep_po_plot / indep_po_sum_plot) + plot_layout(guides = "collect")
```

## Headline table

Arranged by scenario, sample size, and mean MSE (best model first within
each cell).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, model, n, mean_bias, mean_mse, mean_corr, mean_BLP,
         mean_indep_cate, mean_indep_po) %>%
  arrange(scenario, n, mean_mse)

kable(headline, digits = 4,
      col.names = c("Scenario", "Model", "n", "Mean bias", "Mean MSE",
                    "Mean correlation", "Mean BLP p-value",
                    "Mean CATE perm. p-value", "Mean PO perm. p-value"))
```
