---
title: "Missing covariates, binary outcome: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
```

## What does missingness cost, and which handling method costs least?

Covariates are amputated from a dataset the estimators would otherwise have seen
in full, then handed to nine different handling methods. Every arm is scored
against the *same* known true CATE: τ at the **unamputed** covariates.

**Complete and incomplete units are scored separately.** No method can recover
τ for a unit whose effect modifiers are missing, so on those units every
imputation method carries an error floor, while complete cases and IPW are
scored on their complete units alone. Scored on all analysed units, complete
cases would win on MSE whatever they did. So every point metric also comes on
the **complete units** (suffix `_cu`), the subset every method has and the one
the methods are compared on, and on the **incomplete units** (`_iu`), which
show the floor. The split is the amputation's, so the complete units are a
different set under each mechanism: compare mechanisms through the ratios to
`complete_data`, not through absolute `_cu` / `_iu` values.

Unlike the sample-size studies in `continuous/` and `binary/`, this design holds
the sample size fixed and varies the missingness instead. There is therefore no
sample-size axis in any figure below:

| | |
|---|---|
| scenarios | 1–5 — the main study's scenarios 1–5, same numbers |
| n | 500 |
| type | `both` (prognostic and predictive covariates amputated) |
| prop | 0.3 |
| mechanism | MAR, MNAR-Y0, MNAR-τ |
| method | 8 handling methods + `complete_data` reference |
| runs | 100 |

**Scenario numbering is the main study's.** Scenario *k* here is scenario *k*
in `binary/`; this study runs 1–5. Full design in `missing/README.md`.

The outcome is binary, so every effect below is a **risk difference**, and the
truth is on the probability scale. The DGM puts the treatment effect on that
scale directly — control risk bounded in [0.34, 0.70], `P(Y = 1) = m0(x) + W·τ(x)`
— so X1 and X2 are purely prognostic (see `sample_size/binary/README.md`). The oracle
formula from `get_binary_oracle_info()` returns the risk itself.

### Mechanisms

| mechanism | missingness depends on |
|---|---|
| MAR | the observed covariates (correlated with the missing ones) |
| MNAR-Y0 | an unobserved `U`, which also shifts the control risk |
| MNAR-τ | an unobserved `U`, which also enters the treatment effect |

MNAR-τ is not defined for scenario 1 (Null): there is no treatment effect
heterogeneity for `U` to enter. **That facet is empty by design, not by
failure.**

### Handling methods

| method | what it does |
|---|---|
| Complete case | delete every row with any `NA` |
| Single mean imputation | replace `NA` with the column mean |
| Single forest-based imputation | `missForest`, within each arm, Y as a predictor |
| Single model-based imputation | predict the missing values by regression |
| Missing indicators | add binary missing flags, then mean-impute |
| IPW | inverse-probability weighting of the complete cases |
| Multiple imputation (rf) | MICE, 50 imputations within each arm, Y as a predictor, Rubin-combined |
| Inbuilt missingness handling | hand the `NA`s straight to the estimator |
| Complete data (reference) | no missingness introduced at all |

`complete_cases` and `IPW` drop rows from the truth as well as the data, so
estimates and truth stay aligned.

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle`, `dr_superlearner` (`R/cate_models.R`, `profile = "missing"`).
The oracle uses the true outcome model and a known propensity of 0.5; the
semi-oracle knows only the propensity. The SuperLearner arm is skipped whenever
the covariate matrix still carries `NA`s.

Every model carries the BLP and independence tests. The `multiple_imputation`
arm's are saved per imputation but not yet pooled, so it is `NA` in the three
test sections below until a pooling rule is chosen (see `missing/README.md`).

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "missing", "binary")

metrics <- readRDS(file.path(res_path, "bin_miss_metrics.RDS"))

# n / type / prop are single-valued in this design, so nothing facets on them -
# but they stay in the grouping so this survives a grid that gains a level
metrics <- metrics %>%
  apply_labels(MISS_SCENARIO_LABELS) %>%
  mutate(n = factor(n), prop = factor(prop)) %>%
  droplevels()

# derived 0/1 reject-indicators for the power summary - summarise_metrics()'s
# binomial argument expects a real per-run column, not an inline expression
metrics <- metrics %>%
  mutate(BLP_reject = as.numeric(BLP_p < 0.05),
         BLP_os_reject = as.numeric(BLP_p_os < 0.05),
         indep_cate_reject = as.numeric(indep_cate < 0.05),
         indep_po_reject = as.numeric(indep_po < 0.05))

metrics_summary <- summarise_metrics(
  metrics,
  c("scenario", "n", "type", "prop", "mechanism", "method", "model"),
  cols = c(bias = "bias", ate_bias = "ate_bias", mse = "mse", rmse = "rmse",
           mae = "mae", corr = "corr", spearman = "spearman",
           sign_acc = "sign_acc", BLP = "BLP_p", BLP_os = "BLP_p_os",
           indep_cate = "indep_cate",
           indep_po = "indep_po", rel_eff = "rel_efficiency",
           rel_eff_cu = "rel_efficiency_cu", rel_eff_iu = "rel_efficiency_iu",
           bias_diff_cu = "bias_diff_complete_cu",
           bias_cu = "bias_cu", bias_iu = "bias_iu",
           rmse_cu = "rmse_cu", rmse_iu = "rmse_iu", corr_cu = "corr_cu",
           corr_iu = "corr_iu", rmse_r = "rmse_r", rmse_r_cu = "rmse_r_cu",
           rmse_r_iu = "rmse_r_iu", ate_bias_r = "ate_bias_r",
           rel_ate_bias = "rel_ate_bias",
           power_BLP = "BLP_reject", power_BLP_os = "BLP_os_reject",
           power_indep_cate = "indep_cate_reject",
           power_indep_po = "indep_po_reject"),
  binomial = c("power_BLP", "power_BLP_os", "power_indep_cate", "power_indep_po")
)
```

```{r plot-helpers, include=FALSE}
# every summary panel's legend: along the bottom, wrapped over three rows so
# the nine method labels fit the figure width
legend_bottom <- list(
  theme(legend.position = "bottom"),
  guides(colour = guide_legend(nrow = 3)),
  labs(colour = "Method")
)
```

Every panel below shows the mean across runs with its 95% Monte Carlo interval
(mean ± 1.96 × MCSE). Rows are the missingness mechanism, columns the scenario,
colour the handling method.

## Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per run,
against the known truth.

```{r bias-plot, fig.height=9}
point_range_plot(metrics_summary, "bias", "Bias") + legend_bottom
```

## ATE bias

Bias of the *average* treatment effect (`mean(est) - mean(true)`), as opposed to
Bias's per-unit average — the two can differ in sign or magnitude when errors
don't average out evenly across units.

```{r ate-bias-plot, fig.height=9}
point_range_plot(metrics_summary, "ate_bias", "ATE bias") + legend_bottom
```

## MSE of CATE estimates

Mean squared error of the estimated CATE against the known truth, on all
analysed units — about 350 for complete cases and IPW, 500 for the rest, so
not comparable across the two. The comparison across methods is the
complete-units section below.

```{r mse-plot, fig.height=9}
point_range_plot(metrics_summary, "mse", "MSE", facet_scales = "free_y") +
  legend_bottom
```

## RMSE of CATE estimates

`sqrt(mse)` — the same ranking as MSE, shown on the outcome's own scale rather
than squared.

```{r rmse-plot, fig.height=9}
point_range_plot(metrics_summary, "rmse", "RMSE", facet_scales = "free_y") +
  legend_bottom
```

## MAE of CATE estimates

Mean absolute error — a robustness check on MSE, less sensitive to outlying runs.

```{r mae-plot, fig.height=9}
point_range_plot(metrics_summary, "mae", "MAE", facet_scales = "free_y") +
  legend_bottom
```

## Complete and incomplete units

RMSE and bias on the complete units (top: the subset every method has) and the
incomplete units (bottom: complete cases and IPW have none, so draw nothing).
The incomplete-unit error of the imputation methods is mostly the floor that
comes from scoring against τ at covariates they never saw.

```{r rmse-split-plot, fig.height=17}
rmse_cu_sum_plot <- point_range_plot(metrics_summary, "rmse_cu",
                                     "RMSE, complete units", facet_scales = "free_y")
rmse_iu_sum_plot <- point_range_plot(metrics_summary, "rmse_iu",
                                     "RMSE, incomplete units", facet_scales = "free_y")

# the collected legend's position comes from the patchwork's own theme, which
# only `& theme()` sets - legend_bottom on each panel isn't enough on its own
((rmse_cu_sum_plot + legend_bottom) / (rmse_iu_sum_plot + legend_bottom)) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r bias-split-plot, fig.height=17}
bias_cu_sum_plot <- point_range_plot(metrics_summary, "bias_cu", "Bias, complete units")
bias_iu_sum_plot <- point_range_plot(metrics_summary, "bias_iu", "Bias, incomplete units")

((bias_cu_sum_plot + legend_bottom) / (bias_iu_sum_plot + legend_bottom)) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

## Relative efficiency

MSE on the complete units divided by the complete-unit MSE of the same
scenario, mechanism, run and model under `complete_data`
(`rel_efficiency_cu`). The reference arm sits at exactly 1 by construction
(dashed line); anything above 1 is the price of the missingness plus the
handling method. It is a within-mechanism ratio, so it is also how the
mechanisms compare: MNAR-Y0's `U` adds outcome variance to the `complete_data`
arm as well. The all-unit ratio is not defined for complete cases and IPW,
which analyse fewer units than `complete_data`.

```{r rel-eff-plot, fig.height=9}
point_range_plot(metrics_summary, "rel_eff_cu",
                 "Relative efficiency, complete units",
                 facet_scales = "free_y", hline = 1) +
  legend_bottom
```

## Bias difference (vs complete data)

Complete-unit bias minus the complete-unit bias of the same scenario,
mechanism, run and model under `complete_data`: the bias the missingness and
the handling method add. A difference rather than a ratio — the complete-data
bias is often near zero, which made the old ratio `rel_bias_complete`
unstable.

```{r bias-diff-plot, fig.height=9}
point_range_plot(metrics_summary, "bias_diff_cu",
                 "Bias minus complete-data bias, complete units",
                 facet_scales = "free_y") +
  legend_bottom
```

## Relative ATE bias (vs true)

`(mean(est) - mean(true)) / mean(true)`: the ATE's bias (`ate_bias`) as a
fraction of the true ATE — the standard parameter-level "relative bias" from
the simulation literature. Unlike the panel above, this is a *bias fraction*,
not an arm-to-arm ratio, so the reference line is 0 (dashed), the same
convention as the plain Bias/ATE bias panels.

```{r rel-ate-bias-plot, fig.height=9}
if ("rel_ate_bias" %in% names(metrics)) {
  point_range_plot(metrics_summary, "rel_ate_bias",
                   "Relative ATE bias (vs true)", facet_scales = "free_y") +
    legend_bottom
} else {
  cat("rel_ate_bias is not present in this metrics file.")
}
```

The per-unit relative bias `mean((est - true) / true)` (`rel_bias_cate`) is
still in the metrics file but not shown: the true CATE crosses zero in
scenarios 2–5, so the ratio is dominated by the units nearest zero.

## MNAR-τ: the truth given completeness

Under MNAR-τ, `U` drives the missingness and sits in the treatment effect, so
the incomplete units' true effects are shifted away from τ(X). The primary
truth ignores that — it is the CATE as a function of X under every mechanism.
This panel scores against the secondary truth τ + E[U-term | complete or
incomplete] instead, which rewards a method that learns the shift from the
missingness itself (missing indicators, inbuilt handling). See
`missing/ADEMP.md`, "Estimands".

Rows are all analysed units, then the complete and incomplete units
separately. Complete cases and IPW have no incomplete units, so they are
absent from the bottom row.

```{r rmse-r-plot, fig.height=11}
mnar_tau_summary <- filter(metrics_summary,
                           mechanism == MECHANISM_LABELS[["MNAR-tau"]])
r_stems <- c(`All analysed units` = "rmse_r", `Complete units` = "rmse_r_cu",
             `Incomplete units` = "rmse_r_iu")
if (nrow(mnar_tau_summary) > 0 &&
    all(paste0("mean_", r_stems) %in% names(mnar_tau_summary))) {
  # stack the three subsets under one mean/mcse pair, so point_range_plot()
  # can put one row of facets per subset
  mnar_tau_long <- imap(r_stems, function(stem, units) {
    transmute(mnar_tau_summary, scenario, method, model, units = units,
              mean_rmse_r = .data[[paste0("mean_", stem)]],
              mcse_rmse_r = .data[[paste0("mcse_", stem)]])
  }) %>%
    bind_rows() %>%
    mutate(units = factor(units, levels = names(r_stems)))
  point_range_plot(mnar_tau_long, "rmse_r",
                   "RMSE against the truth given completeness",
                   facet_rows = "units", facet_scales = "free_y") +
    legend_bottom
} else {
  cat("No MNAR-tau rows with the secondary truth in this metrics file.")
}
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs. Undefined, and reported as
0, in the Null scenario since there is no heterogeneity to correlate with — so
that scenario is excluded here.

```{r corr-plot, fig.height=9}
corr_summary <- filter(metrics_summary, scenario != "Null")

point_range_plot(corr_summary, "corr", "Pearson correlation",
                 facet_scales = "free_y", hline = NULL) +
  legend_bottom
```

## Spearman rank correlation with the true CATE

Rank correlation alongside the Pearson correlation above — less sensitive to
outlying CATE estimates and to non-linear-but-monotone miscalibration. Same
Null-undefined convention, so that scenario is excluded here too.

```{r spearman-plot, fig.height=9}
point_range_plot(corr_summary, "spearman", "Spearman correlation",
                 facet_scales = "free_y", hline = NULL) +
  legend_bottom
```

## Sign accuracy

Proportion of units where the estimated CATE has the correct sign — a
direction-detection metric, distinct from magnitude accuracy. 0.5 (dashed line)
is the chance level for a coin-flip on sign.

```{r sign-acc-plot, fig.height=9}
point_range_plot(metrics_summary, "sign_acc", "Proportion correct sign",
                 facet_scales = "free_x", hline = 0.5) +
  legend_bottom
```

## True-CATE HTE test evaluation

The BLP and independence tests re-run on the true CATE and true nuisances
instead of an estimator's fitted ones (`bin_miss_true_cate_tests.RDS` —
`run_true_cate_tests()`/`true_cate_test_row()`, `R/cate_models.R`) — the
ceiling each model's own BLP/CATE-permutation test is chasing, since nothing
here is estimated. Shown as an extra "True CATE" x-axis category alongside
the five models in the two sections below (and in "HTE test power"); no
true-CATE counterpart for the PO permutation test, since `dr_oracle`'s own
`indep_po` already tests the true pseudo-outcome. `multiple_imputation` rows
are `NA` here, same as everywhere else in this study (`data` is a list of 50
imputed data frames there, with no single covariate matrix to test against).

BLP never touches the covariates (`run_blp_whole()` takes `Y, W, Y0.hat,
W.hat, tau` only), so its true-CATE series is identical across `method` for
the same run — expected, not a bug. `indep_cate` *does* depend on the
covariates, which genuinely differ by handling method (imputed values,
dropped rows, extra `_missing` indicator columns), so its true-CATE series
varies by method for real: it shows whether a method's reshaping of the
covariates alone (CATE held fixed at the truth) helps or hurts the
independence test's power.

```{r true-cate-data}
true_cate_tests <- readRDS(file.path(res_path, "bin_miss_true_cate_tests.RDS")) %>%
  apply_labels(MISS_SCENARIO_LABELS) %>%
  mutate(n = factor(n), prop = factor(prop)) %>%
  droplevels()

true_cate_summary <- true_cate_tests %>%
  group_by(scenario, n, type, prop, mechanism, method) %>%
  summarise(
    mean_BLP = mean(BLP_p, na.rm = TRUE),
    mcse_BLP = sd(BLP_p, na.rm = TRUE) / sqrt(sum(!is.na(BLP_p))),
    mean_BLP_os = mean(BLP_p_os, na.rm = TRUE),
    mcse_BLP_os = sd(BLP_p_os, na.rm = TRUE) / sqrt(sum(!is.na(BLP_p_os))),
    mean_indep_cate = mean(indep_cate, na.rm = TRUE),
    mcse_indep_cate = sd(indep_cate, na.rm = TRUE) / sqrt(sum(!is.na(indep_cate))),
    # named mean_power_*/mcse_power_*, not power_*/mcse_power_* - matching
    # summarise_metrics()'s own naming convention below, which always prefixes
    # mean_/mcse_ even for its binomial columns
    mean_power_BLP = mean(BLP_p < 0.05, na.rm = TRUE),
    mcse_power_BLP = sqrt(mean_power_BLP * (1 - mean_power_BLP) / sum(!is.na(BLP_p))),
    mean_power_BLP_os = mean(BLP_p_os < 0.05, na.rm = TRUE),
    mcse_power_BLP_os = sqrt(mean_power_BLP_os * (1 - mean_power_BLP_os) /
                               sum(!is.na(BLP_p_os))),
    mean_power_indep_cate = mean(indep_cate < 0.05, na.rm = TRUE),
    mcse_power_indep_cate = sqrt(mean_power_indep_cate * (1 - mean_power_indep_cate) /
                                   sum(!is.na(indep_cate))),
    .groups = "drop"
  ) %>%
  mutate(model = "True CATE")

model_levels_true <- c(levels(metrics$model), "True CATE")

# row-bind the true-CATE reference series onto a summarised data frame,
# extending `model`'s levels so it becomes one more x-axis category
bind_true_cate <- function(df, true_df, extra_cols) {
  bind_rows(
    mutate(df, model = factor(as.character(model), levels = model_levels_true)),
    true_df %>%
      mutate(model = factor(model, levels = model_levels_true)) %>%
      select(scenario, n, type, prop, mechanism, method, model, all_of(extra_cols))
  )
}
```

## BLP test p-values

Best linear predictor test of heterogeneity, against the nominal 5% significance
threshold (dashed line) — where the CATE truly varies, low p-values indicate the
model detects it; in the Null scenario, p-values should sit uniform on [0, 1].
Two versions of the p-value for the BLP's beta.2 (see
`sample_size/continuous/README.md`, "HTE tests"): `BLP_p` is two-sided with
constant-variance SEs, as computed when the models were fitted; `BLP_p_os` is
one-sided (H1: beta.2 > 0) with HC3 SEs, recomputed at metrics time from the
saved nuisances.

```{r blp-plot, fig.height=9}
point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_BLP", "mcse_BLP")),
  "BLP", "p-value", facet_scales = "free_x", hline = 0.05) +
  labs(title = "BLP test (two-sided, constant variance)") +
  legend_bottom
```

```{r blp-os-plot, fig.height=9}
point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_BLP_os", "mcse_BLP_os")),
  "BLP_os", "p-value", facet_scales = "free_x", hline = 0.05) +
  labs(title = "BLP test (one-sided, HC3)") +
  legend_bottom
```

## CATE permutation test p-values

Omnibus independence test of the estimated CATEs against the covariates.

```{r indep-cate-plot, fig.height=9}
point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_indep_cate", "mcse_indep_cate")),
  "indep_cate", "p-value", facet_scales = "free_x", hline = 0.05) +
  legend_bottom
```

## PO permutation test p-values

The same test applied to the potential-outcome pseudo-outcomes rather than the
fitted CATEs. The causal forest has no series here: its pseudo-outcome is
DR-RandomForest's (both use the same random-forest nuisances), so the test is
reported once.

```{r indep-po-plot, fig.height=9}
point_range_plot(metrics_summary, "indep_po", "p-value",
                 facet_scales = "free_x", hline = 0.05) +
  legend_bottom
```

## HTE test power

Power (proportion of runs rejecting at alpha=0.05) alongside the mean p-values
above — the rsimsum-standard measure for a hypothesis test's simulated
behaviour, with the binomial 95% CI rather than the continuous-measure one.
The true-CATE reference series applies to BLP and the CATE permutation test
only, same as the p-value panels above; PO has no true-CATE counterpart
anywhere in this evaluation (`dr_oracle`'s own `indep_po` already covers it).

```{r power-plot, fig.height=9}
BLP_power_plot <- point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary, c("mean_power_BLP", "mcse_power_BLP")),
  "power_BLP", "Power", facet_scales = "free_x", hline = 0.05) +
  labs(title = "BLP test (two-sided, constant variance)") +
  legend_bottom

BLP_os_power_plot <- point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary,
                c("mean_power_BLP_os", "mcse_power_BLP_os")),
  "power_BLP_os", "Power", facet_scales = "free_x", hline = 0.05) +
  labs(title = "BLP test (one-sided, HC3)") +
  legend_bottom

indep_cate_power_plot <- point_range_plot(
  bind_true_cate(metrics_summary, true_cate_summary,
                c("mean_power_indep_cate", "mcse_power_indep_cate")),
  "power_indep_cate", "Power", facet_scales = "free_x", hline = 0.05) +
  legend_bottom

indep_po_power_plot <- point_range_plot(metrics_summary, "power_indep_po", "Power",
                                        facet_scales = "free_x", hline = 0.05) +
  legend_bottom

BLP_power_plot
BLP_os_power_plot
indep_cate_power_plot
indep_po_power_plot
```

Runs with an `NA` test p-value are left out of that test's power and mean
p-value: a degenerate (constant) CATE estimate, for which neither test is
defined, or a test that wasn't run (every `multiple_imputation` row, until a
pooling rule is chosen). Cells with any are listed below.

```{r test-na-table}
test_na_table <- metrics_summary %>%
  select(scenario, mechanism, method, model, n_na_BLP_p, n_na_BLP_p_os, n_na_indep_cate) %>%
  filter(n_na_BLP_p > 0 | n_na_BLP_p_os > 0 | n_na_indep_cate > 0) %>%
  arrange(scenario, mechanism, method, model)

if (nrow(test_na_table) > 0) {
  kable(test_na_table,
        col.names = c("Scenario", "Mechanism", "Method", "Model", "NA BLP p",
                      "NA BLP p (one-sided, HC3)", "NA CATE perm. p"))
} else {
  cat("No NA HTE test p-values in any run.")
}
```

## Missing estimates

`n_na` is a data-quality diagnostic, not a performance metric: the count of `NA`
CATE estimates within a run. Mostly zero, so shown as a table rather than a
distribution plot; only cells with at least one missing estimate are listed.

```{r na-table}
na_table <- metrics %>%
  group_by(scenario, mechanism, method, model) %>%
  summarise(mean_n_na = mean(n_na, na.rm = TRUE),
            total_n_na = sum(n_na, na.rm = TRUE),
            .groups = "drop") %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, mechanism, method, model)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("Scenario", "Mechanism", "Method", "Model",
                      "Mean NA count per run", "Total NA count"))
} else {
  cat("No missing CATE estimates in any run.")
}
```

## Headline table

Arranged by scenario, mechanism, method, and mean MSE (best model first within
each cell).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, mechanism, method, model, mean_bias, mean_rmse,
         mean_rmse_cu, mean_rmse_iu, mean_rel_eff_cu, mean_bias_diff_cu,
         mean_mae, mean_sign_acc, mean_corr, any_of("mean_rmse_r"),
         mean_BLP, mean_BLP_os, mean_indep_cate, mean_indep_po) %>%
  arrange(scenario, mechanism, method, mean_rmse_cu)

kable(headline, digits = 4)
```
