---
title: "Missing covariates, binary outcome: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
```

## What does missingness cost, and which handling method costs least?

Covariates are amputated from a dataset the estimators would otherwise have seen
in full, then handed to nine different handling methods. Every arm is scored
against the *same* known true CATE, so the comparison is like-for-like.

Unlike the sample-size studies in `continuous/` and `binary/`, this design holds
the sample size fixed and varies the missingness instead. There is therefore no
sample-size axis in any figure below:

| | |
|---|---|
| scenarios | 1, 2, 4, 5 — a **renumbered** set, not a subset of the main ten |
| n | 500 |
| type | `both` (prognostic and predictive covariates amputated) |
| prop | 0.3 |
| mechanism | MAR, MNAR, MNAR-Y |
| method | 8 handling methods + `complete_data` reference |
| runs | 100 |

**Scenario numbering here is not the main study's.** Scenario *k* corresponds to
main-study scenario 1, 2, 4, 8, 9 respectively — so "Complex" here means
something different from scenario 4 in `binary/`. Full design in
`missing/README.md`.

The outcome is binary, so every effect below is a **risk difference**, and the
truth is on the probability scale. `get_binary_oracle_info()` wraps its formula
in `plogis()`, which is why `bin_miss_models.R` passes
`oracle_link = "identity"` — the opposite convention to `binary/`.

### Mechanisms

| mechanism | missingness depends on |
|---|---|
| MAR | the observed covariates |
| MNAR | an unobserved `U` |
| MNAR-Y | `U`, which also enters the treatment effect |

MNAR-Y is not defined for scenario 1 (Null): there is no treatment effect
heterogeneity for `U` to enter. **That facet is empty by design, not by
failure.**

### Handling methods

| method | what it does |
|---|---|
| Complete case | delete every row with any `NA` |
| Single mean imputation | replace `NA` with the column mean |
| Single forest-based imputation | `missForest` |
| Single model-based imputation | predict the missing values by regression |
| Missing indicators | add binary missing flags, then mean-impute |
| IPW | inverse-probability weighting of the complete cases |
| Multiple imputation (rf) | MICE, 50 imputations, Rubin-combined |
| Inbuilt missingness handling | hand the `NA`s straight to the estimator |
| Complete data (reference) | no missingness introduced at all |

`complete_cases` and `IPW` drop rows from the truth as well as the data, so
estimates and truth stay aligned.

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle`, `dr_superlearner` (`R/cate_models.R`, `profile = "missing"`).
The oracle uses the true outcome model and a known propensity of 0.5; the
semi-oracle knows only the propensity. The SuperLearner arm is skipped whenever
the covariate matrix still carries `NA`s.

`dr_random_forest` used to carry no BLP or independence test under
`profile = "missing"`, so its p-values were `NA` in the three test sections
below. That is fixed: the flag is now `dr_rf_tests = TRUE`, and the finished
results were back-filled in place rather than re-run (see
`missing/binary/README.md`). If this model is still `NA` below, the patch job
has not been run over these results yet.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "missing", "binary")

metrics <- readRDS(file.path(res_path, "bin_miss_metrics.RDS"))

# n / type / prop are single-valued in this design, so nothing facets on them -
# but they stay in the grouping so this survives a grid that gains a level
metrics <- metrics %>%
  apply_labels(MISS_SCENARIO_LABELS) %>%
  mutate(n = factor(n), prop = factor(prop)) %>%
  droplevels()

metrics_summary <- summarise_metrics(
  metrics,
  c("scenario", "n", "type", "prop", "mechanism", "method", "model"),
  cols = c(bias = "bias", ate_bias = "ate_bias", mse = "mse", rmse = "rmse",
           mae = "mae", corr = "corr", spearman = "spearman",
           sign_acc = "sign_acc", BLP = "BLP_p", indep_cate = "indep_cate",
           indep_po = "indep_po", rel_eff = "rel_efficiency",
           rel_bias = "rel_bias")
)
```

```{r box-plot-helper, include=FALSE}
# the per-run distribution behind each summary panel. distribution_plot() in
# R/figures.R is the sample-size studies' shape (x = n, facet_wrap(~scenario));
# this study has n fixed and two extra factors, so it needs its own. Faceting
# matches point_range_plot() so the two stack cleanly.
miss_box_plot <- function(metrics, metric, y_lab, hline = 0,
                          facet_scales = "free") {
  p <- ggplot(metrics, aes(x = model, y = .data[[metric]], colour = method)) +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(rows = vars(mechanism), cols = vars(scenario),
               scales = facet_scales) +
    drf_scale() +
    drf_theme(rotate_x = TRUE) +
    labs(y = y_lab, x = "Model", colour = "Method")
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p
}
```

Every section below pairs the per-run distribution (top) with the mean ± Monte
Carlo standard error across runs (bottom). Rows are the missingness mechanism,
columns the scenario, colour the handling method.

## Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per run,
against the known truth.

```{r bias-plot, fig.height=16}
bias_plot <- miss_box_plot(metrics, "bias", "Bias")
bias_sum_plot <- point_range_plot(metrics_summary, "bias", "Bias")

(bias_plot / bias_sum_plot) + plot_layout(guides = "collect")
```

## ATE bias

Bias of the *average* treatment effect (`mean(est) - mean(true)`), as opposed to
Bias's per-unit average — the two can differ in sign or magnitude when errors
don't average out evenly across units.

```{r ate-bias-plot, fig.height=16}
ate_bias_plot <- miss_box_plot(metrics, "ate_bias", "ATE bias")
ate_bias_sum_plot <- point_range_plot(metrics_summary, "ate_bias", "ATE bias")

(ate_bias_plot / ate_bias_sum_plot) + plot_layout(guides = "collect")
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the known
truth.

```{r mse-plot, fig.height=16}
mse_plot <- miss_box_plot(metrics, "mse", "MSE", facet_scales = "free_y")
mse_sum_plot <- point_range_plot(metrics_summary, "mse", "MSE",
                                 facet_scales = "free_y")

(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
```

## RMSE of CATE estimates

`sqrt(mse)` — the same ranking as MSE, shown on the outcome's own scale rather
than squared.

```{r rmse-plot, fig.height=16}
rmse_plot <- miss_box_plot(metrics, "rmse", "RMSE", facet_scales = "free_y")
rmse_sum_plot <- point_range_plot(metrics_summary, "rmse", "RMSE",
                                  facet_scales = "free_y")

(rmse_plot / rmse_sum_plot) + plot_layout(guides = "collect")
```

## MAE of CATE estimates

Mean absolute error — a robustness check on MSE, less sensitive to outlying runs.

```{r mae-plot, fig.height=16}
mae_plot <- miss_box_plot(metrics, "mae", "MAE", facet_scales = "free_y")
mae_sum_plot <- point_range_plot(metrics_summary, "mae", "MAE",
                                 facet_scales = "free_y")

(mae_plot / mae_sum_plot) + plot_layout(guides = "collect")
```

## Relative efficiency

`mse` divided by the MSE of the same scenario, mechanism, run and model under
`complete_data`. The reference arm sits at exactly 1 by construction (dashed
line); anything above 1 is the price of the missingness plus the handling
method. This is the metric specific to the missing-data studies — nothing in
`binary/` or `continuous/` has it.

```{r rel-eff-plot, fig.height=16}
if ("rel_efficiency" %in% names(metrics)) {
  rel_eff_plot <- miss_box_plot(metrics, "rel_efficiency",
                                "Relative efficiency", hline = 1,
                                facet_scales = "free_y")
  rel_eff_sum_plot <- point_range_plot(metrics_summary, "rel_eff",
                                       "Relative efficiency",
                                       facet_scales = "free_y") +
    geom_hline(yintercept = 1, linetype = "dashed")

  (rel_eff_plot / rel_eff_sum_plot) + plot_layout(guides = "collect")
} else {
  cat("rel_efficiency is not present in this metrics file.")
}
```

## Relative bias

`bias` divided by the bias of the same scenario, mechanism, run and model under
`complete_data`. Unlike relative efficiency this is a *signed* ratio: 1 (dashed
line) means the same bias as complete data, a negative value means the handling
method flipped the sign of the bias, and the ratio blows up when the
complete-data bias is near zero — read it alongside the plain Bias panel above,
not on its own.

```{r rel-bias-plot, fig.height=16}
if ("rel_bias" %in% names(metrics)) {
  rel_bias_plot <- miss_box_plot(metrics, "rel_bias",
                                 "Relative bias", hline = 1,
                                 facet_scales = "free_y")
  rel_bias_sum_plot <- point_range_plot(metrics_summary, "rel_bias",
                                        "Relative bias",
                                        facet_scales = "free_y") +
    geom_hline(yintercept = 1, linetype = "dashed")

  (rel_bias_plot / rel_bias_sum_plot) + plot_layout(guides = "collect")
} else {
  cat("rel_bias is not present in this metrics file.")
}
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs. Undefined, and reported as
0, in the Null scenario since there is no heterogeneity to correlate with — so
that scenario is excluded here.

```{r corr-plot, fig.height=16}
corr_metrics <- filter(metrics, scenario != "Null")
corr_summary <- filter(metrics_summary, scenario != "Null")

corr_plot <- miss_box_plot(corr_metrics, "corr", "Pearson correlation",
                           hline = NULL, facet_scales = "free_y")
corr_sum_plot <- point_range_plot(corr_summary, "corr", "Pearson correlation",
                                  facet_scales = "free_y")

(corr_plot / corr_sum_plot) + plot_layout(guides = "collect")
```

## Spearman rank correlation with the true CATE

Rank correlation alongside the Pearson correlation above — less sensitive to
outlying CATE estimates and to non-linear-but-monotone miscalibration. Same
Null-undefined convention, so that scenario is excluded here too.

```{r spearman-plot, fig.height=16}
spearman_plot <- miss_box_plot(corr_metrics, "spearman", "Spearman correlation",
                               hline = NULL, facet_scales = "free_y")
spearman_sum_plot <- point_range_plot(corr_summary, "spearman",
                                      "Spearman correlation",
                                      facet_scales = "free_y")

(spearman_plot / spearman_sum_plot) + plot_layout(guides = "collect")
```

## Sign accuracy

Proportion of units where the estimated CATE has the correct sign — a
direction-detection metric, distinct from magnitude accuracy. 0.5 (dashed line)
is the chance level for a coin-flip on sign.

```{r sign-acc-plot, fig.height=16}
sign_acc_plot <- miss_box_plot(metrics, "sign_acc", "Proportion correct sign",
                               hline = 0.5, facet_scales = "free_x")
sign_acc_sum_plot <- point_range_plot(metrics_summary, "sign_acc",
                                      "Proportion correct sign",
                                      facet_scales = "free_x") +
  geom_hline(yintercept = 0.5, linetype = "dashed")

(sign_acc_plot / sign_acc_sum_plot) + plot_layout(guides = "collect")
```

## BLP test p-values

Best linear predictor test of heterogeneity, against the nominal 5% significance
threshold (dashed line) — where the CATE truly varies, low p-values indicate the
model detects it; in the Null scenario, p-values should sit uniform on [0, 1].

`dr_random_forest` appears in these panels only once the HTE back-fill has been
run over the results (`R/patch_hte_tests.R`); before that it carries no BLP test
under `profile = "missing"`.

```{r blp-plot, fig.height=16}
BLP_plot <- miss_box_plot(metrics, "BLP_p", "p-value", hline = 0.05,
                          facet_scales = "free_x")
BLP_sum_plot <- point_range_plot(metrics_summary, "BLP", "p-value",
                                 facet_scales = "free_x") +
  geom_hline(yintercept = 0.05, linetype = "dashed")

(BLP_plot / BLP_sum_plot) + plot_layout(guides = "collect")
```

## CATE permutation test p-values

Omnibus independence test of the estimated CATEs against the covariates.

```{r indep-cate-plot, fig.height=16}
indep_cate_plot <- miss_box_plot(metrics, "indep_cate", "p-value", hline = 0.05,
                                 facet_scales = "free_x")
indep_cate_sum_plot <- point_range_plot(metrics_summary, "indep_cate", "p-value",
                                        facet_scales = "free_x") +
  geom_hline(yintercept = 0.05, linetype = "dashed")

(indep_cate_plot / indep_cate_sum_plot) + plot_layout(guides = "collect")
```

## PO permutation test p-values

The same test applied to the potential-outcome pseudo-outcomes rather than the
fitted CATEs.

```{r indep-po-plot, fig.height=16}
indep_po_plot <- miss_box_plot(metrics, "indep_po", "p-value", hline = 0.05,
                               facet_scales = "free_x")
indep_po_sum_plot <- point_range_plot(metrics_summary, "indep_po", "p-value",
                                      facet_scales = "free_x") +
  geom_hline(yintercept = 0.05, linetype = "dashed")

(indep_po_plot / indep_po_sum_plot) + plot_layout(guides = "collect")
```

## Missing estimates

`n_na` is a data-quality diagnostic, not a performance metric: the count of `NA`
CATE estimates within a run. Mostly zero, so shown as a table rather than a
distribution plot; only cells with at least one missing estimate are listed.

```{r na-table}
na_table <- metrics %>%
  group_by(scenario, mechanism, method, model) %>%
  summarise(mean_n_na = mean(n_na, na.rm = TRUE),
            total_n_na = sum(n_na, na.rm = TRUE),
            .groups = "drop") %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, mechanism, method, model)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("Scenario", "Mechanism", "Method", "Model",
                      "Mean NA count per run", "Total NA count"))
} else {
  cat("No missing CATE estimates in any run.")
}
```

## Headline table

Arranged by scenario, mechanism, method, and mean MSE (best model first within
each cell).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, mechanism, method, model, mean_bias, mean_mse, mean_rmse,
         mean_mae, mean_sign_acc, mean_corr, any_of("mean_rel_eff"),
         mean_BLP, mean_indep_cate, mean_indep_po) %>%
  arrange(scenario, mechanism, method, mean_mse)

kable(headline, digits = 4)
```
