---
title: "Crossfitting comparison: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Does double crossfitting earn its cost?

The simulations in the current project initially set out to use double cross-fitting,
following implementations seen in Hines' TE-VIMS paper. Two-stage estimators fit
nuisances over all `C(V,2)` fold *pairs*: `collate_predictions` assembles an
`n x V` pseudo-outcome matrix whose column `k` is untouched by fold `k`, and
the second stage trains on `po_matrix[, k]` before predicting fold `k`. At
`V = 10` that is 45 nuisance fits instead of 10, a 4.5x cost ($\frac{(n)(n-1)}{2n}$).

Under ordinary (single) crossfitting, a stage-2 training row `i` carries a
pseudo-outcome from a nuisance model trained on every fold except `i`'s - 
which *includes* the held-out test fold. Re-randomising the
stage-2 split cannot remove that dependence. So if double crossfitting
matters, `scf_scf` and `scf_scf_new` should perform alike, and both should
differ from `dcf`. If they don't, the cheap procedure is enough and the rest
of the study can be sped up 4.5x.

### Design

Fixed at **n = 500, V = 10 folds, scenarios Null / Simple / Interaction /
Non-linear, 500 replicates** (2000 array jobs). All variants within a
replicate share one fold assignment, so differences are attributable to the
procedure rather than to the fold draw.

Every arm is scored twice: against the known truth on the **training**
sample, and on an independent **test** sample of 2000 drawn from the same
DGP.

**There is no optimism to detect here, and the train-to-test gap isn't an
overfitting diagnostic.** Optimism is what you see when a model is scored
against the labels it was fit to. Every arm here is scored against the
*known true CATE*, while the label the stage-2 model actually saw is a noisy
pseudo-outcome. The test score measures how well the fitted surface
generalises to fresh covariate draws — the only like-for-like comparison of
the final fitted models across procedures (see [Generalisation
gap](#generalisation-gap)).

**Two test scores, because crossfit and whole-sample arms predict
differently.** A crossfit arm ends up with `V` fitted models and averages
their test predictions — the estimator you'd actually deploy, so `mse` uses
it — but that averaging is a variance-reducing ensemble stacked on top of the
honesty effect being studied. `mse_test_single` scores each fold model on the
test set separately and averages the `V` scores instead, which is the
like-for-like reading against a whole-sample arm's single model. The two
coincide for whole-sample arms, so any gap is the ensembling effect alone
(see [Ensemble vs. single fold model](#ensemble-vs-single-fold-model)).

### Arms compared

**DR-learner, random forest (8 arms)**

| id | stage 1 (nuisances) | stage 2 (final model) |
|---|---|---|
| `dcf` | double CF over fold pairs | crossfit, same folds, column `k` (**status quo**) |
| `scf_scf` | single CF, leave-one-fold-out | crossfit, same folds |
| `scf_scf_new` | single CF | crossfit, **fresh independent** split |
| `scf_oob` | single CF | whole sample, **OOB** predictions |
| `oob_oob` | whole sample, **OOB**, T-learner | whole sample, **OOB** |
| `oob_oob_s` | whole sample, **OOB**, S-learner (`X.orig` workaround) | whole sample, **OOB** |
| `oob_oob_manual` | whole sample, **OOB**, S-learner (manual tree-loop) | whole sample, **OOB** |
| `scf_oob_t` | single CF, T-learner | whole sample, OOB (**control**) |

`oob_oob_s` and `oob_oob_manual` give an S-learner a genuine OOB counterfactual
prediction — matching the S-learner structure used by every other nuisance in
this study, unlike `oob_oob`'s T-learner workaround. `oob_oob_s` uses the
maintainer-endorsed but unsupported `forest$X.orig` shortcut
([grf-labs/grf#307](https://github.com/grf-labs/grf/issues/307));
`oob_oob_manual` gets the same answer through grf's documented
`get_tree()`/`get_leaf_node()` API and exists mainly to check the shortcut
against it. See `crossfitting/README.md` for the full explanation.

**DR-learner, SuperLearner (3 arms):** `dcf`, `scf_scf`, `scf_scf_new` — no
OOB analogue exists for SuperLearner, so the OOB arms and the T-learner
control are dropped.

**Causal forest (4 arms)**

| id | `Y.hat` / `W.hat` | forest |
|---|---|---|
| `cf_dcf` | double-CF matrices, column `k` | fold-wise (**status quo**) |
| `cf_scf` | single-CF vectors | fold-wise |
| `cf_full_oob` | single-CF vectors | whole sample, OOB `tau` |
| `cf_default` | grf's own internal OOB | whole sample, OOB `tau` — plain `causal_forest(X, Y, W)` |

Full design rationale and deviations from the rest of the study are in
`crossfitting/README.md`.

```{r data}
# functions and label dictionaries
source(here("crossfitting", "cf_metrics.R"))

# paths
path <- here()
res_path <- file.path(dirname(path), "results", "crossfitting")

metrics <- readRDS(file.path(res_path, "cf_metrics.RDS"))

# tidy up
metrics <- metrics %>%
  mutate(
    scenario = factor(
      case_when(scenario == 1 ~ "Null",
                scenario == 4 ~ "Simple",
                scenario == 6 ~ "Interaction",
                scenario == 9 ~ "Non-linear"),
      levels = c("Null", "Simple", "Interaction", "Non-linear")),
    family = factor(recode(family, !!!family_labels), levels = unname(family_labels)),
    variant = factor(recode(variant, !!!variant_labels),
                     levels = unname(variant_labels[variant_levels])),
    set = factor(set, levels = c("train", "test"), labels = c("Training sample", "Test sample"))
  ) %>%
  droplevels()

# per variant summaries
metrics_summary <- metrics %>%
  group_by(scenario, family, variant, set) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(n()),
    mean_sign_acc = mean(sign_acc, na.rm = T),
    mcse_sign_acc = sd(sign_acc, na.rm = T) / sqrt(n()),
    mean_mse_single = mean(mse_test_single, na.rm = T),
    mcse_mse_single = sd(mse_test_single, na.rm = T) / sqrt(n()),
    mean_time = mean(time_total, na.rm = T),
    mean_time_nuisance = mean(time_nuisance, na.rm = T),
    mean_time_stage2 = mean(time_stage2, na.rm = T),
    .groups = "drop"
  )
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with MCSE error bars.
# facet_wrap (not facet_grid) because variant is family-specific - each
# family only populates 5-8 of the 13 levels, and facet_grid's free scales
# only vary per row/column, never per panel, so it would still show every
# panel all 13 categories. facet_wrap frees per panel, dropping the ones
# absent from that family's arms.
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = 0) {
  df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]]) %>%
    ggplot(aes(x = variant, y = est, colour = set, ymin = lo, ymax = hi)) +
    geom_hline(yintercept = hline, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_wrap(vars(family, scenario), nrow = n_distinct(df$family), scales = "free") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Crossfitting procedure", colour = "Scored on")
}
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth, on both samples, for every arm.

```{r mse-plot, fig.height=16}
mse_plot <- metrics %>%
  ggplot(aes(x = variant, y = mse, colour = set)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "MSE of CATE estimates by crossfitting procedure",
       y = "MSE", x = "Crossfitting procedure", colour = "Scored on")

mse_sum_plot <- summary_plot(metrics_summary, "mean_mse", "mcse_mse",
                             "Mean MSE of CATE estimates", "Mean MSE")

(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in the Null scenario since there's no heterogeneity to correlate with —
excluded from these plots).

```{r corr-plot, fig.height=16}
corr_sum_plot <- summary_plot(filter(metrics_summary, scenario != "Null"),
                              "mean_corr", "mcse_corr",
                              "Mean correlation with the true CATE", "Mean correlation")

corr_df <- filter(metrics, scenario != "Null")

corr_plot <- corr_df %>%
  ggplot(aes(x = variant, y = corr, colour = set)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  # free_x only: correlation is already on a bounded, comparable -1..1 scale,
  # so y stays shared - only the sparse variant axis needs freeing per panel
  facet_wrap(vars(family, scenario), nrow = n_distinct(corr_df$family), scales = "free_x") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation between estimated and true CATEs",
       y = "Pearson correlation", x = "Crossfitting procedure", colour = "Scored on")

(corr_sum_plot / corr_plot) + plot_layout(guides = "collect")
```

## Generalisation gap {#generalisation-gap}

Test MSE minus training MSE, per run. **This is not an optimism measure** —
every arm is scored against the true CATE rather than the pseudo-outcomes it
was fit to, so there's no optimism to detect; an in-sample prediction is
contaminated by the unit's own noisy pseudo-outcome and so scores worse
against truth, not better. What this plot actually shows is how far each
fitted surface degrades on fresh covariate draws.

```{r gap-plot}
gap <- metrics %>%
  select(scenario, family, variant, run, set, mse) %>%
  pivot_wider(names_from = set, values_from = mse) %>%
  mutate(gap = `Test sample` - `Training sample`)

gap_plot <- gap %>%
  ggplot(aes(x = variant, y = gap, colour = family)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(gap$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  labs(title = "Generalisation gap: test MSE minus training MSE",
       subtitle = "How far each fitted surface degrades on fresh covariate draws",
       y = "Test MSE - training MSE", x = "Crossfitting procedure")

gap_plot
```

## Ensemble vs. single fold model {#ensemble-vs-single-fold-model}

A crossfit arm deploys `V` fold models and averages their test predictions —
that's `mse`, the estimator you'd actually deploy. `mse_test_single` scores
each fold model on the test set separately and averages the `V` scores
instead, the like-for-like reading against a whole-sample arm's single
model. The two coincide for whole-sample arms, so any gap here is the
ensembling effect alone, stacked on top of whatever the crossfitting
procedure itself buys.

```{r ensemble-plot}
ensemble_df <- metrics_summary %>%
  filter(set == "Test sample") %>%
  select(scenario, family, variant, mean_mse, mean_mse_single) %>%
  pivot_longer(c(mean_mse, mean_mse_single),
               names_to = "scoring", values_to = "mse") %>%
  mutate(scoring = recode(scoring,
                          mean_mse = "Averaged fold models (deployed)",
                          mean_mse_single = "Single fold model (like-for-like)"))

ensemble_plot <- ensemble_df %>%
  ggplot(aes(x = variant, y = mse, colour = scoring)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(ensemble_df$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Test MSE: ensemble of fold models against a single fold model",
       subtitle = "Identical for whole-sample arms; the gap on crossfit arms is the ensembling effect",
       y = "Mean test MSE", x = "Crossfitting procedure", colour = "Scoring")

ensemble_plot
```

## Bias

```{r bias-plot, fig.height=10}
bias_plot <- metrics %>%
  ggplot(aes(x = variant, y = bias, colour = set)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(vars(family, scenario), nrow = n_distinct(metrics$family), scales = "free") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Bias of CATE estimates by crossfitting procedure",
       y = "Average bias (estimate - truth)", x = "Crossfitting procedure",
       colour = "Scored on")

bias_plot
```

## Runtime

What the extra fold pairs actually buy, in seconds.

```{r time-plot, fig.height=10}
time_df <- metrics_summary %>%
  filter(set == "Test sample") %>%
  select(scenario, family, variant, mean_time_nuisance, mean_time_stage2) %>%
  pivot_longer(c(mean_time_nuisance, mean_time_stage2),
               names_to = "stage", values_to = "seconds") %>%
  mutate(stage = recode(stage,
                        mean_time_nuisance = "Stage 1 (nuisances)",
                        mean_time_stage2 = "Stage 2 (final model)"))

time_plot <- time_df %>%
  ggplot(aes(x = variant, y = seconds, fill = stage)) +
  geom_col() +
  facet_wrap(vars(family, scenario), nrow = n_distinct(time_df$family), scales = "free") +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Mean runtime per replicate",
       y = "Seconds", x = "Crossfitting procedure", fill = "Stage")

time_plot
```

## Accuracy against cost: the decision plot

Test-sample MSE against runtime, on a log scale — down and to the left is
better. This is the plot that answers the study's headline question: is
double crossfitting (`dcf` / `cf_dcf`) actually on the efficient frontier, or
does a cheaper arm dominate it?

```{r frontier-plot}
frontier_plot <- metrics_summary %>%
  filter(set == "Test sample") %>%
  ggplot(aes(x = mean_time, y = mean_mse, colour = variant, shape = family)) +
  geom_point(size = 3) +
  facet_wrap(~scenario, scales = "free") +
  scale_x_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  labs(title = "Test-sample MSE against runtime",
       subtitle = "Down and to the left is better",
       y = "Mean test MSE", x = "Mean runtime per replicate (s, log scale)",
       colour = "Procedure", shape = "Estimator")

frontier_plot
```

## Headline table

Test-sample only, arranged by family, scenario, and mean MSE (best arm
first within each scenario).

```{r headline-table}
headline <- metrics_summary %>%
  filter(set == "Test sample") %>%
  select(scenario, family, variant, mean_mse, mcse_mse, mean_mse_single,
         mean_corr, mean_time) %>%
  arrange(family, scenario, mean_mse)

kable(headline, digits = 4,
      col.names = c("Scenario", "Family", "Procedure", "Mean MSE", "MCSE (MSE)",
                    "Mean MSE (single fold)", "Mean correlation", "Mean runtime (s)"))
```
