---
title: "Crossfitting comparison: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 11,
  fig.height = 8
)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Does double crossfitting earn its cost?

The simulations in the current project initially set out to use double cross-fitting,
following implementations seen in Hines' TE-VIMS paper. Two-stage estimators fit
nuisances over all `C(V,2)` fold *pairs*: `collate_predictions` assembles an
`n x V` pseudo-outcome matrix whose column `k` is untouched by fold `k`, and
the second stage trains on `po_matrix[, k]` before predicting fold `k`. At
`V = 10` that is 45 nuisance fits instead of 10, a 4.5x cost ($\frac{(n)(n-1)}{2n}$).

Under ordinary (single) crossfitting, a stage-2 training row `i` carries a
pseudo-outcome from a nuisance model trained on every fold except `i`'s - 
which *includes* the held-out test fold.

### Design

Fixed at **n = 500, V = 10 folds, scenarios Null / Simple / Interaction /
Non-linear, 100 replicates** (400 array jobs). Every arm is scored twice:
against the known truth on the **training** sample, and on an independent **test**
sample of 2000 drawn from the same DGP.

The test score measures how well the fitted surface
generalises to fresh covariate draws — the only like-for-like comparison of
the final fitted models across procedures (see [Generalisation
gap](#generalisation-gap)).

**Two test scores, because crossfit and whole-sample arms predict
differently.** A crossfit arm ends up with `V` fitted models and averages
their test predictions — the estimator you'd actually deploy, so `mse` uses
it — but that averaging is a variance-reducing ensemble stacked on top of the
honesty effect being studied. `mse_test_single` scores each fold model on the
test set separately and averages the `V` scores instead, which is the
like-for-like reading against a whole-sample arm's single model. The two
coincide for whole-sample arms, so any gap is the ensembling effect alone
(see [Ensemble vs. single fold model](#ensemble-vs-single-fold-model)).

### Arms compared

**DR-learner, random forest (4 arms)**

| id | stage 1 (nuisances) | stage 2 (final model) |
|---|---|---|
| `dcf` | double CF over fold pairs | crossfit, same folds, column `k` (**status quo**) |
| `scf_scf` | single CF, leave-one-fold-out | crossfit, same folds |
| `scf_oob` | single CF | whole sample, **OOB** predictions |
| `oob_oob` | whole sample, **OOB** | whole sample, **OOB** (production `dr_random_forest`) |

Every DR arm fits its outcome model separately in each treatment arm, as the
production DR-learners do, so the arms differ only in how they split the
sample. See `crossfitting/README.md` for the full explanation.

**DR-learner, SuperLearner (2 arms):** `dcf`, `scf_scf` (production
`dr_superlearner`) — no OOB analogue exists for SuperLearner, so the OOB arms
are dropped.

**Causal forest (4 arms)**

| id | `Y.hat` / `W.hat` | forest |
|---|---|---|
| `cf_dcf` | double-CF matrices, column `k` | fold-wise (**status quo**) |
| `cf_scf` | single-CF vectors | fold-wise |
| `cf_full_oob` | single-CF vectors | whole sample, OOB `tau` |
| `cf_default` | grf's own internal OOB | whole sample, OOB `tau` — plain `causal_forest(X, Y, W)` |

Full design rationale and deviations from the rest of the study are in
`crossfitting/README.md`.

```{r data}
# functions and label dictionaries
source(here("crossfitting", "cf_metrics.R"))

# paths
path <- here()
res_path <- file.path(dirname(path), "results", "crossfitting")

metrics <- readRDS(file.path(res_path, "cf_metrics.RDS"))

# tidy up
metrics <- metrics %>%
  mutate(
    scenario = factor(
      case_when(
        scenario == 1 ~ "Null",
        scenario == 6 ~ "Simple",
        scenario == 8 ~ "Interaction",
        scenario == 4 ~ "Non-linear"
      ),
      levels = c("Null", "Simple", "Interaction", "Non-linear")
    ),
    family = factor(
      recode(family, !!!family_labels),
      levels = unname(family_labels)
    ),
    variant = factor(
      recode(variant, !!!variant_labels),
      levels = unname(variant_labels[variant_levels])
    ),
    set = factor(
      set,
      levels = c("train", "test"),
      labels = c("Training sample", "Test sample")
    )
  ) %>%
  droplevels()

# per variant summaries
# mcse denominator is sqrt(sum(!is.na(x))), matching the non-NA count of that
# column - not sqrt(n()), the whole group's row count regardless of NAs
# (mse_test_single in particular is NA_real_ for whole-sample arms, see
# cf_metrics.R:50, so mcse_mse_single needs the narrower count)
metrics_summary <- metrics %>%
  group_by(scenario, family, variant, set) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(sum(!is.na(bias))),
    mean_ate_bias = mean(ate_bias, na.rm = T),
    mcse_ate_bias = sd(ate_bias, na.rm = T) / sqrt(sum(!is.na(ate_bias))),
    mean_rel_ate_bias = mean(rel_ate_bias, na.rm = T),
    mcse_rel_ate_bias = sd(rel_ate_bias, na.rm = T) /
      sqrt(sum(!is.na(rel_ate_bias))),
    mean_rel_bias_cate = mean(rel_bias_cate, na.rm = T),
    mcse_rel_bias_cate = sd(rel_bias_cate, na.rm = T) /
      sqrt(sum(!is.na(rel_bias_cate))),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(sum(!is.na(mse))),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(sum(!is.na(corr))),
    mean_sign_acc = mean(sign_acc, na.rm = T),
    mcse_sign_acc = sd(sign_acc, na.rm = T) / sqrt(sum(!is.na(sign_acc))),
    mean_mse_single = mean(mse_test_single, na.rm = T),
    mcse_mse_single = sd(mse_test_single, na.rm = T) /
      sqrt(sum(!is.na(mse_test_single))),
    mean_time = mean(time_total, na.rm = T),
    mean_time_nuisance = mean(time_nuisance, na.rm = T),
    mean_time_stage2 = mean(time_stage2, na.rm = T),
    .groups = "drop"
  )
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with a 95% CI (mean +/- qnorm(1 -
# alpha/2) x MCSE) error bar - see R/figures.R's point_range_plot() (its alpha doc) for
# why this isn't a raw +/- 1x MCSE (~68% coverage, not 95%).
# facet_wrap (not facet_grid) because variant is family-specific - each
# family only populates 2-4 of the 8 levels, and facet_grid's free scales
# only vary per row/column, never per panel, so it would still show every
# panel all 8 categories. facet_wrap frees per panel, dropping the ones
# absent from that family's arms.
# hline = NULL omits the reference line.
summary_plot <- function(
  df,
  mean_col,
  mcse_col,
  title,
  ylab,
  hline = 0,
  alpha = 0.05
) {
  z <- qnorm(1 - alpha / 2)
  df %>%
    mutate(
      est = .data[[mean_col]],
      lo = .data[[mean_col]] - z * .data[[mcse_col]],
      hi = .data[[mean_col]] + z * .data[[mcse_col]]
    ) %>%
    ggplot(aes(x = set, y = est, colour = variant, ymin = lo, ymax = hi)) +
    (if (!is.null(hline)) geom_hline(yintercept = hline, linetype = "dashed")) +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(
      position = position_dodge(width = 0.5),
      linewidth = 0.3,
      width = 0.3
    ) +
    facet_wrap(
      vars(family, scenario),
      nrow = n_distinct(df$family),
      scales = "free"
    ) +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black")
    ) +
    labs(title = title, y = ylab, x = "Evaluation set", colour = "Variant")
}
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth, on both samples, for every arm.

```{r mse-plot, fig.height=16}
mse_plot <- metrics %>%
  ggplot(aes(x = variant, y = mse, colour = set)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(metrics$family),
    scales = "free"
  ) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "MSE of CATE estimates by crossfitting procedure",
    y = "MSE",
    x = "Crossfitting procedure",
    colour = "Scored on"
  )

mse_sum_plot <- summary_plot(
  metrics_summary,
  "mean_mse",
  "mcse_mse",
  "Mean MSE of CATE estimates",
  "Mean MSE"
)

#(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
mse_sum_plot
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in the Null scenario since there's no heterogeneity to correlate with —
excluded from these plots).

```{r corr-plot, fig.height=16}
corr_sum_plot <- summary_plot(
  filter(metrics_summary, scenario != "Null"),
  "mean_corr",
  "mcse_corr",
  "Mean correlation with the true CATE",
  "Mean correlation",
  hline = NULL
)

corr_df <- filter(metrics, scenario != "Null")

corr_plot <- corr_df %>%
  ggplot(aes(x = variant, y = corr, colour = set)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  # free_x only: correlation is already on a bounded, comparable -1..1 scale,
  # so y stays shared - only the sparse variant axis needs freeing per panel
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(corr_df$family),
    scales = "free_x"
  ) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Correlation between estimated and true CATEs",
    y = "Pearson correlation",
    x = "Crossfitting procedure",
    colour = "Scored on"
  )

#(corr_sum_plot / corr_plot) + plot_layout(guides = "collect")
corr_sum_plot
```

## Generalisation gap {#generalisation-gap}

Test MSE minus training MSE, per run. **This is not an optimism measure** —
every arm is scored against the true CATE rather than the pseudo-outcomes it
was fit to, so there's no optimism to detect; an in-sample prediction is
contaminated by the unit's own noisy pseudo-outcome and so scores worse
against truth, not better. What this plot actually shows is how far each
fitted surface degrades on fresh covariate draws.

```{r gap-plot}
gap <- metrics %>%
  select(scenario, family, variant, run, set, mse) %>%
  pivot_wider(names_from = set, values_from = mse) %>%
  mutate(gap = `Test sample` - `Training sample`)

gap_plot <- gap %>%
  ggplot(aes(x = variant, y = gap, colour = family)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(gap$family),
    scales = "free"
  ) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Generalisation gap: test MSE minus training MSE",
    subtitle = "How far each fitted surface degrades on fresh covariate draws",
    y = "Test MSE - training MSE",
    x = "Crossfitting procedure"
  )

gap_plot
```

## Ensemble vs. single fold model {#ensemble-vs-single-fold-model}

A crossfit arm deploys `V` fold models and averages their test predictions —
that's `mse`, the estimator you'd actually deploy. `mse_test_single` scores
each fold model on the test set separately and averages the `V` scores
instead, the like-for-like reading against a whole-sample arm's single
model. The two coincide for whole-sample arms, so any gap here is the
ensembling effect alone, stacked on top of whatever the crossfitting
procedure itself buys.

```{r ensemble-plot}
ensemble_df <- metrics_summary %>%
  filter(set == "Test sample") %>%
  select(scenario, family, variant, mean_mse, mean_mse_single) %>%
  pivot_longer(
    c(mean_mse, mean_mse_single),
    names_to = "scoring",
    values_to = "mse"
  ) %>%
  mutate(
    scoring = recode(
      scoring,
      mean_mse = "Averaged fold models (deployed)",
      mean_mse_single = "Single fold model (like-for-like)"
    )
  )

ensemble_plot <- ensemble_df %>%
  ggplot(aes(x = variant, y = mse, colour = scoring)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(ensemble_df$family),
    scales = "free"
  ) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Test MSE: ensemble of fold models against a single fold model",
    subtitle = "Identical for whole-sample arms; the gap on crossfit arms is the ensembling effect",
    y = "Mean test MSE",
    x = "Crossfitting procedure",
    colour = "Scoring"
  )

ensemble_plot
```

## Bias

```{r bias-plot, fig.height=10}
bias_plot <- metrics %>%
  ggplot(aes(x = variant, y = bias, colour = set)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(metrics$family),
    scales = "free"
  ) +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Bias of CATE estimates by crossfitting procedure",
    y = "Average bias (estimate - truth)",
    x = "Crossfitting procedure",
    colour = "Scored on"
  )

bias_sum_plot <- summary_plot(
  metrics_summary,
  "mean_bias",
  "mcse_bias",
  "Mean bias of CATE estimates",
  "Mean bias"
)

bias_sum_plot
```

## Relative bias

`rel_ate_bias` is the parameter-level relative bias (ATE bias divided by the
true ATE); `rel_bias_cate` is the per-unit version, `(est - true) / true`
averaged over units (undefined and set to `NA` wherever the true CATE is
exactly 0 for a unit — see `R/metrics.R::cate_metrics()`).

```{r rel-bias-plot, fig.height=16}
rel_ate_bias_sum_plot <- summary_plot(
  metrics_summary,
  "mean_rel_ate_bias",
  "mcse_rel_ate_bias",
  "Mean relative bias in the ATE",
  "Mean relative ATE bias"
)
rel_bias_cate_sum_plot <- summary_plot(
  metrics_summary,
  "mean_rel_bias_cate",
  "mcse_rel_bias_cate",
  "Mean relative bias in the CATE",
  "Mean relative CATE bias"
)

rel_ate_bias_sum_plot
rel_bias_cate_sum_plot
```

## Runtime

What the extra fold pairs actually buy, in seconds.

```{r time-plot, fig.height=10}
time_df <- metrics_summary %>%
  filter(set == "Test sample") %>%
  select(scenario, family, variant, mean_time_nuisance, mean_time_stage2) %>%
  pivot_longer(
    c(mean_time_nuisance, mean_time_stage2),
    names_to = "stage",
    values_to = "seconds"
  ) %>%
  mutate(
    stage = recode(
      stage,
      mean_time_nuisance = "Stage 1 (nuisances)",
      mean_time_stage2 = "Stage 2 (final model)"
    )
  )

time_plot <- time_df %>%
  ggplot(aes(x = variant, y = seconds, fill = stage)) +
  geom_col() +
  facet_wrap(
    vars(family, scenario),
    nrow = n_distinct(time_df$family),
    scales = "free"
  ) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Mean runtime per replicate",
    y = "Seconds",
    x = "Crossfitting procedure",
    fill = "Stage"
  )

time_plot
```

## Accuracy against cost: the decision plot

Test-sample MSE against runtime, on a log scale — down and to the left is
better. This is the plot that answers the study's headline question: is
double crossfitting (`dcf` / `cf_dcf`) actually on the efficient frontier, or
does a cheaper arm dominate it?

```{r frontier-plot}
frontier_plot <- metrics_summary %>%
  filter(set == "Test sample") %>%
  ggplot(aes(x = mean_time, y = mean_mse, colour = variant, shape = family)) +
  geom_point(size = 3) +
  facet_wrap(~scenario, scales = "free") +
  scale_x_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Test-sample MSE against runtime",
    subtitle = "Down and to the left is better",
    y = "Mean test MSE",
    x = "Mean runtime per replicate (s, log scale)",
    colour = "Procedure",
    shape = "Estimator"
  )

frontier_plot
```

## Headline table

Test-sample only, one table per estimator family, arranged by scenario and
mean MSE (best arm first within each scenario).

```{r headline-table-helper}
headline_table <- function(fam) {
  metrics_summary %>%
    filter(set == "Test sample", family == fam) %>%
    select(
      scenario,
      variant,
      mean_mse,
      mcse_mse,
      mean_mse_single,
      mean_corr,
      mean_time
    ) %>%
    arrange(scenario, mean_mse) %>%
    kable(
      digits = 4,
      col.names = c(
        "Scenario",
        "Procedure",
        "Mean MSE",
        "MCSE (MSE)",
        "Mean MSE (single fold)",
        "Mean correlation",
        "Mean runtime (s)"
      )
    )
}
```

### Causal forest

```{r headline-table-cf}
headline_table(family_labels[["causal_forest"]])
```

### DR-learner (RF)

```{r headline-table-dr-rf}
headline_table(family_labels[["dr_rf"]])
```

### DR-learner (SuperLearner)

```{r headline-table-dr-sl}
headline_table(family_labels[["dr_sl"]])
```

## Decision

The MCSEs above are marginal — they treat each arm's 100 replicates as
independent. But every arm is fit to the same simulated dataset within a
replicate, so dataset-to-dataset noise is shared across arms and cancels in a
per-run difference. Comparing arms on the paired difference
`mse[arm] - mse[reference]` within each run gives a much tighter (and correct)
MCSE than eyeballing overlapping error bars.

**Decision rule** (fixed before reading the results): within each family, use
the **cheapest arm that is non-inferior to the best arm in all four
scenarios**, where non-inferior means the upper 95% bound of its paired
relative test-MSE difference against the best arm in that scenario is below
`ni_margin`. Correlation and bias (above) are guardrails: a picked arm
shouldn't be visibly worse on either.

```{r paired-helper}
ni_margin <- 0.05 # 5% relative test-MSE margin

status_quo <- setNames(
  variant_labels[c("cf_dcf", "dcf", "dcf")],
  family_labels[c("causal_forest", "dr_rf", "dr_sl")]
)

# paired per-run MSE differences against a reference arm
# ref: one row per scenario x family, naming the reference arm in ref_variant
# eval_set: switch to "Training sample" if the main study scores CATEs
# in-sample rather than on fresh draws
paired_diff <- function(ref, eval_set = "Test sample", alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  scored <- filter(metrics, set == eval_set)

  ref_mse <- scored %>%
    inner_join(ref, by = c("scenario", "family")) %>%
    filter(variant == ref_variant) %>%
    select(scenario, family, run, ref_variant, ref_mse = mse)

  # relative difference is a ratio of means, mean(diff) / mean(ref_mse), not
  # mean(diff / ref_mse) - the per-run ratio is heavy-tailed whenever a run's
  # reference MSE is small (Null especially). its MCSE is the delta-method
  # (ratio estimator) linearisation sd(diff - rel * ref_mse) / (sqrt(n) * mean(ref_mse))
  scored %>%
    inner_join(ref_mse, by = c("scenario", "family", "run")) %>%
    mutate(diff = mse - ref_mse) %>%
    filter(!is.na(diff)) %>%
    group_by(scenario, family, variant, ref_variant) %>%
    summarise(
      n_runs = n(),
      mean_diff = mean(diff),
      mcse_diff = sd(diff) / sqrt(n_runs),
      mean_rel_diff = mean_diff / mean(ref_mse),
      mcse_rel_diff = sd(diff - mean_rel_diff * ref_mse) /
        (sqrt(n_runs) * mean(ref_mse)),
      p_better = mean(diff < 0), # share of runs where this arm beats the reference
      .groups = "drop"
    ) %>%
    mutate(
      lo_rel = mean_rel_diff - z * mcse_rel_diff,
      hi_rel = mean_rel_diff + z * mcse_rel_diff
    )
}

ref_status_quo <- metrics %>%
  distinct(scenario, family) %>%
  mutate(ref_variant = unname(status_quo[as.character(family)]))

ref_best <- metrics_summary %>%
  filter(set == "Test sample") %>%
  group_by(scenario, family) %>%
  slice_min(mean_mse, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  transmute(scenario, family, ref_variant = as.character(variant))

vs_status_quo <- paired_diff(ref_status_quo)
vs_best <- paired_diff(ref_best)
```

### Against the status quo

Paired relative difference in test MSE against double crossfitting (`dcf` /
`cf_dcf`), with a 95% CI. Left of zero means the arm beats the status quo;
the dotted line is the non-inferiority margin.

```{r paired-plot, fig.height=9}
paired_plot <- vs_status_quo %>%
  filter(variant != ref_variant) %>%
  ggplot(aes(
    x = 100 * mean_rel_diff,
    y = variant,
    xmin = 100 * lo_rel,
    xmax = 100 * hi_rel
  )) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 100 * ni_margin, linetype = "dotted") +
  geom_pointrange(size = 0.3) +
  # free_y + space so each family row only shows its own arms
  facet_grid(family ~ scenario, scales = "free", space = "free_y") +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "Test MSE relative to double crossfitting (paired by run)",
    subtitle = "Left of zero is better than the status quo; dotted line is the non-inferiority margin",
    x = "Relative difference in test MSE (%)",
    y = NULL
  )

paired_plot
```

### Decision tables

One row per arm, cheapest first. *Worst upper bound* is the largest upper 95%
bound of the paired relative MSE difference against the best arm across the
four scenarios; *Worst P(beats status quo)* is the lowest share of runs, across
scenarios, in which the arm beat double crossfitting. The pick is the cheapest
arm non-inferior in all four scenarios.

```{r decision-table-helper}
decision_table <- function(fam) {
  # kable formats on the call, so the NA placeholder only needs to hold in here
  op <- options(knitr.kable.NA = "—")
  on.exit(options(op))

  runtime <- metrics_summary %>%
    filter(set == "Test sample", family == fam) %>%
    group_by(variant) %>%
    summarise(mean_time = mean(mean_time), .groups = "drop")

  # the status quo arm itself is left out, so its row reads as blank
  sq <- vs_status_quo %>%
    filter(family == fam, variant != ref_variant) %>%
    group_by(variant) %>%
    summarise(worst_p_better = min(p_better), .groups = "drop")

  vs_best %>%
    filter(family == fam) %>%
    group_by(variant) %>%
    summarise(
      worst_hi_rel = max(hi_rel),
      n_ni = sum(hi_rel < ni_margin),
      n_scen = n(),
      .groups = "drop"
    ) %>%
    left_join(sq, by = "variant") %>%
    left_join(runtime, by = "variant") %>%
    arrange(mean_time) %>%
    mutate(
      eligible = n_ni == n_scen,
      pick = ifelse(eligible & cumsum(eligible) == 1, "✓", ""),
      worst_hi_rel = 100 * worst_hi_rel,
      ni = paste0(n_ni, "/", n_scen)
    ) %>%
    select(variant, mean_time, worst_hi_rel, ni, worst_p_better, pick) %>%
    kable(
      digits = c(0, 2, 1, 0, 2, 0),
      col.names = c(
        "Procedure",
        "Mean runtime (s)",
        "Worst upper bound vs best (%)",
        "Non-inferior in",
        "Worst P(beats status quo)",
        "Pick"
      )
    )
}
```

#### Causal forest

```{r decision-table-cf}
decision_table(family_labels[["causal_forest"]])
```

#### DR-learner (RF)

```{r decision-table-dr-rf}
decision_table(family_labels[["dr_rf"]])
```

#### DR-learner (SuperLearner)

```{r decision-table-dr-sl}
decision_table(family_labels[["dr_sl"]])
```
