---
title: "Competing risks: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Can these estimators recover treatment effects when a competing event can pre-empt the event of interest?

Event of interest (EOI) and competing event (CE) are generated from Weibull
distributions via joint hazards (Beyersmann 2009). Seven scenarios vary
*where* the treatment acts — on the EOI, the CE, or both, as an average
effect (ATE) or heterogeneous one (HTE):

| Scenario | Treatment effect |
|---|---|
| 1 | ATE on EOI only |
| 2 | ATE on CE only |
| 3 | HTE on EOI, no ATE on CE |
| 4 | HTE on EOI, ATE on CE |
| 5 | HTE on CE, no ATE on EOI |
| 6 | HTE on CE, ATE on EOI |
| 7 | HTE on both |

Scenarios 2, 5 and 6 are the interesting ones: the treatment moves the
competing event, so a naive analysis of the EOI alone can show an apparent
effect that is really just a change in who survives long enough to have one.
Fixed at **n = 500**, censoring **on/off**, **100 runs**, horizon = 28. Full
design in `competing_risk/README.md`.

**Five estimator frameworks**, several shipping in multiple arms that vary
how their pseudo-values/nuisances are built (see `README.md`'s
"Crossfitting" section for the full rationale):

| Family | Arms | Targets |
|---|---|---|
| `ipw` | causal forest on IPW-transformed RMST | cause-specific RMST |
| `csf_cs` | causal survival forest, competing events censored | cause-specific RMST |
| `csf_sh` | causal survival forest, competing events kept in the risk set (Fine-Gray) | subdistribution RMST |
| Pseudo-value CF | `whole_oob`, `whole_scf`, `cvps_scf` | RMTL |
| Pseudo-value DR | `whole_oob`, `whole_scf`, `cvps_scf` | RMTL |
| SL T-learner | `whole`, `cvps`, `split` | RMTL |
| SL DR-learner | `whole`, `cvps` | RMTL |

**The truth column depends on the family**: `ipw`/`csf_cs` remove competing
events and target the cause-specific (net) RMST; `csf_sh` keeps them in the
risk set and targets the subdistribution RMST; every pseudo-value family
targets RMTL. `surv_metrics.R` already resolves this per row, so `bias`/
`mse`/etc. below are always scored against the correct truth.

**Scenario 5 flips the sign of the `RMTL1` correlation — expect it, it is
not a bug.** `tau_RMST1_cs` is flat in scenario 5 (no cause-specific effect
on the EOI at all), so every bit of `RMTL1` heterogeneity there is the
competing-risk artefact induced by the (near-mirror-image) `RMTL2`
heterogeneity. An estimator that tracks the real signal correctly therefore
lands *anti-correlated* with the induced `RMTL1` estimand. A consequence:
`tau_RMST1_cs` has zero variance in scenario 5, so `corr`/`spearman`/
`c_stat` are genuine `NA` for the `ipw`/`csf_cs` Event 1 cells there — dropped
by `na.rm = TRUE` below, not treated as 0. See `competing_risk/README.md`
for the full argument.

```{r data}
res_path <- file.path(dirname(here()), "results", "competing_risk")
metrics <- readRDS(file.path(res_path, "surv_metrics.RDS"))

# display order and labels - no shared label-dictionary script for this study
# (unlike crossfitting/cf_metrics.R), so these stay inline, matching
# continuous/cts_results.qmd's precedent
scenario_labels <- c(
  `1` = "ATE E1",
  `2` = "ATE E2",
  `3` = "HTE E1",
  `4` = "HTE E1, ATE E2",
  `5` = "HTE E2",
  `6` = "HTE E2, ATE E1",
  `7` = "HTE E1&2"
)

# framework is split into family (the estimator) and variant (the arm),
# mirroring crossfitting/cf_results.qmd's family/variant split - keeps
# facets legible instead of one flat 14-level factor
framework_family <- c(
  ipw = "IPW", csf_cs = "CSF (cause-specific)", csf_sh = "CSF (subdistribution)",
  pseudo_cf_whole_oob = "Pseudo-value CF", pseudo_cf_whole_scf = "Pseudo-value CF",
  pseudo_cf_cvps_scf  = "Pseudo-value CF",
  pseudo_dr_whole_oob = "Pseudo-value DR", pseudo_dr_whole_scf = "Pseudo-value DR",
  pseudo_dr_cvps_scf  = "Pseudo-value DR",
  sl_t_whole = "SL T-learner", sl_t_cvps = "SL T-learner", sl_t_split = "SL T-learner",
  sl_dr_whole = "SL DR-learner", sl_dr_cvps = "SL DR-learner"
)
family_levels <- c("IPW", "CSF (cause-specific)", "CSF (subdistribution)",
                    "Pseudo-value CF", "Pseudo-value DR",
                    "SL T-learner", "SL DR-learner")

framework_variant <- c(
  ipw = "single", csf_cs = "single", csf_sh = "single",
  pseudo_cf_whole_oob = "whole_oob", pseudo_cf_whole_scf = "whole_scf",
  pseudo_cf_cvps_scf  = "cvps_scf",
  pseudo_dr_whole_oob = "whole_oob", pseudo_dr_whole_scf = "whole_scf",
  pseudo_dr_cvps_scf  = "cvps_scf",
  sl_t_whole = "whole", sl_t_cvps = "cvps", sl_t_split = "split",
  sl_dr_whole = "whole", sl_dr_cvps = "cvps"
)
variant_levels <- c("single", "whole_oob", "whole_scf", "cvps_scf", "whole", "cvps", "split")
variant_labels <- c(
  single    = "(single arm)",
  whole_oob = "Whole-sample, OOB",
  whole_scf = "Whole-sample, single CF",
  cvps_scf  = "Crossfit PV, single CF",
  whole     = "Whole-sample PV",
  cvps      = "Crossfit PV",
  split     = "Split PV"
)

# tidy up
metrics <- metrics %>%
  mutate(
    family    = factor(recode(framework, !!!framework_family), levels = family_levels),
    variant   = recode(framework, !!!framework_variant),
    variant   = factor(recode(variant, !!!variant_labels),
                       levels = unname(variant_labels[variant_levels])),
    scenario  = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                       labels = unname(scenario_labels)),
    censoring = factor(censoring, levels = c(TRUE, FALSE),
                       labels = c("Censored", "Uncensored")),
    target    = factor(target, levels = c("Event 1", "Event 2", "Combined"))
  ) %>%
  droplevels()

# remove boring bits
metrics <- metrics %>%
  # remove CSH stuff for conference presentations
  filter(!(family %in% c("CSF (cause-specific)", "IPW"))) %>%
  # remove combined event
  filter(target != "Combined") %>%
  # remove superfluous crossfitting variants
  filter(!(variant %in% c("Split PV", "Whole-sample PV", "Whole-sample, OOB", "Whole-sample, single CF"))) %>%
  droplevels()

# per (scenario, censoring, target, family, variant) summaries
metrics_summary <- metrics %>%
  group_by(scenario, censoring, target, family, variant) %>%
  summarise(
    mean_bias = mean(bias, na.rm = T),
    mcse_bias = sd(bias, na.rm = T) / sqrt(n()),
    mean_mse = mean(mse, na.rm = T),
    mcse_mse = sd(mse, na.rm = T) / sqrt(n()),
    mean_rmse = mean(rmse, na.rm = T),
    mcse_rmse = sd(rmse, na.rm = T) / sqrt(n()),
    mean_mae = mean(mae, na.rm = T),
    mcse_mae = sd(mae, na.rm = T) / sqrt(n()),
    mean_corr = mean(corr, na.rm = T),
    mcse_corr = sd(corr, na.rm = T) / sqrt(n()),
    mean_spearman = mean(spearman, na.rm = T),
    mcse_spearman = sd(spearman, na.rm = T) / sqrt(n()),
    mean_sign_acc = mean(sign_acc, na.rm = T),
    mcse_sign_acc = sd(sign_acc, na.rm = T) / sqrt(n()),
    mean_c_stat = mean(c_stat, na.rm = T),
    mcse_c_stat = sd(c_stat, na.rm = T) / sqrt(n()),
    mean_n_na = mean(n_na, na.rm = T),
    total_n_na = sum(n_na, na.rm = T),
    .groups = "drop"
  )
```

### Results coverage

The array job set is 1,400 runs (7 scenarios x 2 censoring settings x 100
runs); `competing_risk/README.md` documents known partial-failure routes, so
this table is worth checking before reading anything else below.

```{r completeness-table}
completeness <- metrics %>%
  group_by(scenario, censoring) %>%
  summarise(n_runs = n_distinct(run), n_frameworks = n_distinct(family, variant),
            .groups = "drop") %>%
  arrange(scenario, censoring)

kable(completeness,
      col.names = c("Scenario", "Censoring", "Runs collected", "Arms present"))
```

```{r summary-plot-helper, include=FALSE}
# helper: the house summary plot, points with MCSE error bars, faceted by
# family x scenario x censoring - extends crossfitting/cf_results.qmd's
# family x scenario facet with the censoring dimension this study has and
# the precedents don't. facet_wrap (not facet_grid) because variant is
# family-specific, same reasoning as cf_results.qmd.
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = 0) {
  df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]]) %>%
    ggplot(aes(x = family, y = est, ymin = lo, ymax = hi, colour = censoring)) +
    geom_hline(yintercept = hline, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(target), cols = vars(scenario), scales = "free") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
    labs(title = title, y = ylab, x = "Model", colour = "Censoring")
}
```

## Bias

Average bias in the estimated CATE (estimate minus truth), against each
target's own truth column.

```{r bias-plot, fig.height=8}
summary_plot(metrics_summary, "mean_bias", "mcse_bias",
             "Mean bias in CATE estimates", "Mean bias")
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth.

```{r mse-plot, fig.height=10}
summary_plot(metrics_summary, "mean_mse", "mcse_mse",
             "Mean MSE of CATE estimates", "Mean MSE")
```

## RMSE of CATE estimates

`sqrt(mse)` — the same ranking as MSE, on the outcome's own scale.

```{r rmse-plot, fig.height=10}
summary_plot(metrics_summary, "mean_rmse", "mcse_rmse",
             "Mean RMSE of CATE estimates", "Mean RMSE")
```

## MAE of CATE estimates

Mean absolute error — a robustness check on MSE, less sensitive to outlying
runs.

```{r mae-plot, fig.height=10}
summary_plot(metrics_summary, "mean_mae", "mcse_mae",
             "Mean MAE of CATE estimates", "Mean MAE")
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs. Undefined (forced to
0 by `cate_metrics()`) in scenario 1, since there's no heterogeneity to
correlate with — excluded here. Also see the scenario-5 note above: the
`ipw`/`csf_cs` Event 1 cells there are genuine `NA` (zero-variance truth),
dropped by `na.rm = TRUE` rather than shown as 0.

```{r corr-plot, fig.height=10}
summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
             "mean_corr", "mcse_corr",
             "Mean correlation with the true CATE", "Mean correlation")
```

## Spearman rank correlation with the true CATE

Rank correlation alongside the Pearson correlation above — less sensitive to
outlying CATE estimates and non-linear-but-monotone miscalibration. Same
scenario-1-undefined convention, so scenario 1 is excluded here too.

```{r spearman-plot, fig.height=10}
summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
             "mean_spearman", "mcse_spearman",
             "Mean Spearman correlation with the true CATE", "Mean Spearman correlation")
```

## Sign accuracy

Proportion of units where the estimated CATE has the correct sign. 0.5
(dashed line) is chance level.

```{r sign-acc-plot, fig.height=10}
summary_plot(metrics_summary, "mean_sign_acc", "mcse_sign_acc",
             "Mean sign accuracy", "Mean proportion correct sign", hline = 0.5)
```

## C-statistic

`(Kendall's tau_b + 1) / 2`, a rank-concordance measure equivalent to
Harrell's C for a continuous outcome. Forced to 0.5 in scenario 1 (no
heterogeneity to rank), so it's excluded here for the same reason as the
correlation metrics above; 0.5 (dashed line) is chance level elsewhere.

```{r c-stat-plot, fig.height=10}
summary_plot(filter(metrics_summary, scenario != scenario_labels[["1"]]),
             "mean_c_stat", "mcse_c_stat",
             "Mean C-statistic", "Mean C-statistic", hline = 0.5)
```

## Missing estimates

`n_na` is a data-quality diagnostic, not a performance metric: the count of
`NA` CATE estimates within a run (see `competing_risk/README.md`'s "Known
issues" for the failure routes that produce these). Mostly zero, so shown as
a table rather than a distribution plot; only cells with at least one
missing estimate are listed.

```{r na-table}
na_table <- metrics_summary %>%
  select(scenario, censoring, target, family, variant, mean_n_na, total_n_na) %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, censoring, target, family, variant)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("Scenario", "Censoring", "Target", "Family", "Arm",
                      "Mean NA count per run", "Total NA count"))
} else {
  cat("No missing CATE estimates in any collected run.")
}
```

## Headline table

Arranged by scenario, censoring, target, and mean MSE (best arm first
within each cell).

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, censoring, target, family, variant,
         mean_bias, mean_mse, mean_rmse, mean_mae, mean_corr, mean_c_stat) %>%
  arrange(scenario, censoring, target, mean_mse)

kable(headline, digits = 4,
      col.names = c("Scenario", "Censoring", "Target", "Family", "Arm",
                    "Mean bias", "Mean MSE", "Mean RMSE", "Mean MAE",
                    "Mean correlation", "Mean C-statistic"))
```
