---
title: "Model evaluation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Which model evaluation metrics work well in RCT settings?

In real life we cannot evaluate models with true PEHE because the true CATEs are
never observed. Different surrogate metrics designs to select the best CATE model
among several have been designed. Typical application has been in large observational
datasets and rely on ML machinery, so we want to know how well these techniques
will work in RCT settings where sample size is much smaller.
A proxy is useful if, run by run, its argmin is the
same candidate true PEHE's argmin is; and if when it differs, it doesn't
differ expensively.

### Design

Four scenarios crossed with three sample sizes, 30 replicates each — **360
array jobs**. Scenarios are indices into `R/dgm_scenarios.R`'s
`set = "continuous"`; the subset 1 / 4 / 6 / 9 and its display labels are the
same ones `crossfitting/cf_results.qmd` uses, so the two reports are directly
comparable.

| | |
|---|---|
| scenarios | 1 (Null), 4 (Simple), 6 (Interaction), 9 (Non-linear) |
| n | 250, 500, 1000 |
| runs | 30 |
| folds | V = 10, single crossfitting, all n |


### The 9 candidates

Each is fit as its own single-crossfit DR-learner (`me_models.R`). `p` is the
number of covariates, 7–9 across these scenarios.

| id | family | hyperparameters |
|---|---|---|
| `rf1` | random forest | ranger defaults (`mtry = floor(sqrt(p))`) |
| `rf2` | random forest | `mtry = p, max.depth = 5` |
| `rf3` | random forest | `mtry = ceiling(p/2), max.depth = 3` |
| `net1` | elastic net | `alpha = 1` (lasso) |
| `net2` | elastic net | `alpha = 0` (ridge) |
| `net3` | elastic net | `alpha = 0.5` |
| `SL1` | SuperLearner | `glmnet, ranger, earth, gam, mean` |
| `SL2` | SuperLearner | `glmnet, xgboost, cforest, earth, gam, mean` |
| `SL3` | SuperLearner | `svm, nnet, mean` |

### The 8 proxy scores

A second, *independent* nuisance pipeline estimates `mu0_T`, `mu1_T`, `mu_DR`
and `pi`, and is used only to score the candidates — never to fit them. Two
pipelines, two fold regimes, two score types:

| axis | levels |
|---|---|
| pipeline | **XGBoost** (36-combination CV grid) / **H2O AutoML** (up to 20 auto-tuned models) |
| fold regime | **cv** (leave-one-fold-out) / **whole** (no split) |
| score type | **infl** — influence-corrected PEHE proxy (`calc_infl_score`) / **dr** — DR risk, MSE against the AIPW pseudo-outcome `phi` (`calc_dr_risk`) |

All eight, and `true_pehe` itself, are **losses: lower is better**.

**`whole` is resubstitution.** It fits and predicts the scoring nuisances on
the entire sample, so they see the same rows the candidate's `tau_hat` was fit
on. `cv` is the honest regime. This was done to see if there is a tradeoff between
data leakage and more samples for AutoML/XGB accuracy.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "model_evaluation")

metrics <- readRDS(file.path(res_path, "me_metrics.RDS"))

# the proxy columns are whatever me_metrics.R produced that isn't a key or the
# truth - derived rather than hardcoded so that re-adding a fold regime (the
# `infold` one me_nuisance.R removed) needs no edit here
key_cols <- c("scenario", "n", "run", "model")
score_cols <- setdiff(names(metrics), c(key_cols, "true_pehe"))

# labels. Scenario labels match crossfitting/cf_results.qmd, which reports the
# same 1/4/6/9 subset.
scenario_labels <- c(`1` = "Null", `4` = "Simple", `6` = "Interaction",
                     `9` = "Non-linear")

model_levels <- c("rf1", "rf2", "rf3", "net1", "net2", "net3",
                  "SL1", "SL2", "SL3")

family_labels <- c(rf = "Random forest", net = "Elastic net",
                   SL = "SuperLearner")

# proxy display labels, ordered score type first so the infl/dr split reads
# straight off the x axis
proxy_labels <- c(
  infl_cv_xgb       = "Infl / CV / XGB",
  infl_cv_automl    = "Infl / CV / AutoML",
  infl_whole_xgb    = "Infl / whole / XGB",
  infl_whole_automl = "Infl / whole / AutoML",
  dr_cv_xgb         = "DR / CV / XGB",
  dr_cv_automl      = "DR / CV / AutoML",
  dr_whole_xgb      = "DR / whole / XGB",
  dr_whole_automl   = "DR / whole / AutoML"
)

axis_labels <- c(score_type = "Score type", fold_type = "Fold regime",
                 pipeline = "Nuisance pipeline")

level_labels <- c(infl = "Influence", dr = "DR risk", cv = "CV",
                  whole = "Whole", xgb = "XGBoost", automl = "AutoML")

metrics <- metrics %>%
  mutate(
    scenario = factor(scenario_labels[as.character(scenario)],
                      levels = unname(scenario_labels)),
    n = factor(n, levels = sort(unique(n))),
    model = factor(model, levels = model_levels),
    family = factor(family_labels[sub("[0-9]+$", "", as.character(model))],
                    levels = unname(family_labels))
    # maybe adding version: 1,2,3 would make plots nicer?
  )
```

```{r long-form}
# one row per (scenario, n, run, model, proxy). The three design axes are
# recovered from the column name, which me_metrics.R builds as
# <score_type>_<fold_type>_<pipeline>.
long <- metrics %>%
  pivot_longer(all_of(score_cols), names_to = "proxy_id", values_to = "score") %>%
  mutate(
    score_type = sub("_.*$", "", proxy_id),
    fold_type  = sub("^[^_]+_([^_]+)_.*$", "\\1", proxy_id),
    pipeline   = sub("^.*_", "", proxy_id),
    proxy = factor(proxy_labels[proxy_id], levels = unname(proxy_labels))
  )

# a candidate only enters a ranking if both its score and its truth are there
long_ok <- filter(long, !is.na(score), !is.na(true_pehe))
```

```{r selection}
# The core derivation: one row per (scenario, n, run, proxy), summarising how
# that proxy ranked the 9 candidates against how true PEHE ranked them.
#
# `regret` is measured against the best candidate the proxy could actually have
# picked (the minimum over the candidates present in that run), not against a
# hypothetical complete set - so a run with a dropped candidate is scored
# honestly rather than penalised for the gap.
selection <- long_ok %>%
  group_by(scenario, n, run, proxy, proxy_id, score_type, fold_type, pipeline) %>%
  filter(n() >= 2) %>%
  summarise(
    n_candidates = n(),
    rho          = cor(score, true_pehe, method = "spearman"),
    kendall      = cor(score, true_pehe, method = "kendall"),
    pick         = model[which.min(score)],
    best         = model[which.min(true_pehe)],
    hit          = as.numeric(pick == best),
    pick_rank    = rank(true_pehe, ties.method = "min")[which.min(score)],
    regret       = true_pehe[which.min(score)] - min(true_pehe),
    rel_regret   = regret / min(true_pehe),
    spread       = max(true_pehe) - min(true_pehe),
    # expected cost of ignoring the scores and picking one of the candidates
    # uniformly at random - the floor any useful proxy has to beat
    random_regret = mean(true_pehe) - min(true_pehe),
    .groups = "drop"
  )

# per (scenario, n, proxy) summaries, mean +/- Monte Carlo SE over the runs
selection_summary <- selection %>%
  group_by(scenario, n, proxy, score_type, fold_type, pipeline) %>%
  summarise(
    runs = n(),
    mean_rho = mean(rho, na.rm = T),
    mcse_rho = sd(rho, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_rank = mean(pick_rank, na.rm = T),
    mcse_rank = sd(pick_rank, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    mean_rel_regret = mean(rel_regret, na.rm = T),
    mcse_rel_regret = sd(rel_regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )
```

```{r baselines}
# Two reference points, both on the regret scale, both per (scenario, n).
#
# random pick: the expected regret of ignoring the scores entirely.
random_baseline <- selection %>%
  distinct(scenario, n, run, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(random_regret = mean(random_regret, na.rm = T), .groups = "drop")

# best fixed choice: commit to one candidate a priori and always use it. Note
# this is picked WITH hindsight - the best of the nine is chosen using the
# truth - so it is an optimistic upper bound on what a fixed rule could really
# achieve. A proxy that cannot beat it has not earned its cost.
oracle_by_run <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  summarise(oracle_pehe = min(true_pehe), .groups = "drop")

fixed_choice <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  left_join(oracle_by_run, by = c("scenario", "n", "run")) %>%
  mutate(fixed_regret = true_pehe - oracle_pehe) %>%
  group_by(scenario, n, model) %>%
  summarise(mean_regret = mean(fixed_regret, na.rm = T),
            mcse_regret = sd(fixed_regret, na.rm = T) / sqrt(n()),
            .groups = "drop")

best_fixed <- fixed_choice %>%
  group_by(scenario, n) %>%
  slice_min(mean_regret, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  rename(best_fixed_model = model, best_fixed_regret = mean_regret)
```

```{r plot-helpers, include=FALSE}
# The house pair of plots - per-run distribution and mean with MCSE error bars.
# The sweep (the 8 proxies) goes on x and the design on the facets, following
# confidence_intervals/continuous/cts_ci_results.qmd rather than
# cts_results.qmd's facet_wrap(~scenario).

box_plot <- function(df, metric, title, ylab, hline = NULL) {
  p <- ggplot(df, aes(x = proxy, y = .data[[metric]], colour = score_type))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe",
                             labels = c(dr = "DR risk", infl = "Influence")) +
    theme_light() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
    labs(title = title, y = ylab, x = "Proxy score", colour = "Score type")
}

summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = n, y = est, colour = score_type, shape = pipeline, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(fold_type), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe",
                             labels = c(dr = "DR risk", infl = "Influence")) +
    theme_light() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
    labs(title = title, y = ylab, x = "sample size", colour = "Score type", shape = "Machinery")
}

sum_ns_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = fold_type, y = est, colour = score_type, shape = pipeline, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe",
                             labels = c(dr = "DR risk", infl = "Influence")) +
    theme_light() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
    labs(title = title, y = ylab, x = "Fold type", colour = "Score type", shape = "Machinery")
}
```

## Data completeness

The study was still `in_progress` at the last campaign check.
Missing runs shrink the number of rankings each cell is averaged over; missing
*scores* within a present run shrink the number of candidates a proxy ranked.
2 runs in the null $n=250$ configuration have not run, currently troubleshooting.

```{r completeness-runs}
runs_present <- metrics %>%
  distinct(scenario, n, run) %>%
  count(scenario, n, name = "runs") %>%
  pivot_wider(names_from = n, values_from = runs, names_prefix = "n = ")

kable(runs_present, caption = "Replicates present per cell (30 expected)")
```

```{r completeness-na}
na_counts <- metrics %>%
  summarise(across(all_of(c("true_pehe", score_cols)), ~ sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  mutate(pct = 100 * n_missing / nrow(metrics))

kable(na_counts, digits = 2,
      col.names = c("Score column", "Missing values", "% of rows"),
      caption = "Missing scores across all (scenario, n, run, candidate) rows")
```

```{r completeness-candidates}
short_rankings <- selection %>%
  filter(n_candidates < length(model_levels)) %>%
  count(scenario, n, proxy, n_candidates, name = "rankings")

if (nrow(short_rankings) == 0) {
  cat("Every ranking scored all", length(model_levels), "candidates.\n")
} else {
  kable(short_rankings,
        col.names = c("Scenario", "n", "Proxy", "Candidates ranked", "Rankings"),
        caption = "Rankings built from fewer than the full candidate set")
}
```

## Examining True PEHE

If all 9 candidates achieve near-identical true PEHE then picking the wrong one
costs nothing and a high error rate is harmless; if they are far apart, selection
is where the accuracy comes from.

```{r true-pehe-plot, fig.height=10}
pehe_plot <- metrics %>%
  # remove really high values for a nicer looking plot
  filter(true_pehe < 1) %>%
  ggplot(aes(x = n, y = true_pehe, colour = model)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_y") +
  #scale_y_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
  labs(title = "True PEHE of each candidate",
       subtitle = "The values that metrics compared are trying to emulate",
       y = "True PEHE", x = "Sample size", colour = "Model")

pehe_plot
```

```{r winner-plot, fig.height=9}
# how often each candidate is genuinely the best one in its run
winners <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  slice_min(true_pehe, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  count(scenario, n, model, name = "wins", .drop = FALSE) %>%
  group_by(scenario, n) %>%
  mutate(prop = wins / sum(wins)) %>%
  ungroup()

winner_plot <- winners %>%
  ggplot(aes(x = n, y = prop, fill = model)) +
  geom_col() +
  facet_wrap(~scenario) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(colour = "black")) +
  labs(title = "How often each candidate is the true best",
       subtitle = "Proportion of runs in which this candidate has the lowest true PEHE",
       y = "Proportion of runs", x = "Candidate")

winner_plot
```

```{r spread-table}
spread_tab <- selection %>%
  distinct(scenario, n, run, spread, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(mean_spread = mean(spread, na.rm = T),
            mean_random_regret = mean(random_regret, na.rm = T),
            .groups = "drop") %>%
  left_join(select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
            by = c("scenario", "n"))

kable(spread_tab, digits = 4,
      col.names = c("Scenario", "n", "Worst - best true PEHE",
                    "Regret of a random pick", "Best fixed choice",
                    "Its regret"),
      caption = "How much selection is worth, per cell")
```

`Worst - best true PEHE` is the full range on the table (max - min); `regret of a random
pick` is what ignoring the scores altogether costs on average (mean - min). Any proxy has
to beat the latter to be doing anything at all. The `best fixed choice` is based
on the true PEHE.

## Rank agreement

Spearman and Kendall correlation between each proxy score and true PEHE,
computed across the 9 candidates within a run. A proxy that is useful for
selection sits well above 0.

```{r rho-plot, fig.height=9}
rho_box <- box_plot(selection, "rho",
                    "Spearman correlation with true PEHE",
                    "Spearman correlation", hline = NULL)

rho_sum <- summary_plot(selection_summary, "mean_rho", "mcse_rho",
                        "Mean Spearman correlation with true PEHE",
                        "Mean Spearman correlation", hline = 0)
# boxplot is quite hard to read just look at summary
#(rho_box / rho_sum) + plot_layout(guides = "collect")
rho_sum
```

```{r kendall-plot, fig.height=9}
summary_plot(selection_summary, "mean_kendall", "mcse_kendall",
             "Mean Kendall correlation with true PEHE",
             "Mean Kendall tau", hline = 0)
```

## Top-1 selection accuracy

The proportion of times a method selects the best performing model. Reference line
is the accuracy for picking the models at random.

```{r hit-plot, fig.height=9}
summary_plot(selection_summary, "mean_hit", "mcse_hit",
             "Top-1 selection accuracy",
             "Proportion of runs the proxy picks the true best",
             hline = 1 / length(model_levels))
```

Top-1 accuracy alone is harsh, especially when models have similar PEHEs. Let's 
look at true rankings across top selections.

```{r rank-dist-plot, fig.height=12}
rank_dist <- selection %>%
  count(scenario, n, proxy, pick_rank, name = "runs") %>%
  group_by(scenario, n, proxy) %>%
  mutate(prop = runs / sum(runs)) %>%
  ungroup()

# viridis rather than the house qualitative palette: pick_rank is ordinal, and
# rcartocolor::Safe would give 9 unordered hues to an ordered scale
rank_plot <- rank_dist %>%
  ggplot(aes(x = proxy, y = prop, fill = factor(pick_rank))) +
  geom_col() +
  facet_grid(rows = vars(scenario), cols = vars(n)) +
  scale_fill_viridis_d(direction = -1) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
  labs(title = "Where the selected candidate really ranked",
       subtitle = "1 = the proxy picked the true best, 9 = it picked the true worst",
       y = "Proportion of runs", x = "Proxy score",
       fill = "True rank of pick")

rank_plot
```

```{r mean-rank-plot, fig.height=9}
summary_plot(selection_summary, "mean_rank", "mcse_rank",
             "Mean true rank of the selected candidate",
             "Mean rank (1 = best of 9)",
             hline = (1 + length(model_levels)) / 2)
```

The dashed line is the mean rank you would get by picking at random.

## Regret

Regret is the excess true PEHE incurred by deploying the proxy's pick instead
of the true best candidate. The solid reference line in each panel is the regret
of picking at random; the dashed line is the best fixed choice.

```{r regret-plot, fig.height=16}
regret_refs <- list(
  geom_hline(data = random_baseline,
             aes(yintercept = random_regret, linetype = "Random pick"),
             colour = "grey30"),
  geom_hline(data = best_fixed,
             aes(yintercept = best_fixed_regret, linetype = "Best fixed choice"),
             colour = "grey30"),
  scale_linetype_manual(
    name = "Reference",
    breaks = c("Random pick", "Best fixed choice"),
    values = c("Random pick" = "solid", "Best fixed choice" = "dashed")),
  guides(linetype = guide_legend(override.aes = list(colour = "grey30")))
)

regret_box <- box_plot(selection, "regret",
                       "Regret: excess true PEHE from following the proxy",
                       "Regret") +
  regret_refs

regret_sum <- sum_ns_plot(selection_summary, "mean_regret", "mcse_regret",
                           "Mean regret", "Mean regret") +
  regret_refs

#(regret_box / regret_sum) + plot_layout(guides = "collect")
regret_sum
```

Relative regret puts the same quantity on a scale that is comparable across
scenarios and sample sizes: excess PEHE as a fraction of the best achievable
PEHE.

```{r rel-regret-plot, fig.height=9}
summary_plot(selection_summary, "mean_rel_regret", "mcse_rel_regret",
             "Mean relative regret",
             "Excess PEHE as a fraction of the best achievable", hline = 0)
```

## Which axis matters?

The 8 proxies are a 2 x 2 x 2 design. Reading a winner off an 8-level axis
conflates the three questions it actually answers — influence-corrected versus
DR risk, honest crossfitting versus resubstitution, XGBoost versus AutoML — so
here each is marginalised over the other two.

```{r axis-marginals}
axis_long <- selection %>%
  mutate(across(c(score_type, fold_type, pipeline), as.character)) %>%
  pivot_longer(c(score_type, fold_type, pipeline),
               names_to = "axis", values_to = "level") %>%
  mutate(axis = factor(axis_labels[axis], levels = unname(axis_labels)),
         level = factor(level_labels[level], levels = unname(level_labels)))

axis_summary <- axis_long %>%
  group_by(scenario, n, axis, level) %>%
  summarise(
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )

axis_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = level, y = est, colour = n, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(axis), scales = "free_x") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text = element_text(colour = "black")) +
    labs(title = title, y = ylab, x = NULL, colour = "n")
}
```

```{r axis-accuracy-plot, fig.height=16}
hit_axis <- axis_plot(axis_summary, "mean_hit", "mcse_hit",
                      "Top-1 accuracy by design axis",
                      "Mean top-1 accuracy", hline = 1 / length(model_levels))

kendall_axis <- axis_plot(axis_summary, "mean_kendall", "mcse_kendall",
                          "Rank agreement by design axis",
                          "Mean Kendall tau", hline = 1)

(hit_axis / kendall_axis) + plot_layout(guides = "collect")
```

```{r axis-regret-plot, fig.height=9}
axis_plot(axis_summary, "mean_regret", "mcse_regret",
          "Regret by design axis", "Mean regret")
```

## Do the proxies track PEHE in level, not just rank?

Ranking is all selection needs, but a proxy that tracks PEHE in *level* would
also tell you how good the winner is. The scores cannot be compared to PEHE
raw: `calc_infl_score` and `calc_dr_risk` estimate PEHE only up to a
**run-specific additive constant** — the DR risk carries the irreducible
variance of the pseudo-outcome `phi` about `tau`, and the influence score
carries `E[tau^2]` terms that depend on the realised sample, not on the
candidate being scored. Pooling raw scores across runs would produce a
correlation driven entirely by that offset and would say nothing about the
candidates.

Centering within each run tells us how each candidate deviates from its run's
average, in the proxy and in the truth.

```{r calibration-plot, fig.height=16}
centred <- long_ok %>%
  group_by(scenario, n, run, proxy) %>%
  mutate(score_c = score - mean(score),
         pehe_c = true_pehe - mean(true_pehe)) %>%
  ungroup()

calibration_plot <- centred %>%
  # remove large PEHEs to make plots clearer
  filter(true_pehe < 2) %>%
  ggplot(aes(x = pehe_c, y = score_c, colour = scenario)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(size = 0.5, alpha = 0.15) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  facet_wrap(~proxy, ncol = 2, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(strip.background = element_rect(fill = "white"),
        strip.text = element_text(colour = "black"),
        aspect.ratio = 1) +
  labs(title = "Proxy score against true PEHE, both centred within run",
       subtitle = "Centring is required: the proxies estimate PEHE only up to a run-specific constant",
       y = "Centred proxy score", x = "Centred true PEHE", colour = "Scenario")

calibration_plot
```

## Headline table

Per (scenario, n, proxy), arranged so the lowest-regret proxy comes first
within each cell.

```{r headline-table}
headline <- selection_summary %>%
  select(scenario, n, proxy, runs, mean_kendall, mean_rho, mean_hit,
         mean_rank, mean_regret, mean_rel_regret) %>%
  arrange(scenario, n, mean_regret)

kable(headline, digits = 4,
      col.names = c("Scenario", "n", "Proxy", "Runs", "Mean Kendall tau",
                    "Mean Spearman", "Top-1 accuracy", "Mean rank of pick",
                    "Mean regret", "Mean relative regret"))
```

For reference, the two baselines the regret columns should be read against:

```{r baseline-table}
baseline_tab <- random_baseline %>%
  left_join(select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
            by = c("scenario", "n"))

kable(baseline_tab, digits = 4,
      col.names = c("Scenario", "n", "Regret of a random pick",
                    "Best fixed choice", "Its regret"))
```
