---
title: "Model evaluation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 11,
  fig.height = 8
)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Which model evaluation metrics work well in RCT settings?

In real life we cannot evaluate models with true PEHE because the true CATEs are
never observed. Different surrogate metrics designs to select the best CATE model
among several have been designed. Typical application has been in large observational
datasets and rely on ML machinery, so we want to know how well these techniques
will work in RCT settings where sample size is much smaller.
A proxy is useful if, run by run, its argmin is the
same candidate true PEHE's argmin is; and if when it differs, it doesn't
differ expensively.

### Design

Four scenarios crossed with three sample sizes, 30 replicates each — **360
array jobs**. Scenarios are indices into `R/dgm_scenarios.R`'s
`set = "continuous"`; the subset 1 / 4 / 6 / 9 and its display labels are the
same ones `crossfitting/cf_results.qmd` uses, so the two reports are directly
comparable.

| | |
|---|---|
| scenarios | 1 (Null), 4 (Simple), 6 (Interaction), 9 (Non-linear) |
| n | 250, 500, 1000 |
| runs | 30 |
| folds | V = 10, single crossfitting, all n |


### The 9 candidates

Each is fit as its own single-crossfit DR-learner (`me_models.R`). `p` is the
number of covariates, 7–9 across these scenarios.

| id | family | hyperparameters |
|---|---|---|
| `rf1` | random forest | ranger defaults (`mtry = floor(sqrt(p))`) |
| `rf2` | random forest | `mtry = p, max.depth = 5` |
| `rf3` | random forest | `mtry = ceiling(p/2), max.depth = 3` |
| `net1` | elastic net | `alpha = 1` (lasso) |
| `net2` | elastic net | `alpha = 0` (ridge) |
| `net3` | elastic net | `alpha = 0.5` |
| `SL1` | SuperLearner | `glmnet, ranger, earth, gam, mean` |
| `SL2` | SuperLearner | `glmnet, xgboost, cforest, earth, gam, mean` |
| `SL3` | SuperLearner | `svm, nnet, mean` |

### The proxy scores

A second, *independent* nuisance pipeline estimates `mu_DR` and `pi`, and is
used only to score the candidates — never to fit them. Four axes:

| axis | levels |
|---|---|
| pipeline | **XGBoost** (36-combination CV grid) / **H2O AutoML** (up to 20 auto-tuned models) |
| fold regime | **cv** (leave-one-fold-out) / **whole** (no split) |
| score family | **infl** — influence-corrected PEHE proxy (`calc_infl_score`) / **dr** — DR risk, MSE against the AIPW pseudo-outcome `phi` (`calc_dr_risk`) / **cal** — DR calibration (`calc_cal_score`), at K = 5 and K = 10 quantile groups |
| propensity | **estimated** (`pi` from the pipeline) / **fixed at 0.5** (score names carrying `05`) |

Every one of them, and `true_pehe` itself, is a **loss: lower is better**.

**The calibration score** splits the evaluation rows into K quantile groups of
the *candidate's own* `tau_hat`, then compares the group average treatment
effect the candidate implies against the one the DR scores imply:
`M = Σ_k |G_k| · |GATE_k^hat − GATE_k^DR|`. It is weighted by group *counts*,
so it scales with `n` — constant within a run, so it moves no ranking here, but
its raw magnitude is not comparable across sample sizes.

Unlike `infl` and `dr`, **the calibration score is not an estimate of PEHE**,
and it is worth being explicit about what it cannot see. It measures
calibration, not discrimination: a candidate that predicts the *constant* ATE
for every unit has arbitrary quantile groups, so each group's `GATE^DR` is
roughly the overall mean and every discrepancy is near zero — a near-perfect
calibration score, at whatever PEHE the true heterogeneity implies. So expect
the shrunk-to-a-constant candidates (`net2` especially, and everything in
scenario 1) to do *well* here, and read this score as a complement to the PEHE
proxies rather than a substitute. Whether that failure mode actually bites at
these sample sizes is one of the things the study now measures.

**`whole` is resubstitution.** It fits and predicts the scoring nuisances on
the entire sample, so they see the same rows the candidate's `tau_hat` was fit
on. `cv` is the honest regime. This was done to see if there is a tradeoff between
data leakage and more samples for AutoML/XGB accuracy.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "model_evaluation")

metrics <- readRDS(file.path(res_path, "me_metrics.RDS"))
# add for now: remove calibration scores as they're not finished
# also remove sample size 1000 because it's too much info
metrics <- metrics %>%
  select(!all_of(matches("calq|cal05q"))) %>%
  filter(n != 1000)

# the proxy columns are whatever me_metrics.R produced that isn't a key or the
# truth - derived rather than hardcoded so that re-adding a fold regime (the
# `infold` one me_nuisance.R removed) needs no edit here
key_cols <- c("scenario", "n", "run", "model")
score_cols <- setdiff(names(metrics), c(key_cols, "true_pehe"))

# labels. Scenario labels match crossfitting/cf_results.qmd, which reports the
# same 1/4/6/9 subset.
scenario_labels <- c(
  `1` = "Null",
  `4` = "Simple",
  `6` = "Interaction",
  `9` = "Non-linear"
)

model_levels <- c(
  "rf1",
  "rf2",
  "rf3",
  "net1",
  "net2",
  "net3",
  "SL1",
  "SL2",
  "SL3"
)

family_labels <- c(
  rf = "Random forest",
  net = "Elastic net",
  SL = "SuperLearner"
)

# ---- parsing a score column name -------------------------------------------
# me_metrics.R builds every score column as <score_type>_<fold_type>_<pipeline>,
# where score_type is a single token of the form <family>[05][q<K>]:
#
#   infl_cv_xgb           infl05_cv_xgb        (05 = propensity fixed at 0.5)
#   dr_cv_shared_automl   dr05_cv_shared_automl
#   calq5_whole_xgb       cal05q10_holdout_automl
#
# The fold type is EVERYTHING BETWEEN the first and last token, not just the
# first of them - the strategies tree's arms are cv_indep, cv_shared and
# holdout, so a regex that takes one token reads `dr_cv_shared_xgb` as fold
# "cv" and silently merges two arms that the whole study exists to separate.
parse_proxy <- function(proxy_id) {
  parts <- strsplit(proxy_id, "_", fixed = TRUE)
  tibble(
    proxy_id = proxy_id,
    score_type = vapply(parts, function(p) p[1], character(1)),
    fold_type = vapply(
      parts,
      function(p) paste(p[-c(1, length(p))], collapse = "_"),
      character(1)
    ),
    pipeline = vapply(parts, function(p) p[length(p)], character(1))
  ) %>%
    mutate(
      # the family with the propensity marker and group count stripped off.
      # `score_family`, not `family` - the metrics tibble already uses that
      # name for the candidate's own family (rf / net / SL), and the two would
      # collide into family.x/family.y on the join in the long-form chunk.
      score_family = sub("^([a-z]+?)(05)?(q[0-9]+)?$", "\\1", score_type),
      # "05" appears in a score type only when the propensity is fixed - calq10
      # carries a "10", never an "05"
      pi_regime = ifelse(grepl("05", score_type), "fixed", "estimated"),
      k_groups = suppressWarnings(as.integer(sub("^.*q", "", score_type))),
      k_groups = ifelse(grepl("q[0-9]+$", score_type), k_groups, NA_integer_)
    )
}

axis_labels <- c(
  score_type = "Score type",
  score_family = "Score family",
  fold_type = "Fold regime",
  pipeline = "Nuisance pipeline",
  pi_regime = "Propensity",
  k_groups = "Quantile groups"
)

level_labels <- c(
  infl = "Influence",
  dr = "DR risk",
  cal = "Calibration",
  cv = "CV",
  cv_indep = "CV (indep.)",
  cv_shared = "CV (shared)",
  holdout = "Holdout",
  whole = "Whole",
  split = "80:20",
  xgb = "XGBoost",
  automl = "AutoML",
  estimated = "pi estimated",
  fixed = "pi = 0.5"
)

# Display labels are BUILT from the parsed axes rather than typed out: the main
# tree now carries 33 score columns and the strategies tree 65, so a hand-kept
# lookup would be both unreadable and a silent source of NA labels the first
# time a score family is added.
label_or_id <- function(x) ifelse(is.na(level_labels[x]), x, level_labels[x])

proxy_label_of <- function(parsed) {
  paste(
    ifelse(
      is.na(parsed$k_groups),
      label_or_id(parsed$score_family),
      paste0(label_or_id(parsed$score_family), " K=", parsed$k_groups)
    ),
    ifelse(parsed$pi_regime == "fixed", "[pi=.5]", "[pi-hat]"),
    label_or_id(parsed$fold_type),
    label_or_id(parsed$pipeline),
    sep = " / "
  )
}

# one row per score column, in the order me_metrics.R emitted them, which is
# also the order the factor levels below take
proxy_key <- parse_proxy(score_cols) %>%
  mutate(proxy_label = proxy_label_of(.))

proxy_labels <- setNames(proxy_key$proxy_label, proxy_key$proxy_id)

metrics <- metrics %>%
  mutate(
    scenario = factor(
      scenario_labels[as.character(scenario)],
      levels = unname(scenario_labels)
    ),
    n = factor(n, levels = sort(unique(n))),
    model = factor(model, levels = model_levels),
    family = factor(
      family_labels[sub("[0-9]+$", "", as.character(model))],
      levels = unname(family_labels)
    )
    # maybe adding version: 1,2,3 would make plots nicer?
  )
```

```{r long-form}
# one row per (scenario, n, run, model, proxy). The three design axes are
# recovered from the column name, which me_metrics.R builds as
# <score_type>_<fold_type>_<pipeline>.
long <- metrics %>%
  pivot_longer(
    all_of(score_cols),
    names_to = "proxy_id",
    values_to = "score"
  ) %>%
  left_join(proxy_key, by = "proxy_id") %>%
  mutate(proxy = factor(proxy_label, levels = unname(proxy_labels)))

# a candidate only enters a ranking if both its score and its truth are there
long_ok <- filter(long, !is.na(score), !is.na(true_pehe))
```

```{r selection}
# The core derivation: one row per (scenario, n, run, proxy), summarising how
# that proxy ranked the 9 candidates against how true PEHE ranked them.
#
# `regret` is measured against the best candidate the proxy could actually have
# picked (the minimum over the candidates present in that run), not against a
# hypothetical complete set - so a run with a dropped candidate is scored
# honestly rather than penalised for the gap.
selection <- long_ok %>%
  group_by(
    scenario,
    n,
    run,
    proxy,
    proxy_id,
    score_type,
    score_family,
    pi_regime,
    k_groups,
    fold_type,
    pipeline
  ) %>%
  filter(n() >= 2) %>%
  summarise(
    n_candidates = n(),
    rho = cor(score, true_pehe, method = "spearman"),
    kendall = cor(score, true_pehe, method = "kendall"),
    pick = model[which.min(score)],
    best = model[which.min(true_pehe)],
    hit = as.numeric(pick == best),
    pick_rank = rank(true_pehe, ties.method = "min")[which.min(score)],
    regret = true_pehe[which.min(score)] - min(true_pehe),
    rel_regret = regret / min(true_pehe),
    spread = max(true_pehe) - min(true_pehe),
    # expected cost of ignoring the scores and picking one of the candidates
    # uniformly at random - the floor any useful proxy has to beat
    random_regret = mean(true_pehe) - min(true_pehe),
    .groups = "drop"
  )

# per (scenario, n, proxy) summaries, mean +/- Monte Carlo SE over the runs
selection_summary <- selection %>%
  group_by(
    scenario,
    n,
    proxy,
    score_type,
    score_family,
    pi_regime,
    k_groups,
    fold_type,
    pipeline
  ) %>%
  summarise(
    runs = n(),
    mean_rho = mean(rho, na.rm = T),
    mcse_rho = sd(rho, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_rank = mean(pick_rank, na.rm = T),
    mcse_rank = sd(pick_rank, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    mean_rel_regret = mean(rel_regret, na.rm = T),
    mcse_rel_regret = sd(rel_regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )
```

```{r baselines}
# Two reference points, both on the regret scale, both per (scenario, n).
#
# random pick: the expected regret of ignoring the scores entirely.
random_baseline <- selection %>%
  distinct(scenario, n, run, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(random_regret = mean(random_regret, na.rm = T), .groups = "drop")

# best fixed choice: commit to one candidate a priori and always use it. Note
# this is picked WITH hindsight - the best of the nine is chosen using the
# truth - so it is an optimistic upper bound on what a fixed rule could really
# achieve. A proxy that cannot beat it has not earned its cost.
oracle_by_run <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  summarise(oracle_pehe = min(true_pehe), .groups = "drop")

fixed_choice <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  left_join(oracle_by_run, by = c("scenario", "n", "run")) %>%
  mutate(fixed_regret = true_pehe - oracle_pehe) %>%
  group_by(scenario, n, model) %>%
  summarise(
    mean_regret = mean(fixed_regret, na.rm = T),
    mcse_regret = sd(fixed_regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )

best_fixed <- fixed_choice %>%
  group_by(scenario, n) %>%
  slice_min(mean_regret, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  rename(best_fixed_model = model, best_fixed_regret = mean_regret)
```

```{r plot-helpers, include=FALSE}
# The house pair of plots - per-run distribution and mean with MCSE error bars.
# The sweep (the 8 proxies) goes on x and the design on the facets, following
# confidence_intervals/continuous/cts_ci_results.qmd rather than
# cts_results.qmd's facet_wrap(~scenario).

box_plot <- function(df, metric, title, ylab, hline = NULL) {
  p <- ggplot(df, aes(x = proxy, y = .data[[metric]], colour = score_type))
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d(
      "rcartocolor::Safe",
      labels = c(dr = "DR risk", infl = "Influence")
    ) +
    theme_light() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black")
    ) +
    labs(title = title, y = ylab, x = "Proxy score", colour = "Score type")
}

summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(
      est = .data[[mean_col]],
      lo = .data[[mean_col]] - .data[[mcse_col]],
      hi = .data[[mean_col]] + .data[[mcse_col]]
    )
  p <- ggplot(
    df,
    aes(
      x = n,
      y = est,
      colour = score_type,
      shape = pipeline,
      ymin = lo,
      ymax = hi
    )
  )
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(
      position = position_dodge(width = 0.5),
      linewidth = 0.3,
      width = 0.3
    ) +
    facet_grid(
      rows = vars(scenario),
      cols = vars(fold_type),
      scales = "free_y"
    ) +
    scale_colour_paletteer_d(
      "rcartocolor::Safe",
      labels = c(dr = "DR risk", infl = "Influence")
    ) +
    theme_light() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black")
    ) +
    labs(
      title = title,
      y = ylab,
      x = "sample size",
      colour = "Score type",
      shape = "Machinery"
    )
}

sum_ns_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(
      est = .data[[mean_col]],
      lo = .data[[mean_col]] - .data[[mcse_col]],
      hi = .data[[mean_col]] + .data[[mcse_col]]
    )
  p <- ggplot(
    df,
    aes(
      x = fold_type,
      y = est,
      colour = score_type,
      shape = pipeline,
      ymin = lo,
      ymax = hi
    )
  )
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(
      position = position_dodge(width = 0.5),
      linewidth = 0.3,
      width = 0.3
    ) +
    facet_grid(rows = vars(n), cols = vars(scenario), scales = "free_y") +
    scale_colour_paletteer_d(
      "rcartocolor::Safe",
      labels = c(dr = "DR risk", infl = "Influence")
    ) +
    theme_light() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black")
    ) +
    labs(
      title = title,
      y = ylab,
      x = "Fold type",
      colour = "Score type",
      shape = "Machinery"
    )
}
```

## Data completeness

The study was still `in_progress` at the last campaign check.
Missing runs shrink the number of rankings each cell is averaged over; missing
*scores* within a present run shrink the number of candidates a proxy ranked.
2 runs in the null $n=250$ configuration have not run, currently troubleshooting.

```{r completeness-runs}
runs_present <- metrics %>%
  distinct(scenario, n, run) %>%
  count(scenario, n, name = "runs") %>%
  pivot_wider(names_from = n, values_from = runs, names_prefix = "n = ")

kable(runs_present, caption = "Replicates present per cell (30 expected)")
```

```{r completeness-na}
na_counts <- metrics %>%
  summarise(across(all_of(c("true_pehe", score_cols)), ~ sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  mutate(pct = 100 * n_missing / nrow(metrics))

kable(
  na_counts,
  digits = 2,
  col.names = c("Score column", "Missing values", "% of rows"),
  caption = "Missing scores across all (scenario, n, run, candidate) rows"
)
```

```{r completeness-candidates}
short_rankings <- selection %>%
  filter(n_candidates < length(model_levels)) %>%
  count(scenario, n, proxy, n_candidates, name = "rankings")

if (nrow(short_rankings) == 0) {
  cat("Every ranking scored all", length(model_levels), "candidates.\n")
} else {
  kable(
    short_rankings,
    col.names = c("Scenario", "n", "Proxy", "Candidates ranked", "Rankings"),
    caption = "Rankings built from fewer than the full candidate set"
  )
}
```

## Examining True PEHE

If all 9 candidates achieve near-identical true PEHE then picking the wrong one
costs nothing and a high error rate is harmless; if they are far apart, selection
is where the accuracy comes from.

```{r true-pehe-plot, fig.height=10}
pehe_plot <- metrics %>%
  # remove really high values for a nicer looking plot
  filter(true_pehe < 1) %>%
  ggplot(aes(x = n, y = true_pehe, colour = model)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_wrap(~scenario, scales = "free_y") +
  #scale_y_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "True PEHE of each candidate",
    subtitle = "The values that metrics compared are trying to emulate",
    y = "True PEHE",
    x = "Sample size",
    colour = "Model"
  )

pehe_plot
```

```{r winner-plot, fig.height=9}
# how often each candidate is genuinely the best one in its run
winners <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  slice_min(true_pehe, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  count(scenario, n, model, name = "wins", .drop = FALSE) %>%
  group_by(scenario, n) %>%
  mutate(prop = wins / sum(wins)) %>%
  ungroup()

winner_plot <- winners %>%
  ggplot(aes(x = n, y = prop, fill = model)) +
  geom_col() +
  facet_wrap(~scenario) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black")
  ) +
  labs(
    title = "How often each candidate is the true best",
    subtitle = "Proportion of runs in which this candidate has the lowest true PEHE",
    y = "Proportion of runs",
    x = "Candidate"
  )

winner_plot
```

```{r spread-table}
spread_tab <- selection %>%
  distinct(scenario, n, run, spread, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(
    mean_spread = mean(spread, na.rm = T),
    mean_random_regret = mean(random_regret, na.rm = T),
    .groups = "drop"
  ) %>%
  left_join(
    select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
    by = c("scenario", "n")
  )

kable(
  spread_tab,
  digits = 4,
  col.names = c(
    "Scenario",
    "n",
    "Worst - best true PEHE",
    "Regret of a random pick",
    "Best fixed choice",
    "Its regret"
  ),
  caption = "How much selection is worth, per cell"
)
```

`Worst - best true PEHE` is the full range on the table (max - min); `regret of a random
pick` is what ignoring the scores altogether costs on average (mean - min). Any proxy has
to beat the latter to be doing anything at all. The `best fixed choice` is based
on the true PEHE.

## Rank agreement

Spearman and Kendall correlation between each proxy score and true PEHE,
computed across the 9 candidates within a run. A proxy that is useful for
selection sits well above 0.

```{r rho-plot, fig.height=9}
rho_box <- box_plot(
  selection,
  "rho",
  "Spearman correlation with true PEHE",
  "Spearman correlation",
  hline = NULL
)

rho_sum <- sum_ns_plot(
  selection_summary,
  "mean_rho",
  "mcse_rho",
  "Mean Spearman correlation with true PEHE",
  "Mean Spearman correlation",
  hline = 0
)
# boxplot is quite hard to read just look at summary
#(rho_box / rho_sum) + plot_layout(guides = "collect")
rho_sum
```

```{r kendall-plot, fig.height=9}
sum_ns_plot(
  selection_summary,
  "mean_kendall",
  "mcse_kendall",
  "Mean Kendall correlation with true PEHE",
  "Mean Kendall tau",
  hline = 0
)
```

## Top-1 selection accuracy

The proportion of times a method selects the best performing model. Reference line
is the accuracy for picking the models at random.

```{r hit-plot, fig.height=9}
sum_ns_plot(
  selection_summary,
  "mean_hit",
  "mcse_hit",
  "Top-1 selection accuracy",
  "Proportion of runs the proxy picks the true best",
  hline = 1 / length(model_levels)
)
```

Top-1 accuracy alone is harsh, especially when models have similar PEHEs. Let's 
look at true rankings across top selections.

```{r rank-dist-plot, fig.height=12}
# strip the pipeline component out of the proxy label now that pipeline is a
# facet column, not a bar-by-bar distinction — label_or_id()/level_labels
# come from the score-column parsing chunk above; selection already carries
# score_family/pi_regime/k_groups/fold_type per row so no re-parsing needed
rank_dist <- selection %>%
  mutate(
    proxy_sub = paste(
      ifelse(
        is.na(k_groups),
        label_or_id(score_family),
        paste0(label_or_id(score_family), " K=", k_groups)
      ),
      ifelse(pi_regime == "fixed", "[pi=.5]", "[pi-hat]"),
      label_or_id(fold_type),
      sep = " / "
    )
  ) %>%
  count(scenario, n, pipeline, proxy_sub, pick_rank, name = "runs") %>%
  group_by(scenario, n, pipeline, proxy_sub) %>%
  mutate(prop = runs / sum(runs)) %>%
  ungroup()

# viridis rather than the house qualitative palette: pick_rank is ordinal, and
# rcartocolor::Safe would give 9 unordered hues to an ordered scale
#
# One plot per scenario; within each, rows = n, cols = pipeline. scales/space
# = "free_x" so a panel only reserves x-slots for the proxy_sub combos that
# actually occur in it, instead of the union across every panel - that union
# (all fold regimes x families x pi-regimes, most of them belonging to the
# other pipeline) was the source of the blank space.
rank_plots <- rank_dist %>%
  split(.$scenario) %>%
  imap(function(df, scen) {
    ggplot(df, aes(x = proxy_sub, y = prop, fill = factor(pick_rank))) +
      geom_col() +
      facet_grid(
        rows = vars(n),
        cols = vars(pipeline),
        scales = "free_x",
        space = "free_x",
        labeller = labeller(n = label_both)
      ) +
      scale_fill_viridis_d(direction = -1) +
      theme_light() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(colour = "black")
      ) +
      labs(
        title = scen,
        y = "Proportion of runs",
        x = "Proxy score",
        fill = "True rank of pick"
      )
  })

wrap_plots(rank_plots, ncol = 2) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "True rankings of selected candidates",
    subtitle = "1 (best) - 9 (worst)"
  ) &
  theme(legend.position = "bottom")
```

```{r mean-rank-plot, fig.height=9}
sum_ns_plot(
  selection_summary,
  "mean_rank",
  "mcse_rank",
  "Mean true rank of the selected candidate",
  "Mean rank (1 = best of 9)",
  hline = (1 + length(model_levels)) / 2
)
```

The dashed line is the mean rank you would get by picking at random.

## Regret

Regret is the excess true PEHE incurred by deploying the proxy's pick instead
of the true best candidate. The solid reference line in each panel is the regret
of picking at random; the dashed line is the best fixed choice.

```{r regret-plot, fig.height=9}
regret_refs <- list(
  geom_hline(
    data = random_baseline,
    aes(yintercept = random_regret, linetype = "Random pick"),
    colour = "grey30"
  ),
  geom_hline(
    data = best_fixed,
    aes(yintercept = best_fixed_regret, linetype = "Best fixed choice"),
    colour = "grey30"
  ),
  scale_linetype_manual(
    name = "Reference",
    breaks = c("Random pick", "Best fixed choice"),
    values = c("Random pick" = "solid", "Best fixed choice" = "dashed")
  ),
  guides(linetype = guide_legend(override.aes = list(colour = "grey30")))
)

regret_box <- box_plot(
  selection,
  "regret",
  "Regret: excess true PEHE from following the proxy",
  "Regret"
) +
  regret_refs

regret_sum <- sum_ns_plot(
  selection_summary,
  "mean_regret",
  "mcse_regret",
  "Mean regret",
  "Mean regret"
) +
  regret_refs

#(regret_box / regret_sum) + plot_layout(guides = "collect")
regret_sum
```

Relative regret puts the same quantity on a scale that is comparable across
scenarios and sample sizes: excess PEHE as a fraction of the best achievable
PEHE.

```{r rel-regret-plot, fig.height=9}
sum_ns_plot(
  selection_summary,
  "mean_rel_regret",
  "mcse_rel_regret",
  "Mean relative regret",
  "Excess PEHE as a fraction of the best achievable",
  hline = 0
)
```

## Which axis matters?

The proxies are a factorial design, not a flat list. Reading a winner off the
full proxy axis conflates the questions it actually answers — influence-
corrected versus DR risk versus calibration, an estimated propensity versus the
true 0.5, honest crossfitting versus resubstitution, XGBoost versus AutoML — so
here each is marginalised over the others.

`score_family` is used rather than `score_type` because the latter is the
composite token (`cal05q10` is *calibration*, *fixed pi*, *K=10* in one
string); the propensity is pulled out as its own axis. `k_groups` is
deliberately **not** marginalised here: it is defined only for the calibration
columns, so a panel averaging over it would put two thirds of the rows in an
"undefined" level. Compare K = 5 against K = 10 within the calibration family
instead.

```{r axis-marginals}
axis_long <- selection %>%
  mutate(across(
    c(score_family, pi_regime, fold_type, pipeline),
    as.character
  )) %>%
  pivot_longer(
    c(score_family, pi_regime, fold_type, pipeline),
    names_to = "axis",
    values_to = "level"
  ) %>%
  mutate(
    axis = factor(axis_labels[axis], levels = unname(axis_labels)),
    level = factor(level_labels[level], levels = unname(level_labels))
  )

axis_summary <- axis_long %>%
  group_by(scenario, n, axis, level) %>%
  summarise(
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )

axis_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(
      est = .data[[mean_col]],
      lo = .data[[mean_col]] - .data[[mcse_col]],
      hi = .data[[mean_col]] + .data[[mcse_col]]
    )
  p <- ggplot(df, aes(x = level, y = est, colour = n, ymin = lo, ymax = hi))
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(
      position = position_dodge(width = 0.5),
      linewidth = 0.3,
      width = 0.3
    ) +
    facet_grid(rows = vars(scenario), cols = vars(axis), scales = "free_x") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_light() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(colour = "black")
    ) +
    labs(title = title, y = ylab, x = NULL, colour = "n")
}
```

```{r axis-accuracy-plot, fig.height=16}
hit_axis <- axis_plot(
  axis_summary,
  "mean_hit",
  "mcse_hit",
  "Top-1 accuracy by design axis",
  "Mean top-1 accuracy",
  hline = 1 / length(model_levels)
)

kendall_axis <- axis_plot(
  axis_summary,
  "mean_kendall",
  "mcse_kendall",
  "Rank agreement by design axis",
  "Mean Kendall tau",
  hline = 1
)

(hit_axis / kendall_axis) + plot_layout(guides = "collect")
```

```{r axis-regret-plot, fig.height=9}
axis_plot(
  axis_summary,
  "mean_regret",
  "mcse_regret",
  "Regret by design axis",
  "Mean regret"
)
```

## Do the proxies track PEHE in level, not just rank?

(*"Calibration" here means tracking PEHE in level — not the `cal*` score
family, which is a different question entirely.*)

Ranking is all selection needs, but a proxy that tracks PEHE in *level* would
also tell you how good the winner is. The scores cannot be compared to PEHE
raw: `calc_infl_score` and `calc_dr_risk` estimate PEHE only up to a
**run-specific additive constant** — the DR risk carries the irreducible
variance of the pseudo-outcome `phi` about `tau`, and the influence score
carries `E[tau^2]` terms that depend on the realised sample, not on the
candidate being scored. Pooling raw scores across runs would produce a
correlation driven entirely by that offset and would say nothing about the
candidates.

The `cal*` columns are a different matter again: they are not an estimate of
PEHE up to a constant or otherwise, so centring does not put them on PEHE's
scale and the panels below should be read for *rank* agreement only.

Centering within each run tells us how each candidate deviates from its run's
average, in the proxy and in the truth.

```{r calibration-plot, fig.height=12}
centred <- long_ok %>%
  group_by(scenario, n, run, proxy) %>%
  mutate(
    score_c = score - mean(score),
    pehe_c = true_pehe - mean(true_pehe)
  ) %>%
  ungroup()

calibration_plot <- centred %>%
  # remove large PEHEs to make plots clearer
  filter(true_pehe < 2) %>%
  ggplot(aes(x = pehe_c, y = score_c, colour = scenario)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(size = 0.5, alpha = 0.15) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  facet_wrap(~proxy, ncol = 4, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(colour = "black"),
    aspect.ratio = 1
  ) +
  labs(
    title = "Proxy score against true PEHE, both centred within run",
    subtitle = "Centring is required: the proxies estimate PEHE only up to a run-specific constant",
    y = "Centred proxy score",
    x = "Centred true PEHE",
    colour = "Scenario"
  )

calibration_plot
```

## Headline table

Per (scenario, n, proxy), arranged so the lowest-regret proxy comes first
within each cell.

```{r headline-table}
headline <- selection_summary %>%
  select(
    scenario,
    n,
    proxy,
    runs,
    mean_kendall,
    mean_rho,
    mean_hit,
    mean_rank,
    mean_regret,
    mean_rel_regret
  ) %>%
  arrange(scenario, n, mean_regret)

kable(
  headline,
  digits = 4,
  col.names = c(
    "Scenario",
    "n",
    "Proxy",
    "Runs",
    "Mean Kendall tau",
    "Mean Spearman",
    "Top-1 accuracy",
    "Mean rank of pick",
    "Mean regret",
    "Mean relative regret"
  )
)
```

For reference, the two baselines the regret columns should be read against:

```{r baseline-table}
baseline_tab <- random_baseline %>%
  left_join(
    select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
    by = c("scenario", "n")
  )

kable(
  baseline_tab,
  digits = 4,
  col.names = c(
    "Scenario",
    "n",
    "Regret of a random pick",
    "Best fixed choice",
    "Its regret"
  )
)
```
