---
title: "Model evaluation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## Can a cheap proxy loss pick the right CATE learner?

True PEHE — the mean squared error of an estimated CATE against the truth —
is the quantity you would select a CATE learner on if you could. You can't:
outside a simulation the true CATE is never observed. The practical fallback
is a *proxy* loss built from a second, independent set of nuisance estimates,
which is observable. This study asks whether that substitution survives:
**given 9 candidate CATE-learner configurations, do the cheap proxy losses
rank them the way the unobservable true PEHE would?**

That makes this a model-*selection* study, not another "does estimator X
recover the true CATE" study. The unit of analysis throughout is **the ranking
of the 9 candidates within a single (scenario, n, run)** — not any one
candidate's own accuracy. A proxy is useful if, run by run, its argmin is the
same candidate true PEHE's argmin is; and if when it differs, it doesn't
differ expensively.

### Design

Four scenarios crossed with three sample sizes, 30 replicates each — **360
array jobs**. Scenarios are indices into `R/dgm_scenarios.R`'s
`set = "continuous"`; the subset 1 / 4 / 6 / 9 and its display labels are the
same ones `crossfitting/cf_results.qmd` uses, so the two reports are directly
comparable.

| | |
|---|---|
| scenarios | 1 (Null), 4 (Simple), 6 (Interaction), 9 (Non-linear) |
| n | 250, 500, 1000 |
| runs | 30 |
| folds | V = 10, single crossfitting (`scf_scf`), all n |

All 9 candidates share one honesty regime — leave-one-fold-out nuisances with
the stage-2 model trained on the same folds. That is deliberate: mixing OOB
forests with crossfit everything-else would confound learner choice with
crossfitting scheme in the ranking, which is the thing being measured. See
`model_evaluation/README.md`.

### The 9 candidates

Each is fit as its own single-crossfit DR-learner (`me_models.R`). `p` is the
number of covariates, 7–9 across these scenarios.

| id | family | hyperparameters |
|---|---|---|
| `rf1` | random forest | ranger defaults (`mtry = floor(sqrt(p))`) |
| `rf2` | random forest | `mtry = p, max.depth = 5` |
| `rf3` | random forest | `mtry = ceiling(p/2), max.depth = 3` |
| `net1` | elastic net | `alpha = 1` (lasso) |
| `net2` | elastic net | `alpha = 0` (ridge) |
| `net3` | elastic net | `alpha = 0.5` |
| `SL1` | SuperLearner | `glmnet, ranger, earth, gam, mean` |
| `SL2` | SuperLearner | `glmnet, xgboost, cforest, earth, gam, mean` |
| `SL3` | SuperLearner | `svm, nnet, mean` |

### The 8 proxy scores

A second, *independent* nuisance pipeline estimates `mu0_T`, `mu1_T`, `mu_DR`
and `pi`, and is used only to score the candidates — never to fit them. Two
pipelines, two fold regimes, two score types:

| axis | levels |
|---|---|
| pipeline | **XGBoost** (36-combination CV grid) / **H2O AutoML** (up to 20 auto-tuned models) |
| fold regime | **cv** (leave-one-fold-out) / **whole** (no split) |
| score type | **infl** — influence-corrected PEHE proxy (`calc_infl_score`) / **dr** — DR risk, MSE against the AIPW pseudo-outcome `phi` (`calc_dr_risk`) |

All eight, and `true_pehe` itself, are **losses: lower is better**.

**`whole` is resubstitution.** It fits and predicts the scoring nuisances on
the entire sample, so they see the same rows the candidate's `tau_hat` was fit
on. `cv` is the honest regime. Any advantage `whole` shows here is the effect
this study exists to detect, not a free win — read it as a warning, not a
recommendation.

**Scenario 1 (Null) is not degenerate here.** `R/metrics.R::cate_metrics()`
forces its `corr` to 0 in scenario 1 because the true CATE is constant *across
units*, so a unit-level correlation is undefined. This report's correlations
are across the *9 candidates* within a run, and `true_pehe` still varies from
candidate to candidate even when the truth is a constant — so no scenario-1
guard is needed, and none is applied.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "model_evaluation")

metrics <- readRDS(file.path(res_path, "me_metrics.RDS"))

# the proxy columns are whatever me_metrics.R produced that isn't a key or the
# truth - derived rather than hardcoded so that re-adding a fold regime (the
# `infold` one me_nuisance.R removed) needs no edit here
key_cols <- c("scenario", "n", "run", "model")
score_cols <- setdiff(names(metrics), c(key_cols, "true_pehe"))

# labels. Scenario labels match crossfitting/cf_results.qmd, which reports the
# same 1/4/6/9 subset.
scenario_labels <- c(`1` = "Null", `4` = "Simple", `6` = "Interaction",
                     `9` = "Non-linear")

model_levels <- c("rf1", "rf2", "rf3", "net1", "net2", "net3",
                  "SL1", "SL2", "SL3")

family_labels <- c(rf = "Random forest", net = "Elastic net",
                   SL = "SuperLearner")

# proxy display labels, ordered score type first so the infl/dr split reads
# straight off the x axis
proxy_labels <- c(
  infl_cv_xgb       = "Infl / CV / XGB",
  infl_cv_automl    = "Infl / CV / AutoML",
  infl_whole_xgb    = "Infl / whole / XGB",
  infl_whole_automl = "Infl / whole / AutoML",
  dr_cv_xgb         = "DR / CV / XGB",
  dr_cv_automl      = "DR / CV / AutoML",
  dr_whole_xgb      = "DR / whole / XGB",
  dr_whole_automl   = "DR / whole / AutoML"
)

axis_labels <- c(score_type = "Score type", fold_type = "Fold regime",
                 pipeline = "Nuisance pipeline")

level_labels <- c(infl = "Influence", dr = "DR risk", cv = "CV",
                  whole = "Whole", xgb = "XGBoost", automl = "AutoML")

metrics <- metrics %>%
  mutate(
    scenario = factor(scenario_labels[as.character(scenario)],
                      levels = unname(scenario_labels)),
    n = factor(n, levels = sort(unique(n))),
    model = factor(model, levels = model_levels),
    family = factor(family_labels[sub("[0-9]+$", "", as.character(model))],
                    levels = unname(family_labels))
  )
```

```{r long-form}
# one row per (scenario, n, run, model, proxy). The three design axes are
# recovered from the column name, which me_metrics.R builds as
# <score_type>_<fold_type>_<pipeline>.
long <- metrics %>%
  pivot_longer(all_of(score_cols), names_to = "proxy_id", values_to = "score") %>%
  mutate(
    score_type = sub("_.*$", "", proxy_id),
    fold_type  = sub("^[^_]+_([^_]+)_.*$", "\\1", proxy_id),
    pipeline   = sub("^.*_", "", proxy_id),
    proxy = factor(proxy_labels[proxy_id], levels = unname(proxy_labels))
  )

# a candidate only enters a ranking if both its score and its truth are there
long_ok <- filter(long, !is.na(score), !is.na(true_pehe))
```

```{r selection}
# The core derivation: one row per (scenario, n, run, proxy), summarising how
# that proxy ranked the 9 candidates against how true PEHE ranked them.
#
# `regret` is measured against the best candidate the proxy could actually have
# picked (the minimum over the candidates present in that run), not against a
# hypothetical complete set - so a run with a dropped candidate is scored
# honestly rather than penalised for the gap.
selection <- long_ok %>%
  group_by(scenario, n, run, proxy, proxy_id, score_type, fold_type, pipeline) %>%
  filter(n() >= 2) %>%
  summarise(
    n_candidates = n(),
    rho          = cor(score, true_pehe, method = "spearman"),
    kendall      = cor(score, true_pehe, method = "kendall"),
    pick         = model[which.min(score)],
    best         = model[which.min(true_pehe)],
    hit          = as.numeric(pick == best),
    pick_rank    = rank(true_pehe, ties.method = "min")[which.min(score)],
    regret       = true_pehe[which.min(score)] - min(true_pehe),
    rel_regret   = regret / min(true_pehe),
    spread       = max(true_pehe) - min(true_pehe),
    # expected cost of ignoring the scores and picking one of the candidates
    # uniformly at random - the floor any useful proxy has to beat
    random_regret = mean(true_pehe) - min(true_pehe),
    .groups = "drop"
  )

# per (scenario, n, proxy) summaries, mean +/- Monte Carlo SE over the runs
selection_summary <- selection %>%
  group_by(scenario, n, proxy, score_type, fold_type, pipeline) %>%
  summarise(
    runs = n(),
    mean_rho = mean(rho, na.rm = T),
    mcse_rho = sd(rho, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_rank = mean(pick_rank, na.rm = T),
    mcse_rank = sd(pick_rank, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    mean_rel_regret = mean(rel_regret, na.rm = T),
    mcse_rel_regret = sd(rel_regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )
```

```{r baselines}
# Two reference points, both on the regret scale, both per (scenario, n).
#
# random pick: the expected regret of ignoring the scores entirely.
random_baseline <- selection %>%
  distinct(scenario, n, run, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(random_regret = mean(random_regret, na.rm = T), .groups = "drop")

# best fixed choice: commit to one candidate a priori and always use it. Note
# this is picked WITH hindsight - the best of the nine is chosen using the
# truth - so it is an optimistic upper bound on what a fixed rule could really
# achieve. A proxy that cannot beat it has not earned its cost.
oracle_by_run <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  summarise(oracle_pehe = min(true_pehe), .groups = "drop")

fixed_choice <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  left_join(oracle_by_run, by = c("scenario", "n", "run")) %>%
  mutate(fixed_regret = true_pehe - oracle_pehe) %>%
  group_by(scenario, n, model) %>%
  summarise(mean_regret = mean(fixed_regret, na.rm = T),
            mcse_regret = sd(fixed_regret, na.rm = T) / sqrt(n()),
            .groups = "drop")

best_fixed <- fixed_choice %>%
  group_by(scenario, n) %>%
  slice_min(mean_regret, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  rename(best_fixed_model = model, best_fixed_regret = mean_regret)
```

```{r plot-helpers, include=FALSE}
# The house pair of plots - per-run distribution and mean with MCSE error bars.
# The sweep (the 8 proxies) goes on x and the design on the facets, following
# confidence_intervals/continuous/cts_ci_results.qmd rather than
# cts_results.qmd's facet_wrap(~scenario).

box_plot <- function(df, metric, title, ylab, hline = NULL) {
  p <- ggplot(df, aes(x = proxy, y = .data[[metric]], colour = score_type))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe",
                             labels = c(dr = "DR risk", infl = "Influence")) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Proxy score", colour = "Score type")
}

summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = proxy, y = est, colour = score_type, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe",
                             labels = c(dr = "DR risk", infl = "Influence")) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Proxy score", colour = "Score type")
}
```

## Data completeness

The study was still `in_progress` at the last campaign check, and the H2O
AutoML pipeline is the one component that has never completed a full run — so
this is not boilerplate. Missing runs shrink the number of rankings each cell
is averaged over; missing *scores* within a present run shrink the number of
candidates a proxy ranked.

```{r completeness-runs}
runs_present <- metrics %>%
  distinct(scenario, n, run) %>%
  count(scenario, n, name = "runs") %>%
  pivot_wider(names_from = n, values_from = runs, names_prefix = "n = ")

kable(runs_present, caption = "Replicates present per cell (30 expected)")
```

```{r completeness-na}
na_counts <- metrics %>%
  summarise(across(all_of(c("true_pehe", score_cols)), ~ sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  mutate(pct = 100 * n_missing / nrow(metrics))

kable(na_counts, digits = 2,
      col.names = c("Score column", "Missing values", "% of rows"),
      caption = "Missing scores across all (scenario, n, run, candidate) rows")
```

```{r completeness-candidates}
short_rankings <- selection %>%
  filter(n_candidates < length(model_levels)) %>%
  count(scenario, n, proxy, n_candidates, name = "rankings")

if (nrow(short_rankings) == 0) {
  cat("Every ranking scored all", length(model_levels), "candidates.\n")
} else {
  kable(short_rankings,
        col.names = c("Scenario", "n", "Proxy", "Candidates ranked", "Rankings"),
        caption = "Rankings built from fewer than the full candidate set")
}
```

## What is there to get right?

Before asking whether a proxy ranks the candidates correctly, it is worth
knowing whether the ranking matters. If all 9 candidates achieve near-identical
true PEHE then picking the wrong one costs nothing and a high error rate is
harmless; if they are far apart, selection is where the accuracy comes from.

```{r true-pehe-plot, fig.height=13}
pehe_plot <- metrics %>%
  ggplot(aes(x = model, y = true_pehe, colour = family)) +
  geom_boxplot(fill = "transparent", outlier.shape = NA) +
  facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
  scale_y_log10() +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "True PEHE of each candidate",
       subtitle = "The quantity a proxy is trying to reproduce the ordering of; log scale",
       y = "True PEHE (log scale)", x = "Candidate", colour = "Family")

pehe_plot
```

```{r winner-plot, fig.height=9}
# how often each candidate is genuinely the best one in its run
winners <- metrics %>%
  filter(!is.na(true_pehe)) %>%
  group_by(scenario, n, run) %>%
  slice_min(true_pehe, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  count(scenario, n, model, name = "wins", .drop = FALSE) %>%
  group_by(scenario, n) %>%
  mutate(prop = wins / sum(wins)) %>%
  ungroup()

winner_plot <- winners %>%
  ggplot(aes(x = model, y = prop, fill = model)) +
  geom_col() +
  facet_grid(rows = vars(scenario), cols = vars(n)) +
  scale_fill_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(title = "How often each candidate is the true best",
       subtitle = "Proportion of runs in which this candidate has the lowest true PEHE",
       y = "Proportion of runs", x = "Candidate")

winner_plot
```

```{r spread-table}
spread_tab <- selection %>%
  distinct(scenario, n, run, spread, random_regret) %>%
  group_by(scenario, n) %>%
  summarise(mean_spread = mean(spread, na.rm = T),
            mean_random_regret = mean(random_regret, na.rm = T),
            .groups = "drop") %>%
  left_join(select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
            by = c("scenario", "n"))

kable(spread_tab, digits = 4,
      col.names = c("Scenario", "n", "Worst - best true PEHE",
                    "Regret of a random pick", "Best fixed choice",
                    "Its regret"),
      caption = "How much selection is worth, per cell")
```

`Worst - best true PEHE` is the full range on the table; `regret of a random
pick` is what ignoring the scores altogether costs on average. Any proxy has
to beat the latter to be doing anything at all. The `best fixed choice` column
is the strongest a-priori rule available — always use that one candidate — and
it is chosen *with* knowledge of the truth, so it is an optimistic bound. A
proxy that cannot beat it has not earned the two extra nuisance pipelines it
costs.

## Rank agreement

Spearman and Kendall correlation between each proxy score and true PEHE,
computed across the 9 candidates within a run. Kendall is the better-behaved
of the two on 9 items and is the one to trust where they disagree. A proxy
that is useful for selection sits well above 0.

```{r rho-plot, fig.height=16}
rho_box <- box_plot(selection, "rho",
                    "Spearman correlation with true PEHE",
                    "Spearman correlation", hline = 0)

rho_sum <- summary_plot(selection_summary, "mean_rho", "mcse_rho",
                        "Mean Spearman correlation with true PEHE",
                        "Mean Spearman correlation", hline = 0)

(rho_box / rho_sum) + plot_layout(guides = "collect")
```

```{r kendall-plot, fig.height=9}
summary_plot(selection_summary, "mean_kendall", "mcse_kendall",
             "Mean Kendall correlation with true PEHE",
             "Mean Kendall tau", hline = 0)
```

## Top-1 selection accuracy

The decision a proxy is actually used to make: take its argmin and deploy that
candidate. `hit` is the proportion of runs where that is the same candidate
true PEHE's argmin is. The dashed line is chance — 1/9, what you would get by
picking at random.

```{r hit-plot, fig.height=9}
summary_plot(selection_summary, "mean_hit", "mcse_hit",
             "Top-1 selection accuracy",
             "Proportion of runs the proxy picks the true best",
             hline = 1 / length(model_levels))
```

Top-1 accuracy alone is harsh: a proxy that reliably lands on the
second-best candidate is doing a good job, and one that lands on the worst is
doing real damage, but both count as a miss. The rank of the selected
candidate, in the true-PEHE ordering, separates them.

```{r rank-dist-plot, fig.height=12}
rank_dist <- selection %>%
  count(scenario, n, proxy, pick_rank, name = "runs") %>%
  group_by(scenario, n, proxy) %>%
  mutate(prop = runs / sum(runs)) %>%
  ungroup()

# viridis rather than the house qualitative palette: pick_rank is ordinal, and
# rcartocolor::Safe would give 9 unordered hues to an ordered scale
rank_plot <- rank_dist %>%
  ggplot(aes(x = proxy, y = prop, fill = factor(pick_rank))) +
  geom_col() +
  facet_grid(rows = vars(scenario), cols = vars(n)) +
  scale_fill_viridis_d(direction = -1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Where the selected candidate really ranked",
       subtitle = "1 = the proxy picked the true best, 9 = it picked the true worst",
       y = "Proportion of runs", x = "Proxy score",
       fill = "True rank of pick")

rank_plot
```

```{r mean-rank-plot, fig.height=9}
summary_plot(selection_summary, "mean_rank", "mcse_rank",
             "Mean true rank of the selected candidate",
             "Mean rank (1 = best of 9)",
             hline = (1 + length(model_levels)) / 2)
```

The dashed line is the mean rank you would get by picking at random.

## Regret: what does using the proxy cost?

Regret is the excess true PEHE incurred by deploying the proxy's pick instead
of the true best candidate. It is the only one of these quantities on the
scale of the thing you care about — an error the ranking metrics count as a
miss is free if the two candidates were equally good.

The solid reference line in each panel is the regret of picking at random; the
dashed line is the best fixed choice. A proxy is worth its cost only below the
dashed line.

```{r regret-plot, fig.height=16}
regret_box <- box_plot(selection, "regret",
                       "Regret: excess true PEHE from following the proxy",
                       "Regret") +
  geom_hline(data = random_baseline, aes(yintercept = random_regret),
             linetype = "solid", colour = "grey40") +
  geom_hline(data = best_fixed, aes(yintercept = best_fixed_regret),
             linetype = "dashed", colour = "grey20")

regret_sum <- summary_plot(selection_summary, "mean_regret", "mcse_regret",
                           "Mean regret", "Mean regret") +
  geom_hline(data = random_baseline, aes(yintercept = random_regret),
             linetype = "solid", colour = "grey40") +
  geom_hline(data = best_fixed, aes(yintercept = best_fixed_regret),
             linetype = "dashed", colour = "grey20")

(regret_box / regret_sum) + plot_layout(guides = "collect")
```

Relative regret puts the same quantity on a scale that is comparable across
scenarios and sample sizes: excess PEHE as a fraction of the best achievable
PEHE.

```{r rel-regret-plot, fig.height=9}
summary_plot(selection_summary, "mean_rel_regret", "mcse_rel_regret",
             "Mean relative regret",
             "Excess PEHE as a fraction of the best achievable", hline = 0)
```

## Which axis matters?

The 8 proxies are a 2 x 2 x 2 design. Reading a winner off an 8-level axis
conflates the three questions it actually answers — influence-corrected versus
DR risk, honest crossfitting versus resubstitution, XGBoost versus AutoML — so
here each is marginalised over the other two.

```{r axis-marginals}
axis_long <- selection %>%
  mutate(across(c(score_type, fold_type, pipeline), as.character)) %>%
  pivot_longer(c(score_type, fold_type, pipeline),
               names_to = "axis", values_to = "level") %>%
  mutate(axis = factor(axis_labels[axis], levels = unname(axis_labels)),
         level = factor(level_labels[level], levels = unname(level_labels)))

axis_summary <- axis_long %>%
  group_by(scenario, n, axis, level) %>%
  summarise(
    mean_hit = mean(hit, na.rm = T),
    mcse_hit = sd(hit, na.rm = T) / sqrt(n()),
    mean_kendall = mean(kendall, na.rm = T),
    mcse_kendall = sd(kendall, na.rm = T) / sqrt(n()),
    mean_regret = mean(regret, na.rm = T),
    mcse_regret = sd(regret, na.rm = T) / sqrt(n()),
    .groups = "drop"
  )

axis_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL) {
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - .data[[mcse_col]],
           hi = .data[[mean_col]] + .data[[mcse_col]])
  p <- ggplot(df, aes(x = level, y = est, colour = n, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(axis), scales = "free_x") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = NULL, colour = "n")
}
```

```{r axis-accuracy-plot, fig.height=16}
hit_axis <- axis_plot(axis_summary, "mean_hit", "mcse_hit",
                      "Top-1 accuracy by design axis",
                      "Mean top-1 accuracy", hline = 1 / length(model_levels))

kendall_axis <- axis_plot(axis_summary, "mean_kendall", "mcse_kendall",
                          "Rank agreement by design axis",
                          "Mean Kendall tau", hline = 0)

(hit_axis / kendall_axis) + plot_layout(guides = "collect")
```

```{r axis-regret-plot, fig.height=9}
axis_plot(axis_summary, "mean_regret", "mcse_regret",
          "Regret by design axis", "Mean regret")
```

## Do the proxies track PEHE in level, not just rank?

Ranking is all selection needs, but a proxy that tracks PEHE in *level* would
also tell you how good the winner is. The scores cannot be compared to PEHE
raw: `calc_infl_score` and `calc_dr_risk` estimate PEHE only up to a
**run-specific additive constant** — the DR risk carries the irreducible
variance of the pseudo-outcome `phi` about `tau`, and the influence score
carries `E[tau^2]` terms that depend on the realised sample, not on the
candidate being scored. Pooling raw scores across runs would produce a
correlation driven entirely by that offset and would say nothing about the
candidates.

So both axes below are **centred within (scenario, n, run)**. What is left is
the only comparable part: how each candidate deviates from its run's average,
in the proxy and in the truth.

```{r calibration-plot, fig.height=11}
centred <- long_ok %>%
  group_by(scenario, n, run, proxy) %>%
  mutate(score_c = score - mean(score),
         pehe_c = true_pehe - mean(true_pehe)) %>%
  ungroup()

calibration_plot <- centred %>%
  ggplot(aes(x = pehe_c, y = score_c, colour = scenario)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_point(size = 0.5, alpha = 0.15) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  facet_wrap(~proxy, ncol = 4, scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  labs(title = "Proxy score against true PEHE, both centred within run",
       subtitle = "Centring is required: the proxies estimate PEHE only up to a run-specific constant",
       y = "Centred proxy score", x = "Centred true PEHE", colour = "Scenario")

calibration_plot
```

## Headline table

Per (scenario, n, proxy), arranged so the lowest-regret proxy comes first
within each cell.

```{r headline-table}
headline <- selection_summary %>%
  select(scenario, n, proxy, runs, mean_kendall, mean_rho, mean_hit,
         mean_rank, mean_regret, mean_rel_regret) %>%
  arrange(scenario, n, mean_regret)

kable(headline, digits = 4,
      col.names = c("Scenario", "n", "Proxy", "Runs", "Mean Kendall tau",
                    "Mean Spearman", "Top-1 accuracy", "Mean rank of pick",
                    "Mean regret", "Mean relative regret"))
```

For reference, the two baselines the regret columns should be read against:

```{r baseline-table}
baseline_tab <- random_baseline %>%
  left_join(select(best_fixed, scenario, n, best_fixed_model, best_fixed_regret),
            by = c("scenario", "n"))

kable(baseline_tab, digits = 4,
      col.names = c("Scenario", "n", "Regret of a random pick",
                    "Best fixed choice", "Its regret"))
```
