---
title: "Correlated continuous outcome study: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 9)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(knitr)
source(here("R", "figures.R"))
```

## Does correlation between covariates change how well the CATE is recovered?

`continuous/`'s scenarios 1–4, with X1–X5 and X01–X03 drawn from a Gaussian
copula with exchangeable latent correlation ρ, at four sample sizes (100, 250,
500, 1000) and 500 runs each. Every marginal, τ(x) and m0(x) are
`continuous/`'s; only the covariates' joint distribution changes. Full design
in `../README.md`.

| Scenario | CATE structure |
|---|---|
| 1: No HTE | $b_0 + b_1 X_1 + b_2 X_2 + W b_W$ |
| 2: Continuous HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 X_4)$ |
| 3: Two vars + cts x cts | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$ |
| 4: Cosine HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_4 \cos(X_4))$ |

Two values of ρ:

- **ρ = 0**: the copula with no correlation, the paired independent arm.
- **ρ = 0.5**: the correlated arm. The prognostic X1, X2 now move with the
  modifiers X3–X5, and the noise covariates X01–X03 become partial proxies for
  them.

What the correlation does to the design (`Rscript R/calibration_report.R`):

| | ρ = 0 | ρ = 0.5 |
|---|---|---|
| true ATE, n = 100 / 1000 | −0.65 / −0.20 | −0.60 / −0.19 |
| realised power, scenarios 2 / 3 / 4 | 0.65–0.67 / 0.61–0.63 / 0.76–0.77 | 0.65–0.66 / 0.55–0.56 / 0.75–0.76 |
| SD τ, scenario 3 | 1.16 | 1.36 |
| cor(m0, τ), scenarios 2 / 3 / 4 | 0 / 0 / 0 | −0.44 / 0.39 / 0.01 |

Scenario 3's SD(τ) grows with ρ, so its MSE is on a different footing at the
two values; the correlation with the true CATE is scale-free and compares
directly.

**Pairing.** Both values of ρ use the same run seeds, so run *r* shares W, the
raw latent normals, the outcome noise and the categorical covariates across ρ.
Every ρ = 0.5 − ρ = 0 comparison below is a per-run difference, with the
paired MCSE sd(diff)/√runs. The level plots show each ρ's own (unpaired) MCSE,
so the paired differences are the place to judge whether ρ matters.

**Estimators**: as `continuous/`: `causal_forest`, `dr_random_forest`,
`dr_oracle`, `dr_semi_oracle`, `dr_superlearner`, and the T-learners
`t_random_forest` and `t_superlearner`, recovered at metrics time from the DR
learners' saved per-arm outcome models (`R/metrics.R::add_t_learners()`).

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "correlated", "continuous")

metrics <- readRDS(file.path(res_path, "cts_corr_metrics.RDS"))

# the parent study's labels, so the two reports read side by side
scenario_labels <- c(
  `1` = "1: No HTE",
  `2` = "2: Continuous HTE",
  `3` = "3: Two vars + cts x cts",
  `4` = "4: Cosine HTE"
)

# rho stays numeric (the paired differences pivot on it); rho_lab is for facets.
# Also used on the parent study's metrics, which have no rho.
label_corr <- function(df) {
  df <- df %>%
    mutate(
      scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                        labels = unname(scenario_labels)),
      n = factor(n, levels = c(100, 250, 500, 1000))
    )
  if ("rho" %in% names(df)) df$rho_lab <- factor(paste("ρ =", df$rho))
  droplevels(df)
}

metrics <- metrics %>%
  mutate(model = label_factor(model, MODEL_LABELS),
         # each HTE test's rejection at alpha = 0.05, as a per-run 0/1
         # indicator: its mean is the power, with the binomial MCSE, and its
         # per-run difference across rho is what the paired power MCSE needs
         rej_BLP = as.numeric(BLP_p < 0.05),
         rej_BLP_os = as.numeric(BLP_p_os < 0.05),
         rej_indep_cate = as.numeric(indep_cate < 0.05),
         rej_indep_po = as.numeric(indep_po < 0.05)) %>%
  label_corr()

METRIC_COLS <- c(
  bias = "bias", ate_bias = "ate_bias", rel_ate_bias = "rel_ate_bias",
  rel_bias_cate = "rel_bias_cate", mse = "mse", rmse = "rmse", mae = "mae",
  corr = "corr", spearman = "spearman", sign_acc = "sign_acc",
  BLP = "BLP_p", BLP_os = "BLP_p_os", indep_cate = "indep_cate",
  indep_po = "indep_po",
  power_BLP = "rej_BLP", power_BLP_os = "rej_BLP_os",
  power_indep_cate = "rej_indep_cate", power_indep_po = "rej_indep_po"
)
POWER_STEMS <- c("power_BLP", "power_BLP_os", "power_indep_cate", "power_indep_po")

# per (rho, scenario, model, n) mean and MCSE of every metric
# (R/figures.R::summarise_metrics), plus the n_na diagnostic and the count of
# runs with an NA test p-value, which the power and mean p leave out
metrics_summary <- metrics %>%
  summarise_metrics(c("rho", "rho_lab", "scenario", "model", "n"),
                    cols = METRIC_COLS, binomial = POWER_STEMS) %>%
  left_join(metrics %>%
              group_by(rho, scenario, model, n) %>%
              summarise(mean_n_na = mean(n_na, na.rm = T),
                        total_n_na = sum(n_na, na.rm = T), .groups = "drop"),
            by = c("rho", "scenario", "model", "n"))
```

```{r plot-helpers, include=FALSE}
# scenario in rows, rho in columns: facet_grid frees y by row, so each scenario
# keeps its own scale (as the parent report's facet_wrap does) while its two
# rho panels share one, and compare by eye
facet_rho <- function(free_y = TRUE) {
  facet_grid(rows = vars(scenario), cols = vars(rho_lab),
             scales = if (free_y) "free_y" else "fixed")
}

house_style <- function() {
  list(drf_scale(), drf_theme(),
       theme(axis.text.x = element_text(angle = 45, hjust = 1),
             legend.position = "bottom"))
}

# per-run distribution, one box per (n, model)
box_plot <- function(df, col, title, ylab, hline = 0, free_y = TRUE) {
  df %>%
    ggplot(aes(x = n, y = .data[[col]], colour = model)) +
    (if (!is.null(hline)) geom_hline(yintercept = hline, linetype = "dashed")) +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_rho(free_y) +
    house_style() +
    labs(title = title, y = ylab, x = "Sample size", colour = "Model")
}

# the parent report's summary_plot: mean with a 95% CI (mean +/- qnorm(1 -
# alpha/2) x MCSE, not a raw +/- 1x MCSE - see continuous/cts_results.qmd)
summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = 0, alpha = 0.05,
                         free_y = TRUE) {
  z <- qnorm(1 - alpha / 2)
  df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - z * .data[[mcse_col]],
           hi = .data[[mean_col]] + z * .data[[mcse_col]]) %>%
    ggplot(aes(x = n, y = est, colour = model, ymin = lo, ymax = hi)) +
    (if (!is.null(hline)) geom_hline(yintercept = hline, linetype = "dashed")) +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_rho(free_y) +
    house_style() +
    labs(title = title, y = ylab, x = "Sample size", colour = "Model")
}

#' Paired rho = 0.5 - rho = 0 difference in a per-run column, by cell
#'
#' Runs are matched on (by, run). The MCSE is sd(diff) / sqrt(pairs), which is
#' what the shared seeds buy over the unpaired sqrt(mcse_0^2 + mcse_0.5^2). A
#' run with the column NA at either rho (e.g. a failed dr_superlearner fit)
#' drops out of its cell.
paired_diff <- function(df, col, by = c("scenario", "n", "model")) {
  df %>%
    select(all_of(c(by, "run", "rho")), value = all_of(col)) %>%
    pivot_wider(names_from = rho, values_from = value, names_prefix = "rho_") %>%
    mutate(diff = rho_0.5 - rho_0) %>%
    group_by(across(all_of(by))) %>%
    summarise(mean_diff = mean(diff, na.rm = TRUE),
              mcse_diff = sd(diff, na.rm = TRUE) / sqrt(sum(!is.na(diff))),
              pairs = sum(!is.na(diff)), .groups = "drop")
}

# one row of scenarios, as the parent report's summary plots
diff_plot <- function(df, title, ylab, alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  df %>%
    ggplot(aes(x = n, y = mean_diff, colour = model,
               ymin = mean_diff - z * mcse_diff, ymax = mean_diff + z * mcse_diff)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_wrap(~scenario, nrow = 1, scales = "free_y") +
    house_style() +
    labs(title = title, y = ylab, x = "Sample size", colour = "Model")
}
```

## Effect of the correlation: paired ρ = 0.5 − ρ = 0 differences

Each point is the mean over runs of a run's metric at ρ = 0.5 minus the same
run's at ρ = 0, with a 95% CI from the paired MCSE. A CI clear of the dashed
line is a change the correlation made, not Monte Carlo noise. Scenario 1 is
left out of the correlation panels, where its correlation is fixed at 0.

```{r diff-plots, fig.height=4}
diff_plot(paired_diff(metrics, "bias"), "Bias of the CATE", "Δ bias (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics, "ate_bias"), "ATE bias", "Δ ATE bias (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics, "mse"), "MSE of the CATE", "Δ MSE (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(filter(metrics, scenario != scenario_labels[["1"]]), "corr"),
          "Correlation with the true CATE", "Δ Pearson correlation (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(filter(metrics, scenario != scenario_labels[["1"]]), "spearman"),
          "Spearman correlation with the true CATE",
          "Δ Spearman correlation (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics, "sign_acc"), "Sign accuracy",
          "Δ proportion correct sign (ρ = 0.5 − ρ = 0)")
```

The tests' rejection rates at 0.05: a change in type I error in scenario 1,
in power in scenarios 2–4.

```{r diff-power-plots, fig.height=4}
diff_plot(paired_diff(metrics, "rej_BLP_os"), "BLP test power (one-sided, HC3)",
          "Δ rejection rate (ρ = 0.5 − ρ = 0)")
diff_plot(paired_diff(metrics, "rej_indep_cate"), "CATE permutation test power",
          "Δ rejection rate (ρ = 0.5 − ρ = 0)")
```

## Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per
run, against the known truth.

```{r bias-plot}
box_plot(metrics, "bias", "Bias in CATE estimates", "Bias")
summary_plot(metrics_summary, "mean_bias", "mcse_bias",
             "Mean bias in CATE estimates", "Mean bias")
```

## ATE bias

Bias of the *average* treatment effect (`mean(est) - mean(true)`), as
opposed to Bias's per-unit average.

```{r ate-bias-plot}
box_plot(metrics, "ate_bias", "Bias in the average treatment effect", "ATE bias")
summary_plot(metrics_summary, "mean_ate_bias", "mcse_ate_bias",
             "Mean ATE bias", "Mean ATE bias")
```

## Bias relative to the DR-oracle

Each run's ATE bias minus the DR-oracle's on the same dataset, within each ρ.
Every model is fitted to the same dataset in a run, and scenarios share random
numbers, so the noise in a run's outcomes shifts every model's bias together.
The DR-oracle uses the true outcome model and propensity, so its bias is zero
by construction and its realised bias is that common noise. Subtracting it
leaves what each estimator adds, with a paired MCSE much tighter than the ATE
bias section's. The DR-oracle sits at exactly 0 as the reference.

```{r bias-vs-oracle-plot}
oracle_ate_bias <- metrics %>%
  filter(model == "DR-oracle") %>%
  select(rho, scenario, n, run, oracle_ate_bias = ate_bias)

# NA wherever the model's ate_bias is NA, so those runs drop out of the mean
# and the MCSE denominator
metrics_vs_oracle <- metrics %>%
  left_join(oracle_ate_bias, by = c("rho", "scenario", "n", "run")) %>%
  mutate(ate_bias_vs_oracle = ate_bias - oracle_ate_bias)

vs_oracle_summary <- metrics_vs_oracle %>%
  summarise_metrics(c("rho", "rho_lab", "scenario", "model", "n"),
                    cols = c(ate_bias_vs_oracle = "ate_bias_vs_oracle"),
                    count_na = character())

box_plot(metrics_vs_oracle, "ate_bias_vs_oracle", "ATE bias relative to the DR-oracle",
         "ATE bias - DR-oracle's")
summary_plot(vs_oracle_summary, "mean_ate_bias_vs_oracle", "mcse_ate_bias_vs_oracle",
             "Mean ATE bias relative to the DR-oracle", "Mean ATE bias - DR-oracle's")
```

## Relative bias

`rel_ate_bias` is the parameter-level relative bias (ATE bias divided by the
true ATE); `rel_bias_cate` is the per-unit version, `(est - true) / true`
averaged over units (`NA` wherever the true CATE is exactly 0 for a unit — see
`R/metrics.R::cate_metrics()`).

```{r rel-bias-plot}
summary_plot(metrics_summary, "mean_rel_ate_bias", "mcse_rel_ate_bias",
             "Mean relative bias in the ATE", "Mean relative ATE bias")
summary_plot(metrics_summary, "mean_rel_bias_cate", "mcse_rel_bias_cate",
             "Mean relative bias in the CATE", "Mean relative CATE bias")
```

## MSE of CATE estimates

The headline metric: mean squared error of the estimated CATE against the
known truth.

```{r mse-plot}
box_plot(metrics, "mse", "MSE of CATE estimates", "MSE")
summary_plot(metrics_summary, "mean_mse", "mcse_mse", "Mean MSE of CATE estimates",
             "Mean MSE")
```

## RMSE of CATE estimates

```{r rmse-plot}
box_plot(metrics, "rmse", "RMSE of CATE estimates", "RMSE")
summary_plot(metrics_summary, "mean_rmse", "mcse_rmse", "Mean RMSE of CATE estimates",
             "Mean RMSE")
```

## MAE of CATE estimates

Mean absolute error — a robustness check on MSE, less sensitive to outlying
runs.

```{r mae-plot}
box_plot(metrics, "mae", "MAE of CATE estimates", "MAE")
summary_plot(metrics_summary, "mean_mae", "mcse_mae", "Mean MAE of CATE estimates",
             "Mean MAE")
```

## Correlation with the true CATE

Pearson correlation between estimated and true CATEs (undefined, and forced
to 0, in scenario 1 — excluded from these plots).

```{r corr-plot, fig.height=7}
corr_df <- filter(metrics, scenario != scenario_labels[["1"]])
corr_summary <- filter(metrics_summary, scenario != scenario_labels[["1"]])

box_plot(corr_df, "corr", "Correlation between estimated and true CATEs",
         "Pearson correlation", hline = NULL, free_y = FALSE)
summary_plot(corr_summary, "mean_corr", "mcse_corr",
             "Mean correlation with the true CATE", "Mean correlation",
             hline = NULL, free_y = FALSE)
```

## Spearman rank correlation with the true CATE

Same scenario-1 convention, so scenario 1 is excluded here too.

```{r spearman-plot, fig.height=7}
box_plot(corr_df, "spearman", "Spearman rank correlation between estimated and true CATEs",
         "Spearman correlation", hline = NULL, free_y = FALSE)
summary_plot(corr_summary, "mean_spearman", "mcse_spearman",
             "Mean Spearman correlation with the true CATE",
             "Mean Spearman correlation", hline = NULL, free_y = FALSE)
```

## Sign accuracy

Proportion of units where the estimated CATE has the correct sign. 0.5 (dashed
line) is the chance level.

```{r sign-acc-plot}
box_plot(metrics, "sign_acc", "Sign accuracy of CATE estimates",
         "Proportion correct sign", hline = 0.5, free_y = FALSE)
summary_plot(metrics_summary, "mean_sign_acc", "mcse_sign_acc",
             "Mean sign accuracy", "Mean proportion correct sign",
             hline = 0.5, free_y = FALSE)
```

## True-CATE HTE test evaluation

The BLP and CATE independence tests re-run on the true CATE and true
nuisances (`cts_corr_true_cate_tests.RDS`, `R/cate_models.R::true_cate_test_row()`):
the ceiling each model's own test is chasing. Shown as an extra "True CATE"
series in the test sections below. In scenario 1 the true CATE is constant,
so neither test is defined there and "True CATE" has no scenario-1 series.

```{r true-cate-data}
true_cate_tests <- readRDS(file.path(res_path, "cts_corr_true_cate_tests.RDS")) %>%
  mutate(rej_BLP = as.numeric(BLP_p < 0.05),
         rej_BLP_os = as.numeric(BLP_p_os < 0.05),
         rej_indep_cate = as.numeric(indep_cate < 0.05)) %>%
  label_corr()

true_cate_summary <- true_cate_tests %>%
  summarise_metrics(c("rho", "rho_lab", "scenario", "n"),
                    cols = METRIC_COLS, binomial = POWER_STEMS) %>%
  mutate(model = "True CATE")

true_cate_raw <- true_cate_tests %>% mutate(model = "True CATE")

model_levels_true <- c(levels(metrics$model), "True CATE")

# row-bind the true-CATE reference series onto a per-run or summarised data
# frame, extending `model`'s levels so it becomes one more colour
bind_true_cate <- function(df, true_df, extra_cols) {
  bind_rows(
    mutate(df, model = factor(as.character(model), levels = model_levels_true)),
    true_df %>%
      mutate(model = factor(model, levels = model_levels_true)) %>%
      select(rho, rho_lab, scenario, n, model, all_of(extra_cols))
  )
}
```

## BLP test p-values

Best linear predictor test of heterogeneity, against the nominal 5% threshold
(dashed line). `BLP_p` is two-sided with constant-variance SEs, as computed
when the models were fitted; `BLP_p_os` is one-sided (H1: beta.2 > 0) with HC3
SEs, recomputed at metrics time (see `continuous/README.md`, "HTE tests").

```{r blp-plot}
box_plot(bind_true_cate(metrics, true_cate_raw, "BLP_p"), "BLP_p",
         "BLP test p-values (two-sided, constant variance)", "p-value",
         hline = 0.05, free_y = FALSE)
summary_plot(bind_true_cate(metrics_summary, true_cate_summary, c("mean_BLP", "mcse_BLP")),
             "mean_BLP", "mcse_BLP", "Mean BLP test p-value (two-sided, constant variance)",
             "Mean p-value", hline = 0.05, free_y = FALSE)
```

```{r blp-os-plot}
box_plot(bind_true_cate(metrics, true_cate_raw, "BLP_p_os"), "BLP_p_os",
         "BLP test p-values (one-sided, HC3)", "p-value", hline = 0.05, free_y = FALSE)
summary_plot(bind_true_cate(metrics_summary, true_cate_summary,
                            c("mean_BLP_os", "mcse_BLP_os")),
             "mean_BLP_os", "mcse_BLP_os", "Mean BLP test p-value (one-sided, HC3)",
             "Mean p-value", hline = 0.05, free_y = FALSE)
```

## CATE permutation test p-values

```{r indep-cate-plot}
box_plot(bind_true_cate(metrics, true_cate_raw, "indep_cate"), "indep_cate",
         "CATE permutation test p-values", "p-value", hline = 0.05, free_y = FALSE)
summary_plot(bind_true_cate(metrics_summary, true_cate_summary,
                            c("mean_indep_cate", "mcse_indep_cate")),
             "mean_indep_cate", "mcse_indep_cate", "Mean CATE permutation test p-value",
             "Mean p-value", hline = 0.05, free_y = FALSE)
```

## PO permutation test p-values

No causal forest series: its pseudo-outcome is DR-RandomForest's, so the test
is reported once. Likewise the T-learners share their DR learner's.

```{r indep-po-plot}
box_plot(metrics, "indep_po", "PO permutation test p-values", "p-value",
         hline = 0.05, free_y = FALSE)
summary_plot(metrics_summary, "mean_indep_po", "mcse_indep_po",
             "Mean PO permutation test p-value", "Mean p-value",
             hline = 0.05, free_y = FALSE)
```

## HTE test power

Proportion of runs rejecting at alpha = 0.05 (type I error in scenario 1),
with the binomial 95% CI.

```{r power-plot}
summary_plot(bind_true_cate(metrics_summary, true_cate_summary,
                            c("mean_power_BLP", "mcse_power_BLP")),
             "mean_power_BLP", "mcse_power_BLP",
             "BLP test power (two-sided, constant variance)", "Power",
             hline = 0.05, free_y = FALSE)
summary_plot(bind_true_cate(metrics_summary, true_cate_summary,
                            c("mean_power_BLP_os", "mcse_power_BLP_os")),
             "mean_power_BLP_os", "mcse_power_BLP_os", "BLP test power (one-sided, HC3)",
             "Power", hline = 0.05, free_y = FALSE)
summary_plot(bind_true_cate(metrics_summary, true_cate_summary,
                            c("mean_power_indep_cate", "mcse_power_indep_cate")),
             "mean_power_indep_cate", "mcse_power_indep_cate",
             "CATE permutation test power", "Power", hline = 0.05, free_y = FALSE)
summary_plot(metrics_summary, "mean_power_indep_po", "mcse_power_indep_po",
             "PO permutation test power", "Power", hline = 0.05, free_y = FALSE)
```

The true CATE's own power, paired across ρ: how much of a model's change in
power is the HTE itself getting harder to detect (scenario 3's realised power
falls from about 0.62 to 0.55 by design).

```{r true-cate-power-diff}
true_cate_diff <- bind_rows(
  paired_diff(true_cate_tests, "rej_BLP_os", by = c("scenario", "n")) %>%
    mutate(test = "BLP (one-sided, HC3)"),
  paired_diff(true_cate_tests, "rej_indep_cate", by = c("scenario", "n")) %>%
    mutate(test = "CATE permutation")
) %>%
  filter(pairs > 0) %>%
  transmute(scenario, n, test,
            `Δ power (MCSE)` = sprintf("%.3f (%.3f)", mean_diff, mcse_diff)) %>%
  pivot_wider(names_from = test, values_from = `Δ power (MCSE)`) %>%
  arrange(scenario, n)

kable(true_cate_diff, caption = "True-CATE test power, ρ = 0.5 − ρ = 0 (paired)")
```

Runs with an `NA` test p-value are left out of that test's power and mean
p-value: a degenerate (constant) CATE estimate, or a test that wasn't run.
Cells with any are listed below.

```{r test-na-table}
test_na_table <- metrics_summary %>%
  select(rho, scenario, model, n, n_na_BLP_p, n_na_BLP_p_os, n_na_indep_cate) %>%
  filter(n_na_BLP_p > 0 | n_na_BLP_p_os > 0 | n_na_indep_cate > 0) %>%
  arrange(scenario, n, rho, model)

if (nrow(test_na_table) > 0) {
  kable(test_na_table,
        col.names = c("ρ", "Scenario", "Model", "n", "NA BLP p",
                      "NA BLP p (one-sided, HC3)", "NA CATE perm. p"))
} else {
  cat("No NA HTE test p-values in any run.")
}
```

## Missing estimates

`n_na` is the count of `NA` CATE estimates within a run; only cells with at
least one are listed.

```{r na-table}
na_table <- metrics_summary %>%
  select(rho, scenario, model, n, mean_n_na, total_n_na) %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, n, rho, model)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("ρ", "Scenario", "Model", "n", "Mean NA count per run",
                      "Total NA count"))
} else {
  cat("No missing CATE estimates in any run.")
}
```

## Sanity check: ρ = 0 against `continuous/`

The ρ = 0 arm has the same distribution as `continuous/`'s scenarios 1–10, but
different draws: the copula draws in another order, so the two are **not**
paired and the z-scores below use the unpaired MCSE. With this many cells a
few `|z| > 3` are expected by chance; many more point to a DGM or pipeline
difference. Skipped if `continuous/`'s metrics are not on disk.

```{r sanity}
parent_file <- file.path(dirname(path), "results", "continuous", "cts_metrics.RDS")

if (!file.exists(parent_file)) {
  cat("continuous/'s metrics not found - skipped.\n")
} else {
  parent <- readRDS(parent_file) %>%
    mutate(model = label_factor(model, MODEL_LABELS),
           rej_BLP_os = as.numeric(BLP_p_os < 0.05)) %>%
    label_corr()
  cols <- c(bias = "bias", ate_bias = "ate_bias", mse = "mse", corr = "corr",
            power_BLP_os = "rej_BLP_os")
  keys <- c("scenario", "model", "n")
  rho0 <- summarise_metrics(filter(metrics, rho == 0), keys, cols = cols,
                            binomial = "power_BLP_os", count_na = character())
  par <- summarise_metrics(parent, keys, cols = cols, binomial = "power_BLP_os",
                           count_na = character())
  cmp <- inner_join(rho0, par, by = keys, suffix = c("_rho0", "_parent"))
  z <- sapply(names(cols), function(s) {
    (cmp[[paste0("mean_", s, "_rho0")]] - cmp[[paste0("mean_", s, "_parent")]]) /
      sqrt(cmp[[paste0("mcse_", s, "_rho0")]]^2 + cmp[[paste0("mcse_", s, "_parent")]]^2)
  })
  z_df <- bind_cols(select(cmp, all_of(keys)),
                    as.data.frame(round(z, 2)) %>% rename_with(~ paste0("z_", .x)))
  # scenario 1's corr is fixed at 0 in both, so its z is NaN - not a flag
  n_z <- sum(is.finite(z))
  n_big <- sum(abs(z) > 3, na.rm = TRUE)
  cat(sprintf("%d of %d finite z-scores have |z| > 3 (%.1f expected by chance).\n",
              n_big, n_z, n_z * 2 * pnorm(-3)))
  kable(filter(z_df, if_any(starts_with("z_"), ~ abs(.x) > 3)) %>%
          arrange(scenario, n, model),
        caption = "Cells with |z| > 3, ρ = 0 vs continuous/")
}
```

## Headline table

Arranged by scenario, sample size, ρ and mean MSE (best model first within
each cell). The ATE bias relative to the DR-oracle is shown with its paired
MCSE.

```{r headline-table}
headline <- metrics_summary %>%
  left_join(select(vs_oracle_summary, rho, scenario, model, n,
                   mean_ate_bias_vs_oracle, mcse_ate_bias_vs_oracle),
            by = c("rho", "scenario", "model", "n")) %>%
  select(scenario, n, rho, model, mean_bias, mean_ate_bias, mean_ate_bias_vs_oracle,
         mcse_ate_bias_vs_oracle, mean_rel_ate_bias, mean_rel_bias_cate, mean_mse,
         mean_rmse, mean_mae, mean_sign_acc, mean_corr, mean_power_BLP,
         mean_power_BLP_os, mean_power_indep_cate, mean_power_indep_po) %>%
  arrange(scenario, n, rho, mean_mse)

kable(headline, digits = 4,
      col.names = c("Scenario", "n", "ρ", "Model", "Mean bias", "Mean ATE bias",
                    "Mean ATE bias vs DR-oracle", "MCSE (vs DR-oracle)",
                    "Mean relative ATE bias", "Mean relative CATE bias",
                    "Mean MSE", "Mean RMSE", "Mean MAE", "Mean sign accuracy",
                    "Mean correlation", "BLP power", "BLP power (one-sided, HC3)",
                    "CATE perm. power", "PO perm. power"))
```
