---
title: "Binary CI study: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(paletteer)
library(here)
library(patchwork)
library(purrr)
library(knitr)
```

## How should the half-sample bootstrap's subsampling ratio be set?

**This study's underlying data is superseded (Bug A).** `bin_ci_dgms.R` ran on
the *continuous* outcome's coefficient table (`b0 = c(0.4, 0.2, ...)`,
`b1 = -0.05`, `b2 = c(2, 2, ...)`) fed through `plogis`, instead of the binary
table the main `binary/` study uses. With `b2 = 2` on a logit scale the outcome
model saturates — most of the covariate range sits in the flat tails of
`plogis` — so the true CATEs are compressed toward zero and coverage below is
measured against a target the design never intended. `../results/confidence_intervals/binary/`
needs a full re-run under the fixed DGM before these numbers are comparable to
the main binary study; see `confidence_intervals/binary/README.md` for the fix
and status. Read everything below as a caveat pending that re-run, not as a
finished result.

`CI_sf` is the `sample.fraction` passed to the half-sample forests, and it is
the one tuning choice the bootstrap procedure forces on us. Ritzwoller, who
wrote the bootstrap paper with Syrgkanis, reports that the intervals perform
well at a ratio of 0.05. The GRF papers also note that a lower ratio improves
confidence intervals, yet grf's own default is 0.5. That is a wide disagreement
over a parameter that is not knowable a priori — too small and the half-sample
forests are noisy, too large and the half samples stop being independent
enough — so this study sweeps it: 0.05, 0.10, ..., 0.50.

Ten scenarios vary the structure of the CATE, crossed with two sample sizes and
the ten-point `CI_sf` sweep, 100 runs each. Full design in
`confidence_intervals/README.md`; the grid itself is defined once, in
`bin_ci_config.R`.

| | |
|---|---|
| scenarios | 1–10 |
| n | 500, 1000 |
| `CI_sf` | 0.05 to 0.50 in steps of 0.05 |
| runs | 100 |
| array | **20,000 jobs**, split across two jobscripts |

| Scenario | CATE structure |
|---|---|
| 1: No HTE | $b_0 + b_1 X_1 + b_2 X_2 + W b_W$ |
| 2: Binary HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_3 X_3)$ |
| 3: Continuous HTE | $b_0 + b_1 X_1 + b_2 X_2 + W (b_W + b_4 X_4)$ |
| 4: Two HTE vars | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4)$ |
| 5: Cts x binary interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{34} X_3 X_4)$ |
| 6: Two vars + cts x binary | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{34} X_3 X_4)$ |
| 7: Cts x cts interaction | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_{45} X_4 X_5)$ |
| 8: Two vars + cts x cts | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 X_4 + b_{45} X_4 X_5)$ |
| 9: Cosine HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_4 \cos(X_4))$ |
| 10: Exponential HTE | $b_0 + b_1 X_1 + b_2 X_2 + W(b_W + b_3 X_3 + b_4 \exp(-\lvert X_4 \rvert))$ |

The linear predictor above enters `plogis()` to give the binary outcome's
success probability — the same scenario formulas as `continuous/`, on the
logit scale instead of the identity scale.

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_oracle`,
`dr_semi_oracle` — `CI_MODELS` in `R/metrics.R`, fit with `family = binomial()`
and `oracle_link = "logit"` (`bin_ci_models.R`). There is no SuperLearner arm
and no BLP or independence test: this study runs `profile = "ci"`, because the
question is interval coverage rather than heterogeneity detection, and the
bootstrap already dominates the runtime.

### Performance measures

* **Marginal coverage** — the proportion of units their own interval covers.
  Shown both per run and as the mean across runs.
* **Simultaneous coverage** — 0/1 per run: did the band cover *every* unit at
  once. This is the guarantee the half-sample bootstrap is constructed to
  control, and the much stricter target.
* **CI length** — mean interval length within a run. Shown per run and as the
  mean across runs.

Nominal is 0.95 for both coverages.

### Three interval kinds

`bin_ci_metrics.R` scores three intervals, not one, and writes them as separate
rows of the same `model` column. They are split back apart below into an
estimator and an interval kind, so that the comparison between them is on its own
axis rather than mixed into the estimator legend.

| kind | arms | how |
|---|---|---|
| Half-sample bootstrap (per unit) | all 4 | the study's headline interval — simultaneous band over the sample's own units |
| Causal forest variance (pointwise) | `causal_forest` only | grf's own variance estimates turned into a normal-approximation interval by `normal_interval()`, at no extra compute. Written as `causal_forest_inbuilt` |
| Half-sample bootstrap (query grid) | all 4 | the same band evaluated over the fixed covariate-grid query points from `build_query_grid()` rather than the sample's units. Written as `<model>_grid` |

Everything below the interval-kind comparison uses the per-unit bootstrap, the
one interval every arm has.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "confidence_intervals", "binary")

metrics <- readRDS(file.path(res_path, "ci_bin_metrics.RDS"))

# display order and labels - the CI studies have no shared label-dictionary
# script, so these stay inline, matching bin_results.qmd's precedent
scenario_labels <- c(
  `1` = "1: No HTE",
  `2` = "2: Binary HTE",
  `3` = "3: Continuous HTE",
  `4` = "4: Two HTE vars",
  `5` = "5: Cts x binary interaction",
  `6` = "6: Two vars + cts x binary",
  `7` = "7: Cts x cts interaction",
  `8` = "8: Two vars + cts x cts",
  `9` = "9: Cosine HTE",
  `10` = "10: Exponential HTE"
)

# CI_MODELS - no SuperLearner arm in this study
model_labels <- c(
  causal_forest = "Causal forest",
  dr_random_forest = "DR-RandomForest",
  dr_oracle = "DR-oracle",
  dr_semi_oracle = "DR-semi-oracle"
)

interval_labels <- c(
  half_boot      = "Half-sample bootstrap (per unit)",
  grf_normal     = "Causal forest variance (pointwise)",
  half_boot_grid = "Half-sample bootstrap (query grid)"
)

# tidy up. bin_ci_metrics.R writes the alternative intervals as extra *rows* of
# `model` - "causal_forest_inbuilt" and "<model>_grid" - so split that one
# column into the estimator and the interval kind. Without this the _grid rows
# recode to NA and disappear silently, which is what bin_ci_results.Rmd did.
metrics <- metrics %>%
  # filter to 4 interesting scenarios to see figures clearer
  filter(scenario %in% c(1,3,8,9)) %>%
  mutate(
    interval = case_when(
      model == "causal_forest_inbuilt" ~ "grf_normal",
      grepl("_grid$", model)           ~ "half_boot_grid",
      TRUE                             ~ "half_boot"
    ),
    model = sub("_inbuilt$|_grid$", "", model)
  ) %>%
  mutate(
    scenario = factor(scenario, levels = names(scenario_labels) %>% as.integer(),
                      labels = unname(scenario_labels)),
    n = factor(n, levels = c(500, 1000)),
    # the same seq() the study config uses, so the doubles match exactly
    CI_sf = factor(CI_sf, levels = seq(0.05, 0.5, 0.05)),
    model = factor(recode(model, !!!model_labels), levels = unname(model_labels)),
    interval = factor(recode(interval, !!!interval_labels),
                      levels = unname(interval_labels))
  ) %>%
  droplevels()

# per (scenario, n, CI_sf, model, interval) summaries. Computed over every
# interval kind so the comparison section can reuse them; `summary_hb` is the
# bootstrap-only view every other section uses.
#
# mcse denominator is sqrt(sum(!is.na(x))), matching the non-NA count of that
# column - not sqrt(n()), the whole group's row count regardless of NAs.
# interval_metrics() (R/metrics.R) has no internal na.rm, so one failed CI
# bound in a run makes marginal_coverage/mean_ci_length NA for that whole run.
metrics_summary <- metrics %>%
  group_by(scenario, n, CI_sf, model, interval) %>%
  summarise(
    mean_marg_cov = mean(marginal_coverage, na.rm = T),
    mcse_marg_cov = sd(marginal_coverage, na.rm = T) / sqrt(sum(!is.na(marginal_coverage))),
    # simultaneous_coverage is a genuine 0/1 indicator per run (whole-band
    # coverage), so it gets the exact binomial MCSE, not the general one
    # marginal_coverage above uses (that's a proportion over many units
    # within a run, not Bernoulli at the run level)
    mean_simul_cov = mean(simultaneous_coverage, na.rm = T),
    mcse_simul_cov = sqrt(mean_simul_cov * (1 - mean_simul_cov) / sum(!is.na(simultaneous_coverage))),
    mean_ci_len = mean(mean_ci_length, na.rm = T),
    mcse_ci_len = sd(mean_ci_length, na.rm = T) / sqrt(sum(!is.na(mean_ci_length))),
    .groups = "drop"
  )

# everything below the interval-kind comparison is about the per-unit bootstrap
metrics_hb <- filter(metrics, interval == interval_labels[["half_boot"]])
summary_hb <- filter(metrics_summary, interval == interval_labels[["half_boot"]])
```

```{r plot-helpers, include=FALSE}
# helpers: the house pair of plots, per-run distribution and mean with MCSE
# error bars - see cts_results.qmd's summary_plot for the pattern. The sweep
# goes on x and the design on the facets, so scenario x n is a facet_grid
# rather than cts_results.qmd's facet_wrap(~scenario).
#
# hline is optional rather than defaulting to 0 (as in cf_ci_results.qmd): the
# coverages have a nominal 0.95 to sit against, CI width has no target at all.
box_plot <- function(df, metric, title, ylab, hline = NULL) {
  p <- ggplot(df, aes(x = CI_sf, y = .data[[metric]], colour = model))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Subsampling ratio", colour = "Model")
}

summary_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL, alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - z * .data[[mcse_col]],
           hi = .data[[mean_col]] + z * .data[[mcse_col]])
  p <- ggplot(df, aes(x = CI_sf, y = est, colour = model, ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.5), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.5), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Subsampling ratio", colour = "Model")
}

# the interval-kind comparison needs a different assignment of aesthetics: the
# kind takes the colour, so the estimator moves onto the facet columns and n
# onto the point shape.
interval_plot <- function(df, mean_col, mcse_col, title, ylab, hline = NULL, alpha = 0.05) {
  z <- qnorm(1 - alpha / 2)
  df <- df %>%
    mutate(est = .data[[mean_col]],
           lo = .data[[mean_col]] - z * .data[[mcse_col]],
           hi = .data[[mean_col]] + z * .data[[mcse_col]])
  p <- ggplot(df, aes(x = CI_sf, y = est, colour = interval, shape = n,
                      ymin = lo, ymax = hi))
  if (!is.null(hline)) p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  p +
    geom_point(position = position_dodge(width = 0.6), size = 2) +
    geom_errorbar(position = position_dodge(width = 0.6), linewidth = 0.3, width = 0.3) +
    facet_grid(rows = vars(scenario), cols = vars(model), scales = "free_y") +
    scale_colour_paletteer_d("rcartocolor::Safe") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, y = ylab, x = "Subsampling ratio",
         colour = "Interval kind", shape = "Sample size")
}
```

## Marginal coverage

The proportion of units their own interval covers, against the nominal 95%
(dashed line), across the sweep.

```{r marg-cov-plot, fig.height=22}
marg_cov_plot <- box_plot(metrics_hb, "marginal_coverage",
                          "Marginal coverage", "Marginal coverage", hline = 0.95)

marg_cov_sum_plot <- summary_plot(summary_hb, "mean_marg_cov", "mcse_marg_cov",
                                  "Mean marginal coverage", "Marginal coverage",
                                  hline = 0.95)

(marg_cov_plot / marg_cov_sum_plot) + plot_layout(guides = "collect")
```

## Simultaneous coverage

Whether the band covers every unit at once — what the half-sample bootstrap is
actually constructed to control. This is a per-run 0/1 indicator, so only the
mean-proportion summary is shown: a boxplot of a binary outcome carries no
information.

```{r simul-cov-plot, fig.height=18}
simul_cov_sum_plot <- summary_plot(summary_hb, "mean_simul_cov", "mcse_simul_cov",
                                   "Mean simultaneous coverage",
                                   "Simultaneous coverage", hline = 0.95)

simul_cov_sum_plot
```

## CI length

The width that coverage is bought with. No reference line — there is no target
width, only the trade-off against the two coverage sections above. A ratio that
reaches nominal coverage by producing enormous bands has not solved anything.

```{r ci-len-plot, fig.height=22}
ci_len_plot <- box_plot(metrics_hb, "mean_ci_length", "CI length across runs",
                        "CI length")

ci_len_sum_plot <- summary_plot(summary_hb, "mean_ci_len", "mcse_ci_len",
                                "Mean CI length", "Mean CI length")

(ci_len_plot / ci_len_sum_plot) + plot_layout(guides = "collect")
```

## The three interval kinds

The sweep is the same in all three panels below; what changes is which interval
is being scored. Two things to hold on to when reading them:

**The pointwise interval is not competing on simultaneous coverage.** grf's
variance estimates give a per-unit interval, so it should score near zero on the
simultaneous panel — that is the method working as designed, not a defect. The
marginal panel is the fair reading of it, and there it is the free comparator:
does an expensive bootstrap buy anything over the variance grf hands back for
nothing? Note also that it does not depend on `CI_sf` at all — no half-sample
forests are involved — so a flat line across the sweep is expected, and any
visible slope is Monte Carlo noise.

**The query-grid band is a different target, not a harder version of the same
one.** It takes its supremum over the fixed covariate grid from
`build_query_grid()` rather than over the sample's own units, so it is generally
the easier of the two and the numbers are not directly comparable to the per-unit
band.

```{r interval-marg-cov, fig.height=22}
interval_plot(metrics_summary, "mean_marg_cov", "mcse_marg_cov",
              "Mean marginal coverage by interval kind", "Marginal coverage",
              hline = 0.95)
```

```{r interval-simul-cov, fig.height=22}
interval_plot(metrics_summary, "mean_simul_cov", "mcse_simul_cov",
              "Mean simultaneous coverage by interval kind",
              "Simultaneous coverage", hline = 0.95)
```

```{r interval-ci-len, fig.height=22}
interval_plot(metrics_summary, "mean_ci_len", "mcse_ci_len",
              "Mean CI length by interval kind", "Mean CI length")
```

## Choosing `CI_sf`

The point of the sweep. `optimal_sf/` consumes it to pick a single value, so the
question here is which ratio gets simultaneous coverage closest to nominal —
simultaneous rather than marginal, because that is the guarantee the method
claims.

The plot is the distance from nominal, `|mean simultaneous coverage - 0.95|`, so
**lower is better and zero is the target**. The table underneath picks out the
arg-min per scenario, sample size and estimator, and reports the width that
choice costs.

```{r sf-distance-plot, fig.height=18}
sf_dist <- summary_hb %>%
  mutate(dist = abs(mean_simul_cov - 0.95))

sf_dist %>%
  ggplot(aes(x = CI_sf, y = dist, colour = model, group = model)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 2) +
  geom_line(linewidth = 0.3) +
  facet_grid(rows = vars(scenario), cols = vars(n), scales = "free_y") +
  scale_colour_paletteer_d("rcartocolor::Safe") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Distance from nominal simultaneous coverage",
       subtitle = "Lower is better; zero is exactly nominal",
       y = "|mean simultaneous coverage - 0.95|", x = "Subsampling ratio",
       colour = "Model")
```

```{r sf-best-table}
best_sf <- sf_dist %>%
  group_by(scenario, n, model) %>%
  slice_min(dist, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(scenario, n, model, CI_sf, mean_simul_cov, mean_marg_cov, mean_ci_len) %>%
  arrange(scenario, n, model)

kable(best_sf, digits = 4,
      col.names = c("Scenario", "n", "Model", "Best CI_sf",
                    "Mean simultaneous coverage", "Mean marginal coverage",
                    "Mean CI length"))
```

## Headline table

Every cell of the sweep for the per-unit bootstrap, arranged by scenario, sample
size, estimator and subsampling ratio.

```{r headline-table}
headline <- summary_hb %>%
  select(scenario, n, model, CI_sf, mean_marg_cov, mcse_marg_cov,
         mean_simul_cov, mcse_simul_cov, mean_ci_len, mcse_ci_len) %>%
  arrange(scenario, n, model, CI_sf)

kable(headline, digits = 4,
      col.names = c("Scenario", "n", "Model", "CI_sf", "Mean marginal coverage",
                    "MCSE (marg. cov.)", "Mean simultaneous coverage",
                    "MCSE (simul. cov.)", "Mean CI length", "MCSE (CI length)"))
```
