---
title: "Confidence intervals after multiple imputation: results"
author: "Ellie Van Vogt"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                       fig.width = 11, fig.height = 8)
```

```{r libraries, include=FALSE}
library(here)
library(patchwork)
library(purrr)
library(knitr)
source(here("R", "figures.R"))
```

## Once the covariates are imputed, how should the intervals be combined?

A single cell of the missing-data design, asking a narrower question than
`missing/continuous`. Every arm here uses the same mechanism and the same
handling method; what varies is the **pooling strategy**.

| | |
|---|---|
| scenarios | 1–5 (the reduced set — all five, unlike the other two studies) |
| n | 500 |
| type | `both` |
| prop | 0.3 |
| mechanism | MAR only |
| method | `multiple_imputation` only |
| runs | 100 |

Each of 50 imputed datasets gets its own half-sample bootstrap — 200 replicates
at a sample fraction of 0.5 — and `combine_mi_ci()` in `R/bootstrap_ci.R` then
pools them three ways so they can be compared:

| strategy | how |
|---|---|
| Pooled quantiles | empirical quantiles of the bootstrap replicates stacked across imputations |
| MI bootstrap (Rubin) | Rubin's rules — within + between variance, critical value averaged over the per-imputation maxima |
| Hybrid | one variance and one critical value from the stacked draws |

**Estimators**: `causal_forest`, `dr_random_forest`, `dr_semi_oracle`,
`dr_oracle` (`CI_MODELS` — this study has no SuperLearner arm).

Scenario numbering is the missing-data set's, not the main study's. Scenario 3
("Two HTE vars") appears here but not in `missing/continuous` or
`missing/binary`, whose grids run 1, 2, 4, 5 only.

### Two things to know before reading the figures

**The hybrid margin looks like a typo.** It computes
`sqrt(lambda_hat * S_star)` where the other two strategies and
`simultaneous_band()` use `sqrt(lambda_hat) * S_star` — i.e. it takes the square
root of the critical value. Preserved as written; if the hybrid intervals look
oddly narrow below, this is why.

**This study spells the mechanisms `AUX` / `AUX-Y`** where the others say
`MNAR` / `MNAR-Y`. Both are normalised in `R/missingness.R`, and
`MECHANISM_LABELS` maps both spellings to the same display label. Its grid only
ever runs MAR, so the unobserved-`U` branches never fire here.

```{r data}
# paths
path <- here()
res_path <- file.path(dirname(path), "results", "missing", "ci_example")

metrics <- readRDS(file.path(res_path, "cts_miss_ci_metrics.RDS"))

# unlike the chapter script this keeps all five scenarios - this study's grid
# runs 1:5
metrics <- metrics %>%
  apply_labels(MISS_SCENARIO_LABELS) %>%
  mutate(n = factor(n), prop = factor(prop)) %>%
  droplevels()

# sim_cov (simultaneous_coverage) is a genuine 0/1 indicator per run
# (whole-band coverage), so it's flagged for the exact binomial MCSE via
# `binomial=` - mar_cov (marginal_coverage) is a proportion over many units
# within a run, not Bernoulli at the run level, so it stays on the general
# sd(x)/sqrt(n) path.
metrics_summary <- summarise_metrics(
  metrics,
  c("scenario", "n", "type", "prop", "mechanism", "method", "model", "strategy"),
  cols = c(bias = "bias", ate_bias = "ate_bias",
           rel_ate_bias = "rel_ate_bias", rel_bias_cate = "rel_bias_cate",
           mse = "mse", rmse = "rmse",
           mae = "mae", corr = "corr", spearman = "spearman",
           sign_acc = "sign_acc", mar_cov = "marginal_coverage",
           sim_cov = "simultaneous_coverage", ci_len = "mean_ci_length"),
  binomial = c("sim_cov")
)
```

```{r plot-helpers, include=FALSE}
# per-run distribution: strategy on x, one colour per estimator. Only one
# mechanism here, so facet on scenario alone.
ci_box_plot <- function(metrics, metric, y_lab, hline = NULL,
                        facet_scales = "free_y") {
  p <- ggplot(metrics, aes(x = strategy, y = .data[[metric]], colour = model)) +
    geom_boxplot(fill = "transparent", outlier.shape = NA) +
    facet_grid(cols = vars(scenario), scales = facet_scales) +
    drf_scale() +
    drf_theme(rotate_x = TRUE) +
    labs(y = y_lab, x = "Pooling strategy", colour = "Model")
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p
}

# the summary counterpart. point_range_plot()'s faceting is hardcoded to
# mechanism x scenario, degenerate here, so it is overridden at the call site.
ci_sum_plot <- function(summary, metric, y_lab, hline = NULL,
                        facet_scales = "free_y") {
  p <- point_range_plot(summary, metric, y_lab, x = "strategy",
                        colour = "model", facet_scales = facet_scales) +
    facet_grid(cols = vars(scenario), scales = facet_scales) +
    labs(x = "Pooling strategy", colour = "Model")
  if (!is.null(hline)) {
    p <- p + geom_hline(yintercept = hline, linetype = "dashed")
  }
  p
}
```

Every section below pairs the per-run distribution (top) with the mean ± Monte
Carlo standard error across runs (bottom). Columns are the scenario, x the
pooling strategy, colour the estimator.

## Marginal coverage

The proportion of units their own interval covers, against the nominal 95%
(dashed line).

```{r mar-cov-plot, fig.height=14}
mar_cov_plot <- ci_box_plot(metrics, "marginal_coverage", "Marginal coverage",
                            hline = 0.95)
mar_cov_sum_plot <- ci_sum_plot(metrics_summary, "mar_cov", "Marginal coverage",
                                hline = 0.95)

(mar_cov_plot / mar_cov_sum_plot) + plot_layout(guides = "collect")
```

## Simultaneous coverage

Whether the band covers every unit at once — what the half-sample bootstrap is
constructed to control, and a much stricter target than marginal coverage.

Coverage here is evaluated **per-unit, on the sample's own covariates**. The
`confidence_intervals/` studies now also report a grid-based simultaneous
coverage over fixed query points (`build_query_grid()` in `R/dgm_scenarios.R`);
this study does not yet have that, so the two are not directly comparable. See
this study's README.

```{r sim-cov-plot, fig.height=14}
sim_cov_plot <- ci_box_plot(metrics, "simultaneous_coverage",
                            "Simultaneous coverage", hline = 0.95)
sim_cov_sum_plot <- ci_sum_plot(metrics_summary, "sim_cov",
                                "Simultaneous coverage", hline = 0.95)

(sim_cov_plot / sim_cov_sum_plot) + plot_layout(guides = "collect")
```

## CI length

The width that coverage is bought with. No reference line — there is no target
width, only the trade-off against the two coverage panels above.

```{r ci-len-plot, fig.height=14}
ci_len_plot <- ci_box_plot(metrics, "mean_ci_length", "CI length")
ci_len_sum_plot <- ci_sum_plot(metrics_summary, "ci_len", "CI length")

(ci_len_plot / ci_len_sum_plot) + plot_layout(guides = "collect")
```

## Point metrics

The sections below are the same point metrics the other missing-data results
files report. **They are identical across the three strategies within a model**:
all three pool the *intervals* around one shared `model_res$tau`, so the
triplicate points are the same number three times, not three disagreeing
estimates. Only the interval metrics above distinguish the pooling rules.

### Bias

Average bias in the estimated CATE (estimate minus truth), per unit and per run.

```{r bias-plot, fig.height=14}
bias_plot <- ci_box_plot(metrics, "bias", "Bias", hline = 0)
bias_sum_plot <- ci_sum_plot(metrics_summary, "bias", "Bias", hline = 0)

(bias_plot / bias_sum_plot) + plot_layout(guides = "collect")
```

### ATE bias

Bias of the *average* treatment effect (`mean(est) - mean(true)`).

```{r ate-bias-plot, fig.height=14}
ate_bias_plot <- ci_box_plot(metrics, "ate_bias", "ATE bias", hline = 0)
ate_bias_sum_plot <- ci_sum_plot(metrics_summary, "ate_bias", "ATE bias",
                                 hline = 0)

(ate_bias_plot / ate_bias_sum_plot) + plot_layout(guides = "collect")
```

### MSE

```{r mse-plot, fig.height=14}
mse_plot <- ci_box_plot(metrics, "mse", "MSE", hline = 0)
mse_sum_plot <- ci_sum_plot(metrics_summary, "mse", "MSE", hline = 0)

(mse_plot / mse_sum_plot) + plot_layout(guides = "collect")
```

### RMSE

`sqrt(mse)` — the same ranking as MSE, on the outcome's own scale.

```{r rmse-plot, fig.height=14}
rmse_plot <- ci_box_plot(metrics, "rmse", "RMSE", hline = 0)
rmse_sum_plot <- ci_sum_plot(metrics_summary, "rmse", "RMSE", hline = 0)

(rmse_plot / rmse_sum_plot) + plot_layout(guides = "collect")
```

### MAE

Mean absolute error — a robustness check on MSE, less sensitive to outlying runs.

```{r mae-plot, fig.height=14}
mae_plot <- ci_box_plot(metrics, "mae", "MAE", hline = 0)
mae_sum_plot <- ci_sum_plot(metrics_summary, "mae", "MAE", hline = 0)

(mae_plot / mae_sum_plot) + plot_layout(guides = "collect")
```

### Correlation with the true CATE

Pearson correlation between estimated and true CATEs. Undefined, and reported as
0, in the Null scenario — so that scenario is excluded here.

```{r corr-plot, fig.height=14}
corr_metrics <- filter(metrics, scenario != "Null")
corr_summary <- filter(metrics_summary, scenario != "Null")

corr_plot <- ci_box_plot(corr_metrics, "corr", "Pearson correlation")
corr_sum_plot <- ci_sum_plot(corr_summary, "corr", "Pearson correlation")

(corr_plot / corr_sum_plot) + plot_layout(guides = "collect")
```

### Spearman rank correlation with the true CATE

Less sensitive to outlying CATE estimates and to non-linear-but-monotone
miscalibration. Same Null exclusion.

```{r spearman-plot, fig.height=14}
spearman_plot <- ci_box_plot(corr_metrics, "spearman", "Spearman correlation")
spearman_sum_plot <- ci_sum_plot(corr_summary, "spearman",
                                 "Spearman correlation")

(spearman_plot / spearman_sum_plot) + plot_layout(guides = "collect")
```

### Sign accuracy

Proportion of units where the estimated CATE has the correct sign. 0.5 (dashed
line) is the chance level for a coin-flip on sign.

```{r sign-acc-plot, fig.height=14}
sign_acc_plot <- ci_box_plot(metrics, "sign_acc", "Proportion correct sign",
                             hline = 0.5)
sign_acc_sum_plot <- ci_sum_plot(metrics_summary, "sign_acc",
                                 "Proportion correct sign", hline = 0.5)

(sign_acc_plot / sign_acc_sum_plot) + plot_layout(guides = "collect")
```

## Missing estimates

`n_na` is a data-quality diagnostic, not a performance metric: the count of `NA`
CATE estimates within a run. Only cells with at least one missing estimate are
listed.

```{r na-table}
na_table <- metrics %>%
  group_by(scenario, model, strategy) %>%
  summarise(mean_n_na = mean(n_na, na.rm = TRUE),
            total_n_na = sum(n_na, na.rm = TRUE),
            .groups = "drop") %>%
  filter(total_n_na > 0) %>%
  arrange(scenario, model, strategy)

if (nrow(na_table) > 0) {
  kable(na_table, digits = 3,
        col.names = c("Scenario", "Model", "Strategy",
                      "Mean NA count per run", "Total NA count"))
} else {
  cat("No missing CATE estimates in any run.")
}
```

## Headline table

Coverage and width first — those are what this study is for — then the shared
point metrics.

```{r headline-table}
headline <- metrics_summary %>%
  select(scenario, model, strategy, mean_mar_cov, mean_sim_cov, mean_ci_len,
         mean_bias, mean_mse, mean_rmse, mean_mae, mean_sign_acc, mean_corr) %>%
  arrange(scenario, model, strategy)

kable(headline, digits = 4,
      col.names = c("Scenario", "Model", "Strategy", "Mean marginal coverage",
                    "Mean simultaneous coverage", "Mean CI length", "Mean bias",
                    "Mean MSE", "Mean RMSE", "Mean MAE", "Mean sign accuracy",
                    "Mean correlation"))
```
